### 2014-10-01 Paper/ms.tex
Paper

\documentclass[useAMS,usenatbib]{mn2e}

\voffset=-0.8in

% Packages:
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{dsfont}

% Bold symbols
\renewcommand{\btheta}{\{\theta_i\}}
\newcommand{\bx}{\{x_i\}}

% \onecolumn
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[]
{Fast Bayesian Inference for Exoplanet Discovery in Radial Velocity Data}

\author[Brewer]{%
  Brendon~J.~Brewer$^{1}$\thanks{bj.brewer@auckland.ac.nz},
  % Courtney?
  \medskip\\
  $^1$Department of Statistics, The University of Auckland, Private Bag 92019, Auckland 1142, New Zealand}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\date{To be submitted to MNRAS Letters}

\maketitle

\label{firstpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
\end{abstract}

\begin{keywords}
methods: data analysis
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

\vspace{-0.5cm}
\section*{Acknowledgements}
It is a pleasure to thank Courtney Donovan (Auckland) for many helpful
conversations about this subject. BJB is partially
supported by the Marsden Fund (Royal Society of New Zealand).

\begin{thebibliography}{99}
\end{thebibliography}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-10-01 MyDistribution.cpp
Biexponential prior for log periods

@@ -5,18 +5,16 @@
 
 using namespace DNest3;
 
-MyDistribution::MyDistribution(double x_min, double x_max,
-					double mu_min, double mu_max)
-:x_min(x_min)
-,x_max(x_max)
-,mu_min(mu_min)
-,mu_max(mu_max)
+MyDistribution::MyDistribution()
 {
 
 }
 
 void MyDistribution::fromPrior()
 {
+	// Cauchy prior centered on 5.901 = log(365 days).
+	center = 5.901 + tan(M_PI*(0.97*randomU() - 0.485));
+	width = 0.1 + 2.9*randomU();
 	mu = exp(tan(M_PI*(0.97*randomU() - 0.485)));
 }
 
@@ -24,13 +22,29 @@ double MyDistribution::perturb_parameters()
 {
 	double logH = 0.;
 
-	mu = log(mu);
-	mu = (atan(mu)/M_PI + 0.485)/0.97;
-	mu += randh();
-	wrap(mu, 0., 1.);
-	mu = tan(M_PI*(0.97*mu - 0.485));
-	mu = exp(mu);
+	int which = randInt(3);
 
+	if(which == 0)
+	{
+		center = (atan(center - 5.901)/M_PI + 0.485)/0.97;
+		center += randh();
+		wrap(center, 0., 1.);
+		center = 5.901 + tan(M_PI*(0.97*center - 0.485));
+	}
+	else if(which == 1)
+	{
+		width += 2.9*randh();
+		wrap(width, 0.1, 3.);
+	}
+	else
+	{
+		mu = log(mu);
+		mu = (atan(mu)/M_PI + 0.485)/0.97;
+		mu += randh();
+		wrap(mu, 0., 1.);
+		mu = tan(M_PI*(0.97*mu - 0.485));
+		mu = exp(mu);
+	}
 	return logH;
 }
 
@@ -42,18 +56,22 @@ double MyDistribution::perturb_parameters()
 
 double MyDistribution::log_pdf(const std::vector<double>& vec) const
 {
-	if(vec[0] < x_min || vec[0] > x_max || vec[1] < 0. ||
+	if(vec[1] < 0. ||
 			vec[2] < 0. || vec[2] > 2.*M_PI ||
 			vec[3] < 0.4 || vec[3] > 1. ||
 			vec[4] < 0. || vec[4] > 2.*M_PI)
 		return -1E300;
 
-	return -log(mu) - vec[1]/mu;
[…]

### 2014-10-01 MyDistribution.h

@@ -9,9 +9,8 @@
 class MyDistribution:public Distribution
 {
 	private:
-		// Limits
-		double x_min, x_max;
-		double mu_min, mu_max;
+		// Parameters of bi-exponential distribution for log-periods
+		double center, width;
 
 		// Mean of exponential distribution for amplitudes
 		double mu;
@@ -19,8 +18,7 @@ class MyDistribution:public Distribution
 		double perturb_parameters();
 
 	public:
-		MyDistribution(double x_min, double x_max,
-					double mu_min, double mu_max);
+		MyDistribution();
 
 		void fromPrior();
 

### 2014-10-01 MyModel.cpp

@@ -10,7 +10,7 @@ using namespace std;
 using namespace DNest3;
 
 MyModel::MyModel()
-:objects(5, 10, false, MyDistribution(0., 10., 1E-3, 1E3))
+:objects(5, 10, false, MyDistribution())
 ,mu(Data::get_instance().get_t().size())
 {
 

### 2014-10-01 display.py

@@ -4,7 +4,7 @@ import os
 data = loadtxt('exoplanet.txt')
 posterior_sample = atleast_2d(loadtxt('posterior_sample.txt'))
 
-hist(posterior_sample[:,1005], 100)
+hist(posterior_sample[:,1007], 100)
 xlabel('Number of Planets')
 show()
 

### 2014-10-01 Data.h
Set background prior based on data (slightly dodgy)

@@ -2,6 +2,7 @@
 #define _Data_
 
 #include <vector>
+#include <algorithm>
 
 class Data
 {
@@ -15,6 +16,8 @@ class Data
 		// Getters
 		const std::vector<double>& get_t() const { return t; }
 		const std::vector<double>& get_y() const { return y; }
+		double get_y_min() const { return *min_element(y.begin(), y.end()); }
+		double get_y_max() const { return *max_element(y.begin(), y.end()); }
 
 	// Singleton
 	private:

### 2014-10-01 MyModel.cpp

@@ -20,7 +20,8 @@ void MyModel::fromPrior()
 {
 	objects.fromPrior();
 	objects.consolidate_diff();
-	background = -20. + 40.*randomU();
+	background = Data::get_instance().get_y_min() +
+		(Data::get_instance().get_y_max() - Data::get_instance().get_y_min())*randomU();
 	sigma = exp(log(1E-3) + log(1E6)*randomU());
 	nu = exp(log(0.1) + log(1000.)*randomU());
 	calculate_mu();
@@ -93,8 +94,8 @@ double MyModel::perturb()
 		for(size_t i=0; i<mu.size(); i++)
 			mu[i] -= background;
 
-		background += 40.*randh();
-		wrap(background, -20., 20.);
+		background += (Data::get_instance().get_y_max() - Data::get_instance().get_y_min())*randh();
+		wrap(background, Data::get_instance().get_y_min(), Data::get_instance().get_y_max());
 
 		for(size_t i=0; i<mu.size(); i++)
 			mu[i] += background;

### 2014-10-01 display.py
Show periods

@@ -8,6 +8,12 @@ hist(posterior_sample[:,1007], 100)
 xlabel('Number of Planets')
 show()
 
+T = posterior_sample[:,1008:1018]
+T = T[T != 0].flatten()
+hist(T, 200, alpha=0.5)
+xlabel(r'$\ln$(Period/days)')
+show()
+
 t = linspace(data[:,0].min(), data[:,0].max(), 1000)
 
 saveFrames = False # For making movies

### 2014-10-01 display.py
Trim

@@ -10,7 +10,13 @@ show()
 
 T = posterior_sample[:,1008:1018]
 T = T[T != 0].flatten()
-hist(T, 200, alpha=0.5)
+# Trim
+s = sort(T)
+left, middle, right = s[0.25*len(s)], s[0.5*len(s)], s[0.75*len(s)]
+iqr = right - left
+s = s[logical_and(s > middle - 5*iqr, s < middle + 5*iqr)]
+
+hist(s, 200, alpha=0.5)
 xlabel(r'$\ln$(Period/days)')
 show()
 

### 2014-10-01 MyModel.cpp
cauchy based prior for sigma

@@ -22,7 +22,7 @@ void MyModel::fromPrior()
 	objects.consolidate_diff();
 	background = Data::get_instance().get_y_min() +
 		(Data::get_instance().get_y_max() - Data::get_instance().get_y_min())*randomU();
-	sigma = exp(log(1E-3) + log(1E6)*randomU());
+	sigma = exp(tan(M_PI*(0.97*randomU() - 0.485)));
 	nu = exp(log(0.1) + log(1000.)*randomU());
 	calculate_mu();
 }
@@ -80,8 +80,10 @@ double MyModel::perturb()
 	else if(randomU() <= 0.5)
 	{
 		sigma = log(sigma);
-		sigma += log(1E6)*randh();
-		sigma = mod(sigma - log(1E-3), log(1E6)) + log(1E-3);
+		sigma = (atan(sigma)/M_PI + 0.485)/0.97;
+		sigma += randh();
+		wrap(sigma, 0., 1.);
+		sigma = tan(M_PI*(0.97*sigma - 0.485));
 		sigma = exp(sigma);
 
 		nu = log(nu);

### 2014-10-01 Data.cpp
Use errorbars in data

@@ -23,12 +23,14 @@ void Data::load(const char* filename)
 	// Empty the vectors
 	t.clear();
 	y.clear();
+	sig.clear();
 
-	double temp1, temp2;
-	while(fin>>temp1 && fin>>temp2)
+	double temp1, temp2, temp3;
+	while(fin>>temp1 && fin>>temp2 && fin>>temp3)
 	{
 		t.push_back(temp1);
 		y.push_back(temp2);
+		sig.push_back(temp3);
 	}
 	cout<<"# Loaded "<<t.size()<<" data points from file "
 			<<filename<<"."<<endl;

### 2014-10-01 Data.h

@@ -7,7 +7,7 @@
 class Data
 {
 	private:
-		std::vector<double> t, y;
+		std::vector<double> t, y, sig;
 
 	public:
 		Data();
@@ -16,6 +16,7 @@ class Data
 		// Getters
 		const std::vector<double>& get_t() const { return t; }
 		const std::vector<double>& get_y() const { return y; }
+		const std::vector<double>& get_sig() const { return sig; }
 		double get_y_min() const { return *min_element(y.begin(), y.end()); }
 		double get_y_max() const { return *max_element(y.begin(), y.end()); }
 

### 2014-10-01 MyModel.cpp

@@ -22,7 +22,7 @@ void MyModel::fromPrior()
 	objects.consolidate_diff();
 	background = Data::get_instance().get_y_min() +
 		(Data::get_instance().get_y_max() - Data::get_instance().get_y_min())*randomU();
-	sigma = exp(tan(M_PI*(0.97*randomU() - 0.485)));
+	extra_sigma = exp(tan(M_PI*(0.97*randomU() - 0.485)));
 	nu = exp(log(0.1) + log(1000.)*randomU());
 	calculate_mu();
 }
@@ -79,12 +79,12 @@ double MyModel::perturb()
 	}
 	else if(randomU() <= 0.5)
 	{
-		sigma = log(sigma);
-		sigma = (atan(sigma)/M_PI + 0.485)/0.97;
-		sigma += randh();
-		wrap(sigma, 0., 1.);
-		sigma = tan(M_PI*(0.97*sigma - 0.485));
-		sigma = exp(sigma);
+		extra_sigma = log(extra_sigma);
+		extra_sigma = (atan(extra_sigma)/M_PI + 0.485)/0.97;
+		extra_sigma += randh();
+		wrap(extra_sigma, 0., 1.);
+		extra_sigma = tan(M_PI*(0.97*extra_sigma - 0.485));
+		extra_sigma = exp(extra_sigma);
 
 		nu = log(nu);
 		nu += log(1000.)*randh();
@@ -110,10 +110,13 @@ double MyModel::logLikelihood() const
 {
 	// Get the data
 	const vector<double>& y = Data::get_instance().get_y();
+	const vector<double>& sig = Data::get_instance().get_sig();
 
 	double logL = 0.;
+	double sigma;
 	for(size_t i=0; i<y.size(); i++)
 	{
+		sigma = sqrt(pow(sig[i], 2) + pow(extra_sigma, 2));
 		logL += gsl_sf_lngamma(0.5*(nu + 1.)) - gsl_sf_lngamma(0.5*nu)
 			- 0.5*log(M_PI*nu) - log(sigma)
 			- 0.5*(nu + 1.)*log(1. + pow((y[i] - mu[i])/sigma, 2)/nu);
@@ -162,7 +165,7 @@ void MyModel::print(std::ostream& out) const
 
 	for(size_t i=0; i<signal.size(); i++)
 		out<<signal[i]<<' ';
-	out<<sigma<<' '<<nu<<' ';
[…]

### 2014-10-01 MyModel.h

@@ -12,7 +12,7 @@ class MyModel:public DNest3::Model
 		RJObject<MyDistribution> objects;
 
 		double background;
-		double sigma; // Noise scale parameter
+		double extra_sigma; // Noise scale parameter
 		double nu; // Noise shape parameter
 
 		// The signal

### 2014-10-01 display.py

@@ -1,7 +1,7 @@
 from pylab import *
 import os
 
-data = loadtxt('exoplanet.txt')
+data = loadtxt('nu_oph.txt')
 posterior_sample = atleast_2d(loadtxt('posterior_sample.txt'))
 
 hist(posterior_sample[:,1007], 100)

### 2014-10-01 main.cpp

@@ -10,7 +10,7 @@ using namespace DNest3;
 int main(int argc, char** argv)
 {
 	// Load the data
-	Data::get_instance().load("exoplanet.txt");
+	Data::get_instance().load("nu_oph.txt");
 
 	Lookup::get_instance().load();
 

### 2014-10-01 Paper/ms.tex
Two paragraphs

@@
 \section{Introduction}
+The number of known extrasolar planets has exploded in the last two
+decades \citep{}. These discoveries have been driven by improvements in
+all of the different techniques used to detect and characterise exoplanets,
+including the radial velocity method \citep[e.g.][]{},
+the transit method \citep[e.g.]{},
+and gravitational microlensing \citep[e.g.][]{}. This paper focuses on the
+radial velocity method.
+
+Many authors have considered the problem of inferring the properties of an
+exoplanetary system from observational data. With radial velocity data,
+the expected signal due to an exoplanet is periodic, and the goal is to
+infer the number of planets in the system, as well as their properties such
+as orbital periods and eccentricities. Many
+different techniques have been proposed. These techniques fall into two main
+classes: i) those based on periodograms \citep[e.g.][]{}, ii) those based on
+Bayesian inference, to describe the uncertainties probabilistically.
 

### 2014-10-03 Paper/ms.tex
orbits
Parameter table

@@
 
+\section{Inference}
+
+The unknown parameters are:
+\begin{eqnarray}
+\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\phi}\}_{i=1}^N, y_0, s, \nu\right\}
+\end{eqnarray}
+where $\boldsymbol{\alpha} = \{\mu_T, \sigma_T, \mu_A\}$ are the
+hyperparameters, and $\phi_i = \{\}$ are the properties of planet $i$.
+
+In previous studies, the inference was performed by doing separate runs, one
+for each possible value of $N$, and calculating the marginal likelihood
+$p(\boldsymbol{D} | N)$ as a function of $N$.
+
+\begin{table*}
+\begin{tabular}{|c|c|c|}
+\hline
+Quantity	&	Meaning		& Prior\\
+\hline
+$N$		& Number of planets	& Discrete Uniform$(0, 10)$\\
+\hline
+		&	Hyperparameters	&	\\
+$\mu_T$		&			&	\\
+\end{tabular}
+\end{table*}
+
+\section{Orbit Simulations}
+
+The equations of motion were
+\begin{eqnarray}
+\ddot{x} &=& -\frac{x}{r^3} \\
+\ddot{y} &=& -\frac{y}{r^3} \\
+\end{eqnarray}
+where $r = \sqrt{x^2 + y^2}$. The initial conditions
+$\{x(0), y(0), \dot{x}(0), \dot{y}(0)\} = \{1, 0, 0, v\}$.
+
 \vspace{-0.5cm}

### 2014-10-09 Paper/ms.tex
nu oph references

@@
 
+\section{Reanalysis of $\nu$ Oph}
+The $\nu$ Oph system is generally accepted to have two confirmed planets
+\citep[e.g.][]{2011AIPC.1331..102Q, 2012PASJ...64..135S, fengji}, with periods
+of $530.3$ and $3190$ days.
+
 \vspace{-0.5cm}
@@
 \begin{thebibliography}{99}
+\bibitem[\protect\citeauthoryear{Hou, Goodman, 
+\& Hogg}{2014}]{fengjo} Hou F., Goodman J., Hogg D.~W., 2014, arXiv, arXiv:1401.6128
+
+\bibitem[\protect\citeauthoryear{Quirrenbach, Reffert, 
+\& Bergmann}{2011}]{2011AIPC.1331..102Q} Quirrenbach A., Reffert S., Bergmann C., 2011, AIPC, 1331, 102
+
+\bibitem[\protect\citeauthoryear{Sato et al.}{2012}]{2012PASJ...64..135S} 
+Sato B., et al., 2012, PASJ, 64, 135 
 \end{thebibliography}

### 2014-10-09 Paper/ms.tex
A few more refs

@@
 
-In previous studies, the inference was performed by doing separate runs, one
+The number of orbiting planets, $N$, is usually of significant interest.
+Most authors have carried out the inference for $N$ by doing separate runs, one
 for each possible value of $N$, and calculating the marginal likelihood
-$p(\boldsymbol{D} | N)$ as a function of $N$.
+$p(\boldsymbol{D} | N)$ as a function of $N$
+\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}.
 
@@
 \begin{thebibliography}{99}
+\bibitem[\protect\citeauthoryear{Feroz 
+\& Hobson}{2014}]{2014MNRAS.437.3540F} Feroz F., Hobson M.~P., 2014, MNRAS, 437, 3540
+
+\bibitem[\protect\citeauthoryear{Gregory}{2011}]{2011MNRAS.415.2523G} 
+Gregory P.~C., 2011, MNRAS, 415, 2523
+
 \bibitem[\protect\citeauthoryear{Hou, Goodman, 
-\& Hogg}{2014}]{fengjo} Hou F., Goodman J., Hogg D.~W., 2014, arXiv, arXiv:1401.6128
+\& Hogg}{2014}]{fengji} Hou F., Goodman J., Hogg D.~W., 2014, arXiv, arXiv:1401.6128
 

### 2014-10-10 Paper/ms.tex
nu oph figure
Courtney

@@
   Brendon~J.~Brewer$^{1}$\thanks{bj.brewer@auckland.ac.nz},
-  % Courtney?
+  Courtney P. Donovan$^{1}$
   \medskip\\
@@
 
+The posterior distribution for $N$, the number of planets, is shown in
+Figure~\ref{fig:nu_oph_N}.
+
+\begin{figure}
+\includegraphics[scale=0.5]{Figures/nu_oph_N.eps}
+\caption{The posterior distribution for the number of planets $N$ orbiting
+$\nu$ Oph.\label{fig:nu_oph_N}}
+\end{figure}
+
 \vspace{-0.5cm}
 \section*{Acknowledgements}
-It is a pleasure to thank Courtney Donovan (Auckland) for many helpful
-conversations about this subject. BJB is partially
+BJB is partially
 supported by the Marsden Fund (Royal Society of New Zealand).

### 2014-10-10 Paper/ms.tex
periods figure

@@
 
+The posterior distribution for the logarithms of the periods is shown in
+Figure~\ref{fig:nu_oph_periods}. Because of the label-switching degeneracy,
+the posterior samples for all periods were combined to make this figure, which
+shows three prominent peaks. Defining the log-periods by
+$S_i =  \log_{10}(T/\textnormal{1 day})$,
+Figure~\ref{fig:nu_oph_periods} is a Monte Carlo representation
+of the mixture distribution
+\begin{eqnarray}
+f(S) &=& \sum_{N=1}^{10} \frac{1}{N}\sum_{i=1}^N p(S_i | N, \boldsymbol{D}).
+\end{eqnarray}
+
 \begin{figure}
@@
 $\nu$ Oph.\label{fig:nu_oph_N}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.5]{Figures/nu_oph_periods.eps}
+\caption{The posterior distribution for the orbital periods in the $\nu$ Oph
+system.\label{fig:nu_oph_periods}}
 \end{figure}

### 2014-10-10 Paper/ms.tex
An abstract

@@
 \begin{abstract}
+Inferring the number of planets $N$ in an exoplanetary system from radial velocity
+(RV) data is a challenging task. Typically, the posterior distribution for the
+planet parameters is multimodal and may have other awkward features. In
+addition, when $N$ is unknown, the marginal likelihood (or evidence), as a
+function of $N$, is required. In this paper we propose an alternative
+approach using a trans-dimensional Markov Chain Monte Carlo method within
+Nested Sampling. We also implement the likelihood function using a look-up
+table of pre-computed elliptical orbits.
+The posterior distribution for $N$ can be obtained with a
+single run, which takes $\sim$ 10 minutes on a typical modern desktop computer.
+We apply the method to $\nu$ Oph and XXX, and find moderate evidence
+(posterior probability of existence = 70\%) for a third planet around
+$\nu$ Oph with an orbital period of 75.7 $\pm$ 1.8 days.
 \end{abstract}

### 2014-10-10 Paper/ms.tex
Apple pie and bayes

@@
 % Bold symbols
-\renewcommand{\btheta}{\{\theta_i\}}
+\renewcommand{\btheta}{\boldsymbol{\theta}}
+\newcommand{\bdata}{\boldsymbol{D}}
 \newcommand{\bx}{\{x_i\}}
@@
 the transit method \citep[e.g.]{},
-and gravitational microlensing \citep[e.g.][]{}. This paper focuses on the
+and gravitational microlensing
+\citep[e.g.][]{2014ApJ...785..155B, 2014ApJ...790...14Y}. This paper focuses on the
 radial velocity method.
@@
 different techniques have been proposed. These techniques fall into two main
-classes: i) those based on periodograms \citep[e.g.][]{}, ii) those based on
-Bayesian inference, to describe the uncertainties probabilistically.
+classes: i) those based on periodograms \citep[e.g.][]{},
+and ii) those based on model fitting in the Bayesian inference framework,
+to describe the uncertainties probabilistically.
 
 \section{Inference}
+Bayesian inference is the use of probability theory to describe uncertainty
+\citep{}. In this framework, we approach data analysis problems by first
+constructing a {\it hypothesis space}, which is the set of possible answers
+to the problem we are considering. Normally, this is the set of possible
+values of a vector of parameters $\btheta$ whose values we want
+to know. We then assign probability distributions
+called the {\it prior} and the {\it sampling distribution}. The prior
+distribution $p(\btheta)$ describes
+our initial uncertainty about which values of the parameters
+$\btheta$ are plausible, and the
+sampling distribution $p(\bdata | \btheta)$ describes
+our initial uncertainty about the data set we're going to observe, as a
+function of the unknown parameters $\btheta$.
+When the data is known, our state of knowledge about the parameter is
+updated from the prior $p(\theta)$ to the posterior distribution given by
+Bayes' rule:
+\begin{eqnarray}
+p(\btheta | \bdata) &=&
+\frac{p(\btheta)p(\bdata | \btheta)}
+{p(\bdata)}
+\end{eqnarray}
+where the denominator, often called the {\it evidence} or
+{\it marginal likelihood}, is given by:
+\begin{eqnarray}
+p(\bdata) = \int p(\btheta) p(\bdata | \btheta) d^N \btheta
+\end{eqnarray}
 
@@
 for each possible value of $N$, and calculating the marginal likelihood
-$p(\boldsymbol{D} | N)$ as a function of $N$
+$p(\bdata | N)$ as a function of $N$
 \citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}.
@@
 \begin{eqnarray}
-f(S) &=& \sum_{N=1}^{10} \frac{1}{N}\sum_{i=1}^N p(S_i | N, \boldsymbol{D}).
+f(S) &=& \sum_{N=1}^{10} \frac{1}{N}\sum_{i=1}^N p(S_i | N, \bdata).
 \end{eqnarray}
@@
 \begin{thebibliography}{99}
+
+\bibitem[\protect\citeauthoryear{Bennett et 
+al.}{2014}]{2014ApJ...785..155B} Bennett D.~P., et al., 2014, ApJ, 785, 155
+
 \bibitem[\protect\citeauthoryear{Feroz 
@@
 Sato B., et al., 2012, PASJ, 64, 135
+
+\bibitem[\protect\citeauthoryear{Yee et al.}{2014}]{2014ApJ...790...14Y} 
+Yee J.~C., et al., 2014, ApJ, 790, 14
+
 \end{thebibliography}

### 2014-10-14 Paper/ms.tex
Orbit writing

--- 2014-10-10 Paper/ms.tex
@@
 \section{Orbit Simulations}
+The expected (noise-free) signal due to an exoplanet is periodic, but
+non-sinusoidal when the orbit is not perfectly circular. The expected
+shape of the variations can be obtained from mechanical considerations.
+To save time, we pre-computed the properties of orbits as a function of
+ellipticity.
 
-The equations of motion were
+Consider a test particle moving in the $x$-$y$ plane under the influence of a
+point mass at the origin. The equations of motion for the particle are:
 \begin{eqnarray}
-\ddot{x} &=& -\frac{x}{r^3} \\
-\ddot{y} &=& -\frac{y}{r^3} \\
+\frac{d^2x}{dt^2} &=& -\frac{x}{r^3} \\
+\frac{d^2y}{dt^2} &=& -\frac{y}{r^3} \\
 \end{eqnarray}

### 2014-10-14 Paper/ms.tex
v, acknowledgements

@@
 \end{eqnarray}
-where $r = \sqrt{x^2 + y^2}$. The initial conditions
-$\{x(0), y(0), \dot{x}(0), \dot{y}(0)\} = \{1, 0, 0, v\}$.
+where $r = \sqrt{x^2 + y^2}$. The ODEs can be solved for any given set of
+initial conditions for $x$, $y$, $\frac{dx}{dt}$, and $\frac{dy}{dt}$ to yield
+the shape of the orbit. To simulate orbits
+of different ellipticities, we set the initial position to $(1, 0)$, and
+the initial velocity to $(0, v)$ where $v \in [0.4, 1]$.
+If $v=1$, the orbit is circular and as $v$ decreases the orbit becomes more
+elliptical.
 
@@
 \section*{Acknowledgements}
-BJB is partially
-supported by the Marsden Fund (Royal Society of New Zealand).
+It is a pleasure to thank Fengji Hou and David Hogg (NYU) for inspiring me to
+finally work on this problem, and Phil Gregory (UBC) for writing so many
+interesting papers on it. This work was supported by a Marsden Fast Start
+grant from the Royal Society of
+New Zealand.
 

### 2014-10-17 Paper/ms.tex
Orbits
Words

@@
 
+We used hierarchical priors because...
+
 \section{Orbit Simulations}
@@
 If $v=1$, the orbit is circular and as $v$ decreases the orbit becomes more
-elliptical.
+elliptical. For trial values of $v$ ranging from 0.4 to 1 in steps of 0.005,
+we simulated an orbit using a leapfrog integrator, and saved the
+velocities $\frac{dx}{dt}$ and $\frac{dy}{dt}$ as a function of time to disk.
+These saved orbits were used as a lookup table for constructing the expected
+signal $y(t)$ due to a single planet.
+Because of the initial conditions used,
+the simulated orbits were all horizontally
+aligned.
 

### 2014-10-19 Paper/ms.tex
Gliese 581 figures
Abstract tweaks

@@
 approach using a trans-dimensional Markov Chain Monte Carlo method within
-Nested Sampling. We also implement the likelihood function using a look-up
-table of pre-computed elliptical orbits.
-The posterior distribution for $N$ can be obtained with a
-single run, which takes $\sim$ 10 minutes on a typical modern desktop computer.
-We apply the method to $\nu$ Oph and XXX, and find moderate evidence
+Nested Sampling. The posterior distribution for $N$ can be obtained with a
+single run, which takes $\sim$ 15 minutes on a typical modern desktop computer.
+We apply the method to $\nu$ Oph and Gliese 581, finding moderate evidence
 (posterior probability of existence = 70\%) for a third planet around
@@
 
+\begin{figure}
+\includegraphics[scale=0.5]{Figures/gliese581_N.eps}
+\caption{The posterior distribution for the number of planets $N$ orbiting
+Gliese 581.\label{fig:gliese581_N}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.5]{Figures/gliese581_periods.eps}
+\caption{The posterior distribution for the orbital periods in the Gliese 581
+system.\label{fig:gliese581_periods}}
+\end{figure}
+
 \vspace{-0.5cm}

### 2014-10-19 Paper/ms.tex
Prior table

@@
 \begin{table*}
-\begin{tabular}{|c|c|c|}
+\begin{tabular}{|l|l|l|}
 \hline
@@
 \hline
-$N$		& Number of planets	& Discrete Uniform$(0, 10)$\\
+{\bf Hyperparameters}	&	\\
+$\mu_T$		&	Median orbital period	& $\ln(\mu_T) \sim$ Cauchy$(5.9, 1)T(-15.3, 27.1)$\\
+$w_T$		&	Diversity of orbital periods & $w_T \sim$ Uniform$(0.1, 3)$\\
+$\mu_A$		&	Mean amplitude	& $\ln(\mu_A) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
 \hline
-		&	Hyperparameters	&	\\
-$\mu_T$		&			&	\\
+{\bf Planet Parameters}\\
+$T_i$		&	Orbital period	&	$\ln(T_i) \sim $Biexponential($\ln(\mu_T), w_T$)\\
+$A_i$		&	Amplitude of signal	& Exponential$(\mu_A)$\\
+$\phi_i$	&	Phase of signal	&	Uniform($0, 2\pi$)\\
+$v_i$		&	Initial speed of orbit	&	Uniform$(0.4, 1)$\\
+$\theta_i$	&	Viewing angle	&	Uniform$(0, 2\pi)$\\
+\hline
+{\bf Other}\\
+$N$		& Number of planets	& Uniform$(\{0, 1, ..., N_{\rm max}\})$\\
+$\sigma$	& Standard deviation of extra noise	& $\ln(\sigma) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
+$\nu$		& Shape parameter for $t$-distribution for noise & $\ln(\nu) \sim$ Uniform$(\ln(0.01), \ln(1000))$
 \end{tabular}

### 2014-10-21 Paper/ms.tex
Eccentricity
Gliese 581 in abstract

@@
 (RV) data is a challenging task. Typically, the posterior distribution for the
-planet parameters is multimodal and may have other awkward features. In
+planet parameters, such as orbital periods,
+is multimodal and may have other awkward features. In
 addition, when $N$ is unknown, the marginal likelihood (or evidence), as a
@@
 (posterior probability of existence = 70\%) for a third planet around
-$\nu$ Oph with an orbital period of 75.7 $\pm$ 1.8 days. 
+$\nu$ Oph with an orbital period of 75.7 $\pm$ 1.8 days. The results also
+suggest Gliese 581 hosts many (7-15) planets (or causes of other periodic
+signals), but only four have well determined orbital periods. 
 \end{abstract}
@@
 
+Although we parameterise the eccentricity using the initial velocity $v$,
+the standard eccentricity $\epsilon$ can be recovered using the relation
+$\epsilon = 1 - v^2$.
+
 \section{Reanalysis of $\nu$ Oph}

### 2014-10-21 Paper/ms.tex
More references

@@
 and ii) those based on model fitting in the Bayesian inference framework,
-to describe the uncertainties probabilistically.
+to describe the uncertainties probabilistically
+\citep[e.g.][]{2011MNRAS.410...94G, 2014MNRAS.437.3540F,
+2011A&A...528L...5T, fengji}.
 
@@
 
+\bibitem[\protect\citeauthoryear{Gregory}{2011}]{2011MNRAS.410...94G} 
+Gregory P.~C., 2011, MNRAS, 410, 94
+
 \bibitem[\protect\citeauthoryear{Hou, Goodman, 
@@
 
+\bibitem[\protect\citeauthoryear{Tuomi}{2011}]{2011A&A...528L...5T} Tuomi M., 2011, A\&A, 528, L5
+
 \bibitem[\protect\citeauthoryear{Yee et al.}{2014}]{2014ApJ...790...14Y} 

### 2014-10-21 Paper/ms.tex
Writing

--- 2014-10-19 Paper/ms.tex
@@
 non-sinusoidal when the orbit is not perfectly circular. The expected
-shape of the variations can be obtained from mechanical considerations.
+shape $m(t)$ of the variations can be obtained from mechanical considerations.
 To save time, we pre-computed the properties of orbits as a function of
@@
 Consider a test particle moving in the $x$-$y$ plane under the influence of a
-point mass at the origin. The equations of motion for the particle are:
+point mass at the origin. The motion of the test particle represents the
+reflex motion of the host star orbiting around the center of mass of the
+system. The equations of motion for the particle are:
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $r = \sqrt{x^2 + y^2}$. The ODEs can be solved for any given set of
+where $r = \sqrt{x^2 + y^2}$. These ODEs can be solved for any given set of
 initial conditions for $x$, $y$, $\frac{dx}{dt}$, and $\frac{dy}{dt}$ to yield
@@
 signal $y(t)$ due to a single planet.
-Because of the initial conditions used,
-the simulated orbits were all horizontally
-aligned.
+Because of the initial conditions, the simulated orbits were all horizontally
+aligned. If the observer is located on the $x$-axis a large distance
+from the origin, they will measure $m(t) = \dot{x}(t)$. However, if the
+observer is located at
+
+Although we parameterised the orbits using the initial velocity $v$,
+the standard eccentricity $\epsilon$ can be recovered using the relation
+$\epsilon = 1 - v^2$.
 

### 2014-10-21 Paper/ms.tex
Marginal likelihoods, words on Gliese 581

@@
 \begin{eqnarray}
-p(\bdata) = \int p(\btheta) p(\bdata | \btheta) d^N \btheta
+\mathcal{Z} = p(\bdata) = \int p(\btheta) p(\bdata | \btheta) d^N \btheta
 \end{eqnarray}
@@
 
+The marginal likelihood was $\ln(\mathcal{Z}) \approx -217.04$.
+
 \begin{figure}
@@
 
+\section{Reanalysis of Gliese 581}
+
+The red dwarf star Gliese 581 is thought to host several planets. Exactly
+how many is a matter of considerable debate. According to \citet{}, there
+are two planets whose existence is not doubted, two more whose existence is
+doubted by some, and another two thought not to exist by most groups.
+
+Our model cannot distinguish between oscillations due to planets and
+oscillations due to stellar activity: any oscillations found in the dataset
+will be described as ``planets'' by the model.
+
 \begin{figure}
@@
 \end{figure}
+
+The marginal likelihood was $\ln(\mathcal{Z}) \approx -618.04$.
 

### 2014-10-21 Paper/ms.tex
Data in table

@@
 {\bf Planet Parameters}\\
-$T_i$		&	Orbital period	&	$\ln(T_i) \sim $Biexponential($\ln(\mu_T), w_T$)\\
+$T_i$		&	Orbital period	&	$\ln(T_i) \sim $ Biexponential($\ln(\mu_T), w_T$)\\
 $A_i$		&	Amplitude of signal	& Exponential$(\mu_A)$\\
@@
 $N$		& Number of planets	& Uniform$(\{0, 1, ..., N_{\rm max}\})$\\
-$\sigma$	& Standard deviation of extra noise	& $\ln(\sigma) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
-$\nu$		& Shape parameter for $t$-distribution for noise & $\ln(\nu) \sim$ Uniform$(\ln(0.01), \ln(1000))$
+$\sigma_{\rm extra}$	& ``Extra noise'' parameter	& $\ln(\sigma) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
+$\nu$		& Shape parameter for $t$-distribution for noise & $\ln(\nu) \sim$ Uniform$(\ln(0.01), \ln(1000))$\\
+\hline
+{\bf Data}\\
+$Y_i$		& Radial velocity measurements	&
+		Student-$t\left(m(t_i), \sqrt{\sigma_i^2 + \sigma_{\rm extra}^2}, \nu\right)$
 \end{tabular}

### 2014-10-21 Paper/ms.tex
Theta

@@
 $\nu$ Oph with an orbital period of 75.7 $\pm$ 1.8 days. The results also
-suggest Gliese 581 hosts many (7-15) planets (or causes of other periodic
+suggest Gliese 581 hosts many (7-15) planets (or other causes of other periodic
 signals), but only four have well determined orbital periods. 
@@
 from the origin, they will measure $m(t) = \dot{x}(t)$. However, if the
-observer is located at
+observer is located at an angle $\theta$ with respect to the $x$-axis, then
+the radial velocity measured will instead be
+$m(t) = \cos(\theta)\dot{x}(t) + \sin(\theta)\dot{y}(t)$.
+Since the our orientation with respect to the orbits is unknown, each planet
+requires a ``viewing angle'' parameter $\theta$.
 

### 2014-10-22 Paper/ms.tex
Papers
More refs

@@
 Bayesian inference is the use of probability theory to describe uncertainty
-\citep{}. In this framework, we approach data analysis problems by first
+\citep{sivia, ohagan}. In this framework, we approach data analysis problems by first
 constructing a {\it hypothesis space}, which is the set of possible answers
@@
 \citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}.
+
+Diffusive Nested Sampling \citep{dnest} is a variant of
+Nested Sampling \citep{skilling}. RJObject \citep{rjobject}
 
@@
 
+\bibitem[\protect\citeauthoryear{Brewer, P{\'a}rtay,
+\& Cs{\'a}nyi}{2011}]{dnest} Brewer B.~J., P{\'a}rtay L.~B., Cs{\'a}nyi G., 2011,
+Statistics and Computing, 21, 4, 649-656. arXiv:0912.2380
+
+\bibitem[\protect\citeauthoryear{Brewer}{2014}]{rjobject} Brewer, B. J., 2014,
+submitted.
+
 \bibitem[\protect\citeauthoryear{Feroz 
@@
 
+\bibitem[\protect\citeauthoryear{O'Hagan and Forster}{2004}]{ohagan}
+O'Hagan, A., Forster,~J., 2004, Bayesian inference. London: Arnold.
+
 \bibitem[\protect\citeauthoryear{Quirrenbach, Reffert, 
@@
 Sato B., et al., 2012, PASJ, 64, 135
+
+\bibitem[\protect\citeauthoryear{Sivia \& Skilling}{2006}]{sivia} Sivia, 
+D.~ S., Skilling, J., 2006, Data Analysis: A Bayesian Tutorial, 2nd 
+Edition, Oxford University Press
+
+\bibitem[\protect\citeauthoryear{Skilling}{2006}]{skilling} Skilling, 
+J., 2006, ``Nested Sampling for General Bayesian Computation'', Bayesian 
+Analysis 4, pp. 833-860
 

### 2014-10-25 Paper/ms.tex
Ref

@@
 and gravitational microlensing
-\citep[e.g.][]{2014ApJ...785..155B, 2014ApJ...790...14Y}. This paper focuses on the
-radial velocity method.
+\citep[e.g.][]{2014ApJ...785..155B, 2014ApJ...790...14Y}.
 
-Many authors have considered the problem of inferring the properties of an
-exoplanetary system from observational data. With radial velocity data,
+The problem of inferring the properties of an
+exoplanetary system from observational data can be challenging.
+With radial velocity data,
 the expected signal due to an exoplanet is periodic, and the goal is to
@@
 as orbital periods and eccentricities. Many
-different techniques have been proposed. These techniques fall into two main
-classes: i) those based on periodograms \citep[e.g.][]{},
+different techniques have been proposed for doing this.
+These techniques fall into two main
+classes: i) those based on periodograms \citep[e.g.][]{2009A&A...496..577Z},
 and ii) those based on model fitting in the Bayesian inference framework,
@@
 
+\bibitem[\protect\citeauthoryear{Zechmeister \& K{\"u}rster}{2009}]{2009A&A...496..577Z} Zechmeister M., K{\"u}rster M., 2009, A\&A, 496, 577
+
 \end{thebibliography}

### 2014-10-26 Paper/ms.tex
More talk

@@
 addition, when $N$ is unknown, the marginal likelihood (or evidence), as a
-function of $N$, is required. In this paper we propose an alternative
+function of $N$, is required. Rather than doing separate runs with different
+values of $N$, we propose an alternative
 approach using a trans-dimensional Markov Chain Monte Carlo method within
@@
 \begin{eqnarray}
-\mathcal{Z} = p(\bdata) = \int p(\btheta) p(\bdata | \btheta) d^N \btheta
+\mathcal{Z} = p(\bdata) = \int p(\btheta) p(\bdata | \btheta) d^n \btheta
 \end{eqnarray}
+where the integral is over the entire $n$-dimensional parameter space.
+
+The number of orbiting planets, $N$, is an important parameter.
+To calculate the posterior distribution for $N$, most authors
+consider various trial values of $N$, and calculate the marginal likelihood
+\begin{eqnarray}
+p(\bdata | N) &=& \int p(\btheta | N) p(\bdata | \btheta, N) \, d^n \theta
+\end{eqnarray}
+for each possible value of $N$
+\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}. Popular
+methods for calculating the marginal likelihood are Nested Sampling
+\citep{skilling} and ideas related to thermodynamic integration
+\citep[e.g.][]{neal}.
 
@@
 hyperparameters, and $\phi_i = \{\}$ are the properties of planet $i$.
-
-The number of orbiting planets, $N$, is usually of significant interest.
-Most authors have carried out the inference for $N$ by doing separate runs, one
-for each possible value of $N$, and calculating the marginal likelihood
-$p(\bdata | N)$ as a function of $N$
-\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}.
 
@@
 
+\bibitem[\protect\citeauthoryear{Neal}{2001}]{neal}
+Neal, R.~M., 2001, Statistics and Computing 11.2, 125-139.
+
 \bibitem[\protect\citeauthoryear{O'Hagan and Forster}{2004}]{ohagan}

### 2014-10-26 Paper/ms.tex
Posterior for N (traditional way)

@@
 \end{eqnarray}
-where the denominator, often called the {\it evidence} or
-{\it marginal likelihood}, is given by:
+where $p(\bdata | \btheta)$ is called the likelihood, once the actual dataset
+is substituted in.
+The denominator, often called the {\it evidence} or
+{\it marginal likelihood}, is given by the expected value of the likelihood
+with respect to the prior:
 \begin{eqnarray}
@@
 for each possible value of $N$
-\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}. Popular
+\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}.
+The posterior distribution for $N$ can then be found straightforwardly by
+using Bayes' rule with $N$ as the only unknown parameter:
+\begin{eqnarray}
+p(N | \bdata) &=& \frac{p(N)p(\bdata | N)}{\sum_N p(N)p(\bdata | N)}.
+\end{eqnarray}
+Popular
 methods for calculating the marginal likelihood are Nested Sampling

### 2014-10-26 Paper/ms.tex
Keywords
Forceful words

@@
 \begin{keywords}
-methods: data analysis
+stars: planetary systems --- techniques: radial velocities ---
+methods: data analysis --- methods:statistical
 \end{keywords}
@@
 
+\subsection{Inferring the number of planets}
 The number of orbiting planets, $N$, is an important parameter.
@@
 \begin{eqnarray}
-p(\bdata | N) &=& \int p(\btheta | N) p(\bdata | \btheta, N) \, d^n \theta
+p(\bdata | N) &=& \int p(\btheta | N) p(\bdata | \btheta, N) \, d^n \btheta
 \end{eqnarray}
 for each possible value of $N$
-\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}.
+\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}, marginalising
+over all other model parameters.
 The posterior distribution for $N$ can then be found straightforwardly by
@@
 \citep[e.g.][]{neal}.
+
+This traditional approach can be very time consuming.
+Methods for calculating the marginal likelihood are
+already more intensive than standard MCMC methods for sampling the posterior,
+because they usually involve a sequence of probability distributions
+(e.g. the constrained priors in Nested Sampling, or the annealed distributions
+in thermodynamic integration) rather than a single distribution (the posterior).
+This intensive process needs to be run many times, for $N=0$, $N=1$, $N=2$, and
+so on.
+
+The traditional approach also contradicts standard practice in Bayesian
+computation. Imagine we are trying to compute the posterior distribution for
+a parameter $a$ in the presence of a nuisance parameter $b$. This is solved
+by exploring the joint posterior for $a$ and $b$, and then only looking at the
+generated values of $a$. Nobody would suggest the wasteful alternative
+of using a discrete grid of possible $a$ values and doing an entire Nested
+Sampling run for each, to get the marginal likelihood as a function of $a$.
+Yet when the hypothesis space for $a$ is already discrete, the wasteful
+alternative somehow becomes standard practice.
 

### 2014-10-26 Paper/ms.tex
Gliese 581 planets

@@
 The red dwarf star Gliese 581 is thought to host several planets. Exactly
-how many is a matter of considerable debate. According to \citet{}, there
-are two planets whose existence is not doubted, two more whose existence is
-doubted by some, and another two thought not to exist by most groups.
+how many is a matter of considerable debate. According to
+\citet{2014Sci...345..440R}, there are two planets
+(b and c, with periods 5.36 and 12.91 days respectively)
+whose existence is generally
+accepted, two more
+(d and e, with periods 66 and 3.15 days respectively)
+whose existence is mostly accepted, and another two
+(f and g, with periods 433 and 36.5 days respectively)
+whose existence is generally doubted.
 
@@
 
+\bibitem[\protect\citeauthoryear{Robertson et 
+al.}{2014}]{2014Sci...345..440R} Robertson P., Mahadevan S., Endl M., Roy 
+A., 2014, Sci, 345, 440
+
 \bibitem[\protect\citeauthoryear{Sato et al.}{2012}]{2012PASJ...64..135S} 

### 2014-10-26 Paper/ms.tex
Wording tweaks

@@
 
-The traditional approach also contradicts standard practice in Bayesian
+The traditional approach also contradicts the normal approach to Bayesian
 computation. Imagine we are trying to compute the posterior distribution for
-a parameter $a$ in the presence of a nuisance parameter $b$. This is solved
+a parameter $a$ in the presence of a nuisance parameter $b$. This is usually solved
 by exploring the joint posterior for $a$ and $b$, and then only looking at the
@@
 Sampling run for each, to get the marginal likelihood as a function of $a$.
-Yet when the hypothesis space for $a$ is already discrete, the wasteful
-alternative somehow becomes standard practice.
+Yet when the hypothesis space for $a$ is discrete to begin with, the wasteful
+alternative has somehow become standard practice.
 

### 2014-10-26 Paper/ms.tex
Minor edits

@@
 stars: planetary systems --- techniques: radial velocities ---
-methods: data analysis --- methods:statistical
+methods: data analysis --- methods: statistical
 \end{keywords}
@@
 The number of known extrasolar planets has exploded in the last two
-decades \citep{}. These discoveries have been driven by improvements in
+decades, driven by improvements in
 all of the different techniques used to detect and characterise exoplanets,
@@
 exoplanetary system from observational data can be challenging.
-With radial velocity data,
+In the particular case of radial velocity data,
 the expected signal due to an exoplanet is periodic, and the goal is to
@@
 \bibitem[\protect\citeauthoryear{Brewer}{2014}]{rjobject} Brewer, B. J., 2014,
-submitted.
+Bayesian Analysis, submitted.
 

### 2014-10-26 Paper/ms.tex
Paragraph

@@
 all of the different techniques used to detect and characterise exoplanets,
-including the radial velocity method \citep[e.g.][]{},
+including the radial velocity (RV) method \citep[e.g.][]{},
 the transit method \citep[e.g.]{},
@@
 \citep[e.g.][]{2011MNRAS.410...94G, 2014MNRAS.437.3540F,
-2011A&A...528L...5T, fengji}.
+2011A&A...528L...5T, fengji}. Bayesian model fitting via Markov Chain Monte
+Carlo (MCMC) tends to be very computationally intensive, especially if we
+want to calculate the posterior distribution for $N$, the number of planets.
+
+In this letter we introduce a trans-dimensional MCMC approach to inferring
+$N$. When $N$ is treated as just another model parameter, we can obtain its
+posterior distribution in a single run. Computing the expected RV signal from
+a planet in an elliptical orbit can be slow. Therefore we use a lookup table
+of precomputed orbits, allowing us to do $\sim 15000$ MCMC steps per second
+per CPU core. The exploration is done using Diffusive Nested Sampling
+\citep{dnest} which replaces the posterior distribution with an alternative
+{\it mixture of constrained priors}, allowing mixing between separated modes.
 

### 2014-10-26 Paper/ms.tex
Birth death reference

@@
 
-In this letter we introduce a trans-dimensional MCMC approach to inferring
-$N$. When $N$ is treated as just another model parameter, we can obtain its
+In this letter we introduce a trans-dimensional birth-death MCMC approach
+\citep{birthdeath} to inferring $N$.
+When $N$ is treated as just another model parameter, we can obtain its
 posterior distribution in a single run. Computing the expected RV signal from
@@
 
+\bibitem[Stephens(2000)]{birthdeath} Stephens, M., 2000, ``Bayesian analysis of mixture models with an unknown number of components-an alternative to reversible jump methods.'', Annals of Statistics, 40-74.
+
 \bibitem[\protect\citeauthoryear{Tuomi}{2011}]{2011A&A...528L...5T} Tuomi M., 2011, A\&A, 528, L5

### 2014-10-27 Paper/ms.tex
Paragraphs on trans dimensional stuff

@@
 
+\subsection{Trans-dimensional MCMC}
+Trans-dimensional MCMC methods such as birth-death MCMC \citep{birthdeath} or the
+more general reversible jump MCMC \citep{green} treat the model dimension
+$N$ as just another model parameter. At fixed $N$, standard techniques such
+as the Metropolis algorithm can be used to explore the posterior distribution.
+Additional moves that propose to change the value of $N$ are also defined. The
+simplest of these are birth-death moves. More complicated moves, such as
+split-and-merge \citep[e.g.][]{umstatter} are possible but usually not necessary.
+
+In the exoplanet context, a birth move
+proposes to add one more planet to the model. The new planet's properties
+(period, amplitude, eccentricity, etc) are drawn from their prior distribution
+which may depend on other other model parameters.
+The corresponding death move simply chooses a planet currently in the model,
+and removes it.
+
 The unknown parameters are:
@@
 
+\bibitem[\protect\citeauthoryear{Green}{1995}]{green}
+Green, P.~J., 1995, Reversible Jump Markov Chain Monte Carlo Computation and Bayesian Model Determination, Biometrika 82 (4): 711–732.
+
 \bibitem[\protect\citeauthoryear{Gregory}{2011}]{2011MNRAS.415.2523G} 
@@
 
+\bibitem[\protect\citeauthoryear{Umst{\"a}tter et 
+al.}{2005}]{umstatter} Umst{\"a}tter R., Christensen N., Hendry 
+M., Meyer R., Simha V., Veitch J., Vigeland S., Woan G., 2005, PhRvD, 72, 
+022001
+
 \bibitem[\protect\citeauthoryear{Yee et al.}{2014}]{2014ApJ...790...14Y} 

### 2014-11-10 Paper/ms.tex
Speed

@@
 We apply the method to $\nu$ Oph and Gliese 581, finding moderate evidence
-(posterior probability of existence = 70\%) for a third planet around
+(posterior probability of existence = 70\%) for a third signal in
 $\nu$ Oph with an orbital period of 75.7 $\pm$ 1.8 days. The results also
 suggest Gliese 581 hosts many (7-15) planets (or other causes of other periodic
-signals), but only four have well determined orbital periods. 
+signals), but only four have well determined periods. 
 \end{abstract}
@@
 {\it mixture of constrained priors}, allowing mixing between separated modes.
+As a result, we are able to sample the posterior distribution for $N$ in
+a single run which takes about 15 minutes, whereas the approach of
+\citep{gregory} takes approximately 30 minutes per planet
+(Gregory, priv. comm.).
 

### 2014-11-10 Paper/ms.tex
rjobject para

@@
 The corresponding death move simply chooses a planet currently in the model,
-and removes it.
+and removes it. The acceptance probability for these moves is 1 if we want
+to explore the prior. To implement these moves in Nested Sampling
+\citep{skilling}, where the target distribution is proportional to the prior but with a hard likelihood
+constraint, then the acceptance probability is 1 if the proposed move
+satisfies the likelihood constraint, and 0 if it does not.
 
@@
 
-Diffusive Nested Sampling \citep{dnest} is a variant of
-Nested Sampling \citep{skilling}. RJObject \citep{rjobject}
+A recent paper \citep{rjobject} introduced a general approach to implementing
+trans-dimensional models within Diffusive Nested Sampling \citep{dnest}, a
+general MCMC algorithm. The \citet{rjobject} software predefines the
+Metropolis proposals for exploring trans-dimensional target distributions,
+including when the prior for the properties of each model component (i.e. each
+planet) is defined hierarchically.
 
@@
 \end{table*}
-
-We used hierarchical priors because...
 

### 2014-11-10 Paper/ms.tex
parameters

@@
 
-The unknown parameters are:
-\begin{eqnarray}
-\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\phi}\}_{i=1}^N, y_0, s, \nu\right\}
-\end{eqnarray}
-where $\boldsymbol{\alpha} = \{\mu_T, \sigma_T, \mu_A\}$ are the
-hyperparameters, and $\phi_i = \{\}$ are the properties of planet $i$.
-
 A recent paper \citep{rjobject} introduced a general approach to implementing
@@
 planet) is defined hierarchically.
+
+To fit a planet model to radial velocity data, we need parameters to describe
+the properties of each planet. We also include
+
+The unknown parameters are:
+\begin{eqnarray}
+\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\psi}\}_{i=1}^N, y_0, s, \nu\right\}
+\end{eqnarray}
+where $N$ is the number of planets,
+$\boldsymbol{\alpha} = \{\mu_T, \sigma_T, \mu_A\}$ are the
+hyperparameters used to define the prior for the properties of the planets,
+and $\psi_i = \{T_i, A_i, \phi_i, v_i, \theta_i\}$
+are the properties of planet $i$.
+The parameter $y_0$ describes a DC offset in the data, and $s$ and $\nu$ are
+parameters allowing for a non-gaussian distribution for the noise in the
+measurements.
 

### 2014-11-10 Paper/ms.tex
Included H
Figure sizes

@@
 
-The marginal likelihood was $\ln(\mathcal{Z}) \approx -217.04$.
+The marginal likelihood was $\ln(\mathcal{Z}) \approx -217.04$ and the
+information was $\mathcal{H} \approx 80.0$ nats.
 
 \begin{figure}
-\includegraphics[scale=0.5]{Figures/nu_oph_N.eps}
+\includegraphics[scale=0.45]{Figures/nu_oph_N.eps}
 \caption{The posterior distribution for the number of planets $N$ orbiting
@@
 \begin{figure}
-\includegraphics[scale=0.5]{Figures/nu_oph_periods.eps}
+\includegraphics[scale=0.45]{Figures/nu_oph_periods.eps}
 \caption{The posterior distribution for the orbital periods in the $\nu$ Oph
@@
 \begin{figure}
-\includegraphics[scale=0.5]{Figures/gliese581_N.eps}
+\includegraphics[scale=0.45]{Figures/gliese581_N.eps}
 \caption{The posterior distribution for the number of planets $N$ orbiting
@@
 \begin{figure}
-\includegraphics[scale=0.5]{Figures/gliese581_periods.eps}
+\includegraphics[scale=0.45]{Figures/gliese581_periods.eps}
 \caption{The posterior distribution for the orbital periods in the Gliese 581
@@
 
-The marginal likelihood was $\ln(\mathcal{Z}) \approx -618.04$.
+The marginal likelihood was $\ln(\mathcal{Z}) \approx -618.04$ and the
+information was $\mathcal{H} \approx 140.2$ nats.
 

### 2014-11-11 Paper/ms.tex
Ref
Transit ref

@@
 including the radial velocity (RV) method \citep[e.g.][]{},
-the transit method \citep[e.g.]{},
+the transit method \citep[e.g.]{2014PNAS..11112647B},
 and gravitational microlensing
@@
 a single run which takes about 15 minutes, whereas the approach of
-\citep{gregory} takes approximately 30 minutes per planet
+\citep{2011MNRAS.410...94G} takes approximately 30 minutes per planet
 (Gregory, priv. comm.).
@@
 \begin{thebibliography}{99}
+\bibitem[\protect\citeauthoryear{Batalha}{2014}]{2014PNAS..11112647B} 
+Batalha N.~M., 2014, PNAS, 111, 12647
 

### 2014-11-11 Paper/ms.tex
Paras about parameters

@@
 {\it mixture of constrained priors}, allowing mixing between separated modes.
-As a result, we are able to sample the posterior distribution for $N$ in
-a single run which takes about 15 minutes, whereas the approach of
+As a result, we are able to sample the posterior distribution for $N$, and
+evaluate the marginal likelihood (including the sum over $N$) in
+a single run which takes about 10 minutes on a 2-3 planet system.
+On the other hand the approach of
 \citep{2011MNRAS.410...94G} takes approximately 30 minutes per planet
@@
 
-The traditional approach also contradicts the normal approach to Bayesian
+The traditional approach to inferring $N$ also contradicts
+fundamental ideas in Bayesian
 computation. Imagine we are trying to compute the posterior distribution for
@@
 (period, amplitude, eccentricity, etc) are drawn from their prior distribution
-which may depend on other other model parameters.
+which may depend on other other model parameters or hyperparameters.
 The corresponding death move simply chooses a planet currently in the model,
@@
 
-To fit a planet model to radial velocity data, we need parameters to describe
-the properties of each planet. We also include
+To fit a planet model to RV data, we need parameters to describe
+the properties of each planet. For simplicity, we describe each planet by
+five parameters: $T_i$, the orbital period, $A_i$, the amplitude (in metres
+per second) of the RV signal, $\phi_i$, the phase of the signal, $v_i$, an
+``initial velocity'' parameter describing the eccentricity of the orbit
+(see Section~\ref{sec:orbits}), and $\theta_i$, a ``viewing angle'' parameter
+describing the position of the observer with respect to the major axis of the
+star's elliptical trajectory.
+
+We assigned hierarchical priors to the planet's parameters. This allows the
+model to capture the idea that knowing the values of some planet's parameters
+provides some information about the parameters of another planet. Not using
+hierarchical priors usually implies a strong prior commitment to the hypothesis
+that the properties of the planets are spread out across the whole domain of
+possible values, which is not necessarily the case.
 
@@
 parameters allowing for a non-gaussian distribution for the noise in the
-measurements.
+measurements. When the model-predicted RV signal is $m(t)$, the sampling
+distribution for each measurement was a student-$t$ distribution with
+scale parameter $K\sigma_i$ and shape parameter $\nu$. In the limit
+$(s_0 \to 0, \nu \to \inf)$ the sampling distribution reduces to the standard
+assumption of a normal distribution with standard deviation $\sigma_i$ known
+from the error bars in the data set.
+
+All of the model assumptions are specified in detail in
+Table~\ref{tab:priors}.
 
@@
 {\bf Hyperparameters}	&	\\
+$N$		& Number of planets	& Uniform$(\{0, 1, ..., N_{\rm max}\})$\\
 $\mu_T$		&	Median orbital period	& $\ln(\mu_T) \sim$ Cauchy$(5.9, 1)T(-15.3, 27.1)$\\
@@
 {\bf Other}\\
-$N$		& Number of planets	& Uniform$(\{0, 1, ..., N_{\rm max}\})$\\
 $\sigma_{\rm extra}$	& ``Extra noise'' parameter	& $\ln(\sigma) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
@@
 \end{tabular}
+\caption{All of the prior distributions in our Bayesian model.
+The priors for the planet parameters are defined conditional on the values
+of the hyperparameters. Uniform priors were used for parameters like phases.
+For parameters where a rough initial guess is possible, heavy-tailed Cauchy
+distributions were used so this information could be taken into account
+in a non-dogmatic way.\label{tab:priors}}
 \end{table*}
 
-\section{Orbit Simulations}
+\section{Orbit Simulations}\label{sec:orbits}
 The expected (noise-free) signal due to an exoplanet is periodic, but

### 2014-11-11 Paper/ms.tex
A few edits

@@
 On the other hand the approach of
-\citep{2011MNRAS.410...94G} takes approximately 30 minutes per planet
-(Gregory, priv. comm.).
+\citet{2011MNRAS.410...94G} takes approximately 30 minutes per planet
+(Gregory, priv. comm.). A C++ implementation of our method is available online
+at {\tt https://github.com/eggplantbren/Exoplanet} under the terms of the
+GNU General Public Licence version 3.
 
@@
 distribution for each measurement was a student-$t$ distribution with
-scale parameter $K\sigma_i$ and shape parameter $\nu$. In the limit
+scale parameter $\sqrt{\sigma_i^2 + s_0^2}$ and shape parameter $\nu$. In the limit
 $(s_0 \to 0, \nu \to \inf)$ the sampling distribution reduces to the standard
@@
 To save time, we pre-computed the properties of orbits as a function of
-ellipticity.
+ellipticity. We also made the standard assumption that the planets do not
+interact, so the expected signal due to several planets is the sum of the
+contributions of each planet.
 
@@
 requires a ``viewing angle'' parameter $\theta$.
-
 Although we parameterised the orbits using the initial velocity $v$,
@@
 \citep[e.g.][]{2011AIPC.1331..102Q, 2012PASJ...64..135S, fengji}, with periods
-of $530.3$ and $3190$ days.
-
+of $530.3$ and $3190$ days. To test our approach we applied it to the RV
+data from \citep{2012PASJ...64..135S}.
 The posterior distribution for $N$, the number of planets, is shown in
-Figure~\ref{fig:nu_oph_N}.
+Figure~\ref{fig:nu_oph_N}, showing the most probable number of planets is 3.
 

### 2014-11-11 Paper/ms.tex
more typing

@@
 The posterior distribution for $N$, the number of planets, is shown in
-Figure~\ref{fig:nu_oph_N}, showing the most probable number of planets is 3.
+Figure~\ref{fig:nu_oph_N}, showing the most probable number of planets is 3,
+and the posterior probability that $N \geq 3$ is about 95\%.
 
@@
 \begin{eqnarray}
-f(S) &=& \sum_{N=1}^{10} \frac{1}{N}\sum_{i=1}^N p(S_i | N, \bdata).
+f(S) &=& \sum_{N=0}^{10} \frac{1}{N}\sum_{i=1}^N p(S_i | N, \bdata).
 \end{eqnarray}
+The two rightmost peaks are the commonly accepted periods, and the third
+prominent peak corresponds to a $75.7 $\pm$ 1.8$ day period. The probability
+this signal exists is 70\%, and given that it exists, its velocity amplitude
+is about 15 m s$^{-1}$.
 

### 2014-11-11 Paper/ms.tex
Fixed bug
Various edits

@@
 shows three prominent peaks. Defining the log-periods by
-$S_i =  \log_{10}(T/\textnormal{1 day})$,
+$S_i =  \log_{10}\left[T/(\textnormal{1 day})\right]$,
 Figure~\ref{fig:nu_oph_periods} is a Monte Carlo representation
@@
 \begin{eqnarray}
-f(S) &=& \sum_{N=0}^{10} \frac{1}{N}\sum_{i=1}^N p(S_i | N, \bdata).
+f(S) &=& \sum_{N=0}^{10} p(N | \bdata)\sum_{i=1}^N p(S_i | N, \bdata).
 \end{eqnarray}
 The two rightmost peaks are the commonly accepted periods, and the third
-prominent peak corresponds to a $75.7 $\pm$ 1.8$ day period. The probability
-this signal exists is 70\%, and given that it exists, its velocity amplitude
-is about 15 m s$^{-1}$.
+prominent peak corresponds to a 75.7 $\pm$ 1.8 day period. The probability
+this signal exists (technically, the probability a period exists between
+67.6 and 85.1 days)
+is 70\%, and given that it exists, its velocity amplitude
+is 18.6 $\pm$ 8.3 m s$^{-1}$ and its eccentricity (or velocity parameter $v$)
+is not well determined.
 
@@
 \section{Reanalysis of Gliese 581}
-
 The red dwarf star Gliese 581 is thought to host several planets. Exactly
@@
 whose existence is generally doubted.
+To run our code on the combined dataset from the HARPS and HIRES spectrographs,
+we extended the model to include separate DC offsets for each instrument, as
+well as separate ``extra noise'' parameters $s_0$ and $\nu$.
 
+The posterior distribution for $N$ is shown in Figure~\ref{fig:gliese581_N},
+and shows strong evidence for at least eight periods
+($P(N \leq 7 | \bdata) = 0.004$).
+It is now recognised that
+there are many possible sources for oscillations in a data set and not all
+such oscillations should be claimed as planets \citep{2014Sci...345..440R}. 
 Our model cannot distinguish between oscillations due to planets and
 oscillations due to stellar activity: any oscillations found in the dataset
-will be described as ``planets'' by the model.
+will be described as ``planets'' by the model. However it is interesting that
+we find many more signals in the data than previous authors.
+
+By inspecting the posterior distribution for the periods
+(Figure~\ref{fig:gliese581_periods}), we see that only four of the periods
+are well determined, corresponding to the known periods of Gliese 581 b, c, d,
+and e. The other ``periods'' are very uncertain. It is possible that a
+non-sinusoidal signal due to stellar activity is being modelled as several
+periods \citep{astero}, and if the model were extended to include a ``stochastic''
+oscillation \citep{gaussproc}, the number of periods detected may be reduced
+substantially.
 
@@
 al.}{2014}]{2014ApJ...785..155B} Bennett D.~P., et al., 2014, ApJ, 785, 155
+
+\bibitem[\protect\citeauthoryear{Brewer 
+\& Stello}{2009}]{gaussproc} Brewer B.~J., Stello D., 2009, MNRAS, 395, 2226
+
+\bibitem[\protect\citeauthoryear{Brewer et al.}{2007}]{astero} 
+Brewer B.~J., Bedding T.~R., Kjeldsen H., Stello D., 2007, ApJ, 654, 551
 

### 2014-11-11 Paper/ms.tex
Another Feroz reference

@@
 for each possible value of $N$
-\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}, marginalising
+\citep[e.g.][]{2011MNRAS.415.2523G, 2011MNRAS.415.3462F, 2014MNRAS.437.3540F, fengji}, marginalising
 over all other model parameters.
@@
 
+\bibitem[\protect\citeauthoryear{Feroz, Balan, 
+\& Hobson}{2011}]{2011MNRAS.415.3462F} Feroz F., Balan S.~T., Hobson M.~P., 2011, MNRAS, 415, 3462
+
 \bibitem[\protect\citeauthoryear{Green}{1995}]{green}

### 2014-11-13 Paper/ms.tex
Priors

@@
 
-We assigned hierarchical priors to the planet's parameters. This allows the
+We assigned hierarchical priors to some of the planet's parameters. This allows the
 model to capture the idea that knowing the values of some planet's parameters
@@
 All of the model assumptions are specified in detail in
-Table~\ref{tab:priors}.
+Table~\ref{tab:priors}. Cauchy distributions were used when there is a
+preferred value, but since these have very heavy tails, the assumption is
+quite fail safe. For example, the prior for $\mu_T$, the typical orbital period,
+is centered around 1 year.
+
+An apparently strange choice is the conditional prior for the orbital
+periods, which is a biexponential distribution. A more conventional choice
+would have been a normal distribution, however the biexponential has
+analytical cumulative distribution and inverse, which is required for
+implementing it in the \citet{rjobject} framework.
+
+The prior for the velocity amplitudes $A_i$ is important and will affect the
+inferred number of planets $N$. For example, if the prior allows for very
+low amplitude planets to exist, the data will not rule this out. The prior
+will influence how many of these low amplitude planets will exist: if we
+believe there are many, and the data are uninformative about low amplitude
+planets, then the posterior distribution for $N$ will also indicate that there
+may be many low amplitude planets. However, their other properties, such as
+their orbital periods, will not be well determined.
 
@@
 distributions were used so this information could be taken into account
-in a non-dogmatic way.\label{tab:priors}}
+in a non-dogmatic way. Time units are in days and amplitude units are
+in metres per second.\label{tab:priors}}
 \end{table*}

### 2014-11-13 Paper/ms.tex
Some more work

@@
 non-sinusoidal when the orbit is not perfectly circular. The expected
-shape $m(t)$ of the variations can be obtained from mechanical considerations.
+shape $m(t)$ of the variations is needed in order to evaluate the likelihood
+function for any proposed setting of the parameters.
 To save time, we pre-computed the properties of orbits as a function of
@@
 
+By precomputing a set of orbits before running the MCMC, we are able to
+do of order 10,000 likelihood evaluations per second per CPU core.
+
 \section{Reanalysis of $\nu$ Oph}
@@
 
-The marginal likelihood was $\ln(\mathcal{Z}) \approx -217.04$ and the
-information was $\mathcal{H} \approx 80.0$ nats.
+The marginal likelihood for our model was
+$\ln(\mathcal{Z}) \approx -217.04$ and the information
+(Kullback-Leibler divergence from the prior to the posterior)
+was $\mathcal{H} \approx 80.0$ nats.
 

### 2014-11-13 Paper/ms.tex
Conclusions etc

@@
 \citep{skilling} and ideas related to thermodynamic integration
-\citep[e.g.][]{neal}.
+\citep[e.g.][]{neal}. Note that thermodynamic integration can easily fail if
+there is a phase transition in the problem \citep{skilling}.
 
@@
 in a non-dogmatic way. Time units are in days and amplitude units are
-in metres per second.\label{tab:priors}}
+in metres per second. The maximum number of planets, $N_{\rm max}$, was
+set to 15.\label{tab:priors}}
 \end{table*}
@@
 The marginal likelihood was $\ln(\mathcal{Z}) \approx -618.04$ and the
-information was $\mathcal{H} \approx 140.2$ nats.
+information was $\mathcal{H} \approx 140.2$ nats. This compares favourably
+to the marginal likelihood of $-640.1$ found by \citet{fengji}, although it
+is unclear whether we used exactly the same dataset.
+
+\section{Conclusions}
+In this letter we introduced a trans-dimensional MCMC approach to inferring
+the number of planets $N$ in an exoplanetary system from radial velocity data.
+The MCMC was implemented using the framework of \citet{rjobject} which
+defines trans-dimensional birth and death moves, and does the sampling
+with respect to a Nested Sampling target distribution, rather than directly
+sampling the posterior. This approach allows us to compute the results in a
+single run, which provides posterior samples and an estimate of the marginal
+likelihood.
+
+We applied the code to two well-studied RV datasets, $\nu$ Oph and Gliese 581.
+In $\nu$ Oph, we found some evidence for a signal with a period of
+75.7 $\pm$ 1.8 days: the posterior probability it exists, given our assumptions,
+is 70\%. With the combined HIRES+HARPS dataset from Gliese 581, we found
+evidence for a large number of ``planets'', although only four have well
+determined periods, corresponding to the Gliese 581 b, c, d,
+and e. Since our model does not include any possibility of stellar variability,
+any such periodic signals will be attributed to planets. Including non-planetary
+stellar variability is a crucial next step.
 
@@
 finally work on this problem, and Phil Gregory (UBC) for writing so many
-interesting papers on it. This work was supported by a Marsden Fast Start
+interesting papers on it. Tom Loredo (Cornell) also provided interesting
+conversations. This work was supported by a Marsden Fast Start
 grant from the Royal Society of

### 2014-11-13 Paper/ms.tex
Fengji's Z

@@
 information was $\mathcal{H} \approx 140.2$ nats. This compares favourably
-to the marginal likelihood of $-640.1$ found by \citet{fengji}, although it
-is unclear whether we used exactly the same dataset.
+to the marginal likelihood of $-640.1$ (for a 6-planet model)
+found by \citet{fengji}, although it is unclear whether we used exactly the
+same dataset.
 

### 2014-11-13 Paper/ms.tex
Talking about phase transitions

@@
 
-\date{To be submitted to MNRAS Letters}
+\date{To be submitted to MNRAS}
 
@@
 same dataset.
+Interestingly, the log-likelihood curve (Figure~\ref{fig:logl})
+shows this problem has two phase transitions.
+
+\begin{figure}
+\includegraphics[scale=0.45]{Figures/logl.eps}
+\caption{Log likelihood vs. enclosed prior mass for the Gliese 581 analysis.
+The two concave-up regions (at $\log(X) \approx -25$ and -45) correspond to
+phase transitions. Thermal approaches to this problem would produce misleading
+estimtes of the marginal likelihood because they would mix poorly at temperatures
+around 11 and 4.
+\label{fig:logl}}
+\end{figure}
 
 \section{Conclusions}
-In this letter we introduced a trans-dimensional MCMC approach to inferring
+In this paper we introduced a trans-dimensional MCMC approach to inferring
 the number of planets $N$ in an exoplanetary system from radial velocity data.

### 2014-11-13 Paper/ms.tex
Waffling about phase transitions

@@
 
-The marginal likelihood for our model was
-$\ln(\mathcal{Z}) \approx -217.04$ and the information
-(Kullback-Leibler divergence from the prior to the posterior)
-was $\mathcal{H} \approx 80.0$ nats.
-
 \begin{figure}
@@
 \end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.45]{Figures/logl0.eps}
+\caption{Log likelihood vs. enclosed prior mass (a standard output of Nested
+Sampling algorithms) for the $\nu$ Oph analysis.
+There are several phase transitions (concave-up regions) present. The one at
+$\log(X) \approx -75$ separates models which contain the third signal from
+models that do not. Without using Nested Sampling it would be very difficult
+to mix between these two situations and calculate the posterior probability
+for the existence of the third signal.
+\label{fig:logl0}}
+\end{figure}
+
+Figure~\ref{fig:logl0} shows the relationship between the likelihood $L$
+and the enclosed prior mass $X$ for the $\nu$ Oph analysis. These plots are
+a standard output of Nested Sampling \citep{skilling} analyses, and provide
+insights into the structure of the problem. Concave-up regions of this curve
+indicate phase transitions which can cause severe problems for annealing-based
+methods, and sometimes even for sampling the posterior distribution. In this
+analysis, the models with a third signal exist to the left of the phase
+transition at $\log(X) \approx -75$, and models without the signal exist to the
+right of the phase transition. Mixing between these two phases is crucial for
+accurately computing the posterior probability that the third signal exists.
+However, if we had simply tried to sample the posterior distribution, it would
+not mix between these two phases because the posterior is composed of a
+large-volume low-likelihood ``slab'' and a low-volume high-likelihood ``spike''.
+It is difficult for standard MCMC methods to escape the spike because of the
+likelihood contrast, and it is also difficult to escape the slab and enter the
+spike because of its small volume. Thus, even unimodal posterior distributions
+can cause mixing problems analogous to multimodal situations if a phase
+transition is present.
+
+The marginal likelihood for our model was
+$\ln(\mathcal{Z}) \approx -217.04$ and the information
+(Kullback-Leibler divergence from the prior to the posterior)
+was $\mathcal{H} \approx 80.0$ nats.
 
@@
 Interestingly, the log-likelihood curve (Figure~\ref{fig:logl})
-shows this problem has two phase transitions.
+shows this problem has two phase transitions. While these do not affect the
+posterior distribution (as they did for $\nu$ Oph), they would cause difficulties
+if we tried to calculte the marginal likelihood using thermal annealing.
 
@@
 single run, which provides posterior samples and an estimate of the marginal
-likelihood.
+likelihood. By embedding the MCMC in a Nested Sampling algorithm, instead of
+directly trying to sample the posterior distribution, we can overcome difficult
+features in the problem, such as phase transitions.
 
@@
 75.7 $\pm$ 1.8 days: the posterior probability it exists, given our assumptions,
-is 70\%. With the combined HIRES+HARPS dataset from Gliese 581, we found
+is 70\%. The posterior distribution contains models with three signals and
+models with two, however, these are separated by a phase transition. Therefore
+mixing between the two situations would be very difficult if we simply tried
+to sample the posterior distribution.
+
+With the combined HIRES+HARPS dataset from Gliese 581, we found
 evidence for a large number of ``planets'', although only four have well
@@
 and e. Since our model does not include any possibility of stellar variability,
-any such periodic signals will be attributed to planets. Including non-planetary
-stellar variability is a crucial next step.
+any such periodic signals will be attributed to ``planets''.
+Including non-planetary stellar variability is a crucial next step.
 

### 2014-11-13 Paper/ms.tex
Fixed typo

@@
 phase transitions. Thermal approaches to this problem would produce misleading
-estimtes of the marginal likelihood because they would mix poorly at temperatures
+estimates of the marginal likelihood because they would mix poorly at temperatures
 around 11 and 4.

### 2014-11-13 Paper/ms.tex
A little tidying up

@@
 
-\author[Brewer]{%
+\author[Brewer and Donovan]{%
   Brendon~J.~Brewer$^{1}$\thanks{bj.brewer@auckland.ac.nz},
@@
 all of the different techniques used to detect and characterise exoplanets,
-including the radial velocity (RV) method \citep[e.g.][]{},
+including the radial velocity (RV) method \citep[e.g.][]{2012PASJ...64..135S},
 the transit method \citep[e.g.]{2014PNAS..11112647B},
@@
 When $N$ is treated as just another model parameter, we can obtain its
-posterior distribution in a single run. Computing the expected RV signal from
-a planet in an elliptical orbit can be slow. Therefore we use a lookup table
-of precomputed orbits, allowing us to do $\sim 15000$ MCMC steps per second
-per CPU core. The exploration is done using Diffusive Nested Sampling
+posterior distribution in a single run.
+The exploration is done using Diffusive Nested Sampling
 \citep{dnest} which replaces the posterior distribution with an alternative
@@
 GNU General Public Licence version 3.
+
+Computing the expected RV signal from
+a planet in an elliptical orbit can be slow. Therefore we use a lookup table
+of precomputed orbits, allowing us to do $\sim 15000$ MCMC steps per second
+per CPU core.
 
@@
 of $530.3$ and $3190$ days. To test our approach we applied it to the RV
-data from \citep{2012PASJ...64..135S}.
+data from \citet{2012PASJ...64..135S}.
 The posterior distribution for $N$, the number of planets, is shown in

### 2014-11-14 Paper/ms.tex
Editing abstract

@@
 Inferring the number of planets $N$ in an exoplanetary system from radial velocity
-(RV) data is a challenging task. Typically, the posterior distribution for the
+(RV) data is a challenging task. Recently, it has become clear that RV data
+can contain periodic signals due to stellar activity, which can be difficult
+to distinguish from planetary signals. However, even doing the inference
+under a given set of siplifying assumptions (e.g. no stellar activity) can
+be difficult. It is common for the posterior distribution for the
 planet parameters, such as orbital periods,
-is multimodal and may have other awkward features. In
+to be multimodal and to have other awkward features. In
 addition, when $N$ is unknown, the marginal likelihood (or evidence), as a
@@
 suggest Gliese 581 hosts many (7-15) planets (or other causes of other periodic
-signals), but only four have well determined periods. 
+signals), but only four have well determined periods. The analysis of both
+of these datasets shows phase transitions exist which are difficult to
+negotiate without Nested Sampling.
 \end{abstract}

### 2014-11-14 Paper/ms.tex
Intro paragraph

@@
 $\nu$ Oph with an orbital period of 75.7 $\pm$ 1.8 days. The results also
-suggest Gliese 581 hosts many (7-15) planets (or other causes of other periodic
+suggest Gliese 581 hosts many (7-15) ``planets'' (or other causes of other periodic
 signals), but only four have well determined periods. The analysis of both
@@
 
-In this letter we introduce a trans-dimensional birth-death MCMC approach
+It is now well known that RV datasets can contain periodic signals
+resulting from stellar activity rather than planets \citep{2014Sci...345..440R}.
+Therefore, it is important to develop models which attempt to distinguish
+stellar activity signals from Keplerian planet signals based on the shape of
+the oscillations and/or additional data constraining the periods of any
+stellar activity signals. We do not address this important challenge in the
+present paper. Rather, we consider the problem of inferring the number $N$ of
+Keplerian signals in an RV dataset in a computationally efficient way.
+
+We introduce a trans-dimensional birth-death MCMC approach
 \citep{birthdeath} to inferring $N$.
@@
 The exploration is done using Diffusive Nested Sampling
-\citep{dnest} which replaces the posterior distribution with an alternative
+\citep[DNS][]{dnest} which replaces the posterior distribution with an alternative
 {\it mixture of constrained priors}, allowing mixing between separated modes.

### 2014-11-14 Paper/ms.tex
Spikes and slabs

@@
 \citep{skilling} and ideas related to thermodynamic integration
-\citep[e.g.][]{neal}. Note that thermodynamic integration can easily fail if
-there is a phase transition in the problem \citep{skilling}.
+\citep[e.g.][]{neal}.
 
@@
 
+\subsection{Phase transitions}
+It is well known that Bayesian computation (using MCMC for example) can be
+difficult when the posterior distribution is multimodal or has strong
+dependencies between parameters. An uncommon but less well-known difficulty
+is the existence of {\it phase transitions} \citep{skilling}.
+
+Imagine a high-dimensional unimodal posterior distribution that is composed of a
+broad, high volume but low density ``slab'' with a narrow, low volume but high
+density ``spike'' on top of it.
+An example is a mixture of two concentric high-dimensional gaussians with
+different widths. If you ran MCMC on such a posterior, it would be difficult
+to jump between the slab and the spike components. If the MCMC is currently
+in the spike region (or phase) it will be unable to escape: a proposed move
+into the slab will be rejected because of the ratio of densities. Conversely,
+if the MCMC was in the slab region, it would be unlikely to go into the spike
+region, because its volume is so small: it would be very unlikely to
+{\it propose} to move into the spike.
+
+\subsection{Parameters and Priors}
 To fit a planet model to RV data, we need parameters to describe

### 2014-11-14 Paper/ms.tex
Working...

@@
 % Bold symbols
-\renewcommand{\btheta}{\boldsymbol{\theta}}
-\newcommand{\bdata}{\boldsymbol{D}}
+%\renewcommand{	heta}{\boldsymbol{\theta}}
+%\newcommand{D}{\boldsymbol{D}}
 \newcommand{\bx}{\{x_i\}}
@@
 to the problem we are considering. Normally, this is the set of possible
-values of a vector of parameters $\btheta$ whose values we want
+values of a vector of parameters $	heta$ whose values we want
 to know. We then assign probability distributions
 called the {\it prior} and the {\it sampling distribution}. The prior
-distribution $p(\btheta)$ describes
+distribution $p(	heta)$ describes
 our initial uncertainty about which values of the parameters
-$\btheta$ are plausible, and the
-sampling distribution $p(\bdata | \btheta)$ describes
+$	heta$ are plausible, and the
+sampling distribution $p(D | 	heta)$ describes
 our initial uncertainty about the data set we're going to observe, as a
-function of the unknown parameters $\btheta$.
+function of the unknown parameters $	heta$.
 When the data is known, our state of knowledge about the parameter is
@@
 \begin{eqnarray}
-p(\btheta | \bdata) &=&
-\frac{p(\btheta)p(\bdata | \btheta)}
-{p(\bdata)}
+p(	heta | D) &=&
+\frac{p(	heta)p(D | 	heta)}
+{p(D)}
 \end{eqnarray}
-where $p(\bdata | \btheta)$ is called the likelihood, once the actual dataset
+where $p(D | 	heta)$ is called the likelihood, once the actual dataset
 is substituted in.
@@
 \begin{eqnarray}
-\mathcal{Z} = p(\bdata) = \int p(\btheta) p(\bdata | \btheta) d^n \btheta
+\mathcal{Z} = p(D) = \int p(	heta) p(D | 	heta) d^n 	heta
 \end{eqnarray}
@@
 \begin{eqnarray}
-p(\bdata | N) &=& \int p(\btheta | N) p(\bdata | \btheta, N) \, d^n \btheta
+p(D | N) &=& \int p(	heta | N) p(D | 	heta, N) \, d^n 	heta
 \end{eqnarray}
@@
 \begin{eqnarray}
-p(N | \bdata) &=& \frac{p(N)p(\bdata | N)}{\sum_N p(N)p(\bdata | N)}.
+p(N | D) &=& \frac{p(N)p(D | N)}{\sum_N p(N)p(D | N)}.
 \end{eqnarray}
@@
 region, because its volume is so small: it would be very unlikely to
-{\it propose} to move into the spike.
+{\it propose} to move into the spike. Thus, the situation behaves much like
+a multimodal posterior, despite only being unimodal.
+
+Even if the slab contains a very small amount of posterior mass, its existence
+can still cause problems with the calculation of the marginal likelihood
+if annealing methods are used. The thermodynamic integral formula gives
+the log of the marginal likelihood as an average of log likelihoods:
+\begin{eqnarray}
+\log(\mathcal{Z}) &=& \int_0^1 \left<\log\left[L(\theta)\right]\right>_\beta \, d\beta
+\end{eqnarray}
+where the expectation is taken with respect to the
+distribution with ``inverse temperature'' $\beta$, proportional to
+$\pi(\theta)L(\theta)^\beta$.
+
+Nested Sampling \citep{skilling}, and variants such as DNS, are not affected
+by phase transitions.
 
@@
 $N$		& Number of planets	& Uniform$(\{0, 1, ..., N_{\rm max}\})$\\
-$\mu_T$		&	Median orbital period	& $\ln(\mu_T) \sim$ Cauchy$(5.9, 1)T(-15.3, 27.1)$\\
+$\mu_T$		&	Median orbital period	& $\log(\mu_T) \sim$ Cauchy$(5.9, 1)T(-15.3, 27.1)$\\
 $w_T$		&	Diversity of orbital periods & $w_T \sim$ Uniform$(0.1, 3)$\\
-$\mu_A$		&	Mean amplitude	& $\ln(\mu_A) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
+$\mu_A$		&	Mean amplitude	& $\log(\mu_A) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
 \hline
 {\bf Planet Parameters}\\
-$T_i$		&	Orbital period	&	$\ln(T_i) \sim $ Biexponential($\ln(\mu_T), w_T$)\\
+$T_i$		&	Orbital period	&	$\log(T_i) \sim $ Biexponential($\log(\mu_T), w_T$)\\
 $A_i$		&	Amplitude of signal	& Exponential$(\mu_A)$\\
@@
 {\bf Other}\\
-$\sigma_{\rm extra}$	& ``Extra noise'' parameter	& $\ln(\sigma) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
-$\nu$		& Shape parameter for $t$-distribution for noise & $\ln(\nu) \sim$ Uniform$(\ln(0.01), \ln(1000))$\\
+$\sigma_{\rm extra}$	& ``Extra noise'' parameter	& $\log(\sigma) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
+$\nu$		& Shape parameter for $t$-distribution for noise & $\log(\nu) \sim$ Uniform$(\log(0.01), \log(1000))$\\
 \hline
@@
 \begin{eqnarray}
-f(S) &=& \sum_{N=0}^{10} p(N | \bdata)\sum_{i=1}^N p(S_i | N, \bdata).
+f(S) &=& \sum_{N=0}^{10} p(N | D)\sum_{i=1}^N p(S_i | N, D).
 \end{eqnarray}
@@
 However, if we had simply tried to sample the posterior distribution, it would
-not mix between these two phases because the posterior is composed of a
-large-volume low-likelihood ``slab'' and a low-volume high-likelihood ``spike''.
-It is difficult for standard MCMC methods to escape the spike because of the
-likelihood contrast, and it is also difficult to escape the slab and enter the
-spike because of its small volume. Thus, even unimodal posterior distributions
-can cause mixing problems analogous to multimodal situations if a phase
-transition is present.
+not mix easily between these two phases and would give a misleading posterior
+probability for the existence of the third signal.
 
 The marginal likelihood for our model was
-$\ln(\mathcal{Z}) \approx -217.04$ and the information
+$\log(\mathcal{Z}) \approx -217.04$ and the information
 (Kullback-Leibler divergence from the prior to the posterior)
@@
 and shows strong evidence for at least eight periods
-($P(N \leq 7 | \bdata) = 0.004$).
+($P(N \leq 7 | D) = 0.004$).
 It is now recognised that
@@
 
-The marginal likelihood was $\ln(\mathcal{Z}) \approx -618.04$ and the
+The marginal likelihood was $\log(\mathcal{Z}) \approx -618.04$ and the
 information was $\mathcal{H} \approx 140.2$ nats. This compares favourably

### 2014-11-14 Paper/ms.tex
phase transition talk

@@
 % Bold symbols
-%\renewcommand{	heta}{\boldsymbol{\theta}}
+%\renewcommand{\theta}{\boldsymbol{\theta}}
 %\newcommand{D}{\boldsymbol{D}}
@@
 to the problem we are considering. Normally, this is the set of possible
-values of a vector of parameters $	heta$ whose values we want
+values of a vector of parameters $\theta$ whose values we want
 to know. We then assign probability distributions
 called the {\it prior} and the {\it sampling distribution}. The prior
-distribution $p(	heta)$ describes
+distribution $p(\theta)$ describes
 our initial uncertainty about which values of the parameters
-$	heta$ are plausible, and the
-sampling distribution $p(D | 	heta)$ describes
+$\theta$ are plausible, and the
+sampling distribution $p(D | \theta)$ describes
 our initial uncertainty about the data set we're going to observe, as a
-function of the unknown parameters $	heta$.
+function of the unknown parameters $\theta$.
 When the data is known, our state of knowledge about the parameter is
@@
 \begin{eqnarray}
-p(	heta | D) &=&
-\frac{p(	heta)p(D | 	heta)}
+p(\theta | D) &=&
+\frac{p(\theta)p(D | \theta)}
 {p(D)}
 \end{eqnarray}
-where $p(D | 	heta)$ is called the likelihood, once the actual dataset
+where $p(D | \theta)$ is called the likelihood, once the actual dataset
 is substituted in.
@@
 \begin{eqnarray}
-\mathcal{Z} = p(D) = \int p(	heta) p(D | 	heta) d^n 	heta
+\mathcal{Z} = p(D) = \int p(\theta) p(D | \theta) d^n \theta
 \end{eqnarray}
 where the integral is over the entire $n$-dimensional parameter space.
+In the context of Bayesian computation, the prior is often denoted $\pi(\theta)$,
+the likelihood $L(\theta)$, and the marginal likelihood $\mathcal{Z}$.
 
@@
 \begin{eqnarray}
-p(D | N) &=& \int p(	heta | N) p(D | 	heta, N) \, d^n 	heta
+p(D | N) &=& \int p(\theta | N) p(D | \theta, N) \, d^n \theta
 \end{eqnarray}
@@
 distribution with ``inverse temperature'' $\beta$, proportional to
-$\pi(\theta)L(\theta)^\beta$.
+$\pi(\theta)L(\theta)^\beta$. Even if the slab contains virtually zero mass
+at $\beta=1$ (i.e. the posterior), at some inverse temperature between 0 and 1
+the slab and the spike ought to contribute equally to the distribution. At
+this critical temperature the MCMC will fail to mix and will give a misleadingly
+low estimate of the average log likelihood at that temperature.
 
-Nested Sampling \citep{skilling}, and variants such as DNS, are not affected
-by phase transitions.
+In data analysis problems, phase transitions can exist when the data contains
+a ``big'' effect which provides a lot of information about some parameters, and
+a ``small'' subtle effect.
+Nested Sampling, and variants such as DNS, are not affected
+by phase transitions. Part of their output, the relationship between the
+likelihood $L$ and the enclosed prior mass $X$, can be used to diagnose whether the
+problem contains a phase transition. In particular, if the graph of $\log(L)$
+vs. $\log(X)$ is convex at some point, then a phase transition exists.
 

### 2014-11-14 Paper/ms.tex
More editing

@@
 periods, which is a biexponential distribution. A more conventional choice
-would have been a normal distribution, however the biexponential has
-analytical cumulative distribution and inverse, which is required for
-implementing it in the \citet{rjobject} framework.
+would have been a normal distribution. However, the \citet{rjobject} software
+needs to know the corresponding cumulative distribution and its inverse, which
+are not available in closed form for the normal distribution.
 
@@
 \end{eqnarray}
-where $r = \sqrt{x^2 + y^2}$. These ODEs can be solved for any given set of
-initial conditions for $x$, $y$, $\frac{dx}{dt}$, and $\frac{dy}{dt}$ to yield
-the shape of the orbit. To simulate orbits
-of different ellipticities, we set the initial position to $(1, 0)$, and
+where $r = \sqrt{x^2 + y^2}$. The solution to these ODEs are elliptical orbits
+with the focus at the origin. We set the initial position to $(1, 0)$, and
 the initial velocity to $(0, v)$ where $v \in [0.4, 1]$.
@@
 elliptical. For trial values of $v$ ranging from 0.4 to 1 in steps of 0.005,
-we simulated an orbit using a leapfrog integrator, and saved the
+we calculated the orbit, and saved the
 velocities $\frac{dx}{dt}$ and $\frac{dy}{dt}$ as a function of time to disk.
@@
 By precomputing a set of orbits before running the MCMC, we are able to
-do of order 10,000 likelihood evaluations per second per CPU core.
+do $\sim$ 15,000 likelihood evaluations per second per CPU core.
 
@@
 Figure~\ref{fig:nu_oph_N}, showing the most probable number of planets is 3,
-and the posterior probability that $N \geq 3$ is about 95\%.
+and the posterior probability that $N \geq 3$ is about 95\%. Of course, this
+third signal is not necessarily a planet but a feature in the data which is
+better explained by a periodic signal than by noise.
 
@@
 and the enclosed prior mass $X$ for the $\nu$ Oph analysis. These plots are
-a standard output of Nested Sampling \citep{skilling} analyses, and provide
+a standard output of Nested Sampling analyses, and provide
 insights into the structure of the problem. Concave-up regions of this curve
@@
 posterior distribution (as they did for $\nu$ Oph), they would cause difficulties
-if we tried to calculte the marginal likelihood using thermal annealing.
+if we tried to calculte the marginal likelihood using annealing.
 
@@
 finally work on this problem, and Phil Gregory (UBC) for writing so many
-interesting papers on it. Tom Loredo (Cornell) also provided interesting
-conversations. This work was supported by a Marsden Fast Start
+interesting papers on it. We also thank Tom Loredo (Cornell) and Dan
+Foreman-Mackey (NYU) for interesting conversations and feedback.
+
+This work was supported by a Marsden Fast Start
 grant from the Royal Society of

### 2014-11-14 Paper/ms.tex
New figure included in paper
Preparing for next figure

@@
 is not well determined.
+An example model fit to the data is shown in Figure~\ref{fig:nuoph}.
 
@@
 system.\label{fig:nu_oph_periods}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.45]{Figures/nuoph.eps}
+\caption{The $\nu$ Oph radial velocity data and an example model fit which
+includes the third period.\label{fig:nuoph}}
 \end{figure}
@@
 oscillation \citep{gaussproc}, the number of periods detected may be reduced
-substantially.
+substantially. An example model fit for Gliese 581 is shown in
+Figure~\ref{fig:gliese581}.
 
@@
 system.\label{fig:gliese581_periods}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.45]{Figures/gliese581.eps}
+\caption{A portion of the Gliese 581 data and an example model fit.
+\label{fig:gliese581}}
 \end{figure}

### 2014-11-14 Paper/ms.tex
nmax
DC offset in table, minor edits

@@
 $T_i$		&	Orbital period	&	$\log(T_i) \sim $ Biexponential($\log(\mu_T), w_T$)\\
-$A_i$		&	Amplitude of signal	& Exponential$(\mu_A)$\\
+$A_i$		&	Amplitude of signal	& Exponential$(\textnormal{mean}=\mu_A)$\\
 $\phi_i$	&	Phase of signal	&	Uniform($0, 2\pi$)\\
@@
 {\bf Other}\\
+$m_0$		&	Constant DC offset	&	Uniform$(m_{0, \rm min}, m_{0, \rm max})$\\
 $\sigma_{\rm extra}$	& ``Extra noise'' parameter	& $\log(\sigma) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
@@
 in a non-dogmatic way. Time units are in days and amplitude units are
-in metres per second. The maximum number of planets, $N_{\rm max}$, was
-set to 15.\label{tab:priors}}
+in metres per second. The maximum number of planets, $N_{\rm max}$, was set
+to 10. The prior limits for $m_0$ were set to the minimum and maximum value
+in the dataset, which is not strictly a valid strategy.\label{tab:priors}}
 \end{table*}
@@
 \end{eqnarray}
-where $r = \sqrt{x^2 + y^2}$. The solution to these ODEs are elliptical orbits
+where $r = \sqrt{x^2 + y^2}$. The solutions to this system of equations are elliptical orbits
 with the focus at the origin. We set the initial position to $(1, 0)$, and
@@
 we extended the model to include separate DC offsets for each instrument, as
-well as separate ``extra noise'' parameters $s_0$ and $\nu$.
+well as separate ``extra noise'' parameters $s_0$ and $\nu$. We also increased
+$N_{\rm max}$ from 10 to 15 for this system.
 

### 2014-11-14 Paper/ms.tex
Captions

@@
 \caption{The posterior distribution for the orbital periods in the $\nu$ Oph
-system.\label{fig:nu_oph_periods}}
+system. The two dashed vertical lines are the commonly accepted periods
+of $530.3$ and $3190$ days.\label{fig:nu_oph_periods}}
 \end{figure}
@@
 \caption{The posterior distribution for the orbital periods in the Gliese 581
-system.\label{fig:gliese581_periods}}
+system. The solid lines are Gliese 581 b and c, the dashed lines are
+``planets'' d and e, and the dotted lines are f and g.
+\label{fig:gliese581_periods}}
 \end{figure}

### 2014-11-14 MyDistribution.cpp
Phil's eccentricity prior

@@ -58,12 +58,13 @@ double MyDistribution::log_pdf(const std::vector<double>& vec) const
 {
 	if(vec[1] < 0. ||
 			vec[2] < 0. || vec[2] > 2.*M_PI ||
-			vec[3] < 0.4 || vec[3] > 1. ||
+			vec[3] < 0. || vec[3] > 0.8189776 ||
 			vec[4] < 0. || vec[4] > 2.*M_PI)
 		return -1E300;
 
 	return  -log(2.*width) - abs(vec[0] - center)/width
-		-log(mu) - vec[1]/mu;
+		-log(mu) - vec[1]/mu
+		+ 2.1*log(1. - vec[3]/0.995);
 }
 
 void MyDistribution::from_uniform(std::vector<double>& vec) const
@@ -74,7 +75,7 @@ void MyDistribution::from_uniform(std::vector<double>& vec) const
 		vec[0] = center - width*log(2. - 2.*vec[0]);
 	vec[1] = -mu*log(1. - vec[1]);
 	vec[2] = 2.*M_PI*vec[2];
-	vec[3] = 0.4 + 0.6*vec[3];
+	vec[3] = 1. - pow(1. - 0.995*vec[3], 1./3.1);
 	vec[4] = 2.*M_PI*vec[4];
 }
 
@@ -86,7 +87,7 @@ void MyDistribution::to_uniform(std::vector<double>& vec) const
 		vec[0] = 1. - 0.5*exp((center - vec[0])/width);
 	vec[1] = 1. - exp(-vec[1]/mu);
 	vec[2] = vec[2]/(2.*M_PI);
-	vec[3] = (vec[3] - 0.4)/0.6;
+	vec[3] = 1. - pow(1. - vec[3]/0.995, 3.1);
 	vec[4] = vec[4]/(2.*M_PI);
 }
 

### 2014-11-14 MyModel.cpp

@@ -56,7 +56,7 @@ void MyModel::calculate_mu()
 		T = exp(components[j][0]);
 		A = components[j][1];
 		phi = components[j][2];
-		v0 = components[j][3];
+		v0 = sqrt(1. - components[j][3]);
 		viewing_angle = components[j][4];
 		arg = t;
 		for(size_t i=0; i<arg.size(); i++)
@@ -153,7 +153,8 @@ void MyModel::print(std::ostream& out) const
 		T = exp(components[j][0]);
 		A = components[j][1];
 		phi = components[j][2];
-		v0 = components[j][3];
+		v0 = sqrt(1. - components[j][3]);
+
 		viewing_angle = components[j][4];
 		arg = t;
 		for(size_t i=0; i<arg.size(); i++)

### 2014-11-14 display.py
Nicer figures

@@ -11,15 +11,17 @@ show()
 T = posterior_sample[:,1008:1018]
 T = T[T != 0].flatten()
 # Trim
-s = sort(T)
-left, middle, right = s[0.25*len(s)], s[0.5*len(s)], s[0.75*len(s)]
-iqr = right - left
-s = s[logical_and(s > middle - 5*iqr, s < middle + 5*iqr)]
+#s = sort(T)
+#left, middle, right = s[0.25*len(s)], s[0.5*len(s)], s[0.75*len(s)]
+#iqr = right - left
+#s = s[logical_and(s > middle - 5*iqr, s < middle + 5*iqr)]
 
-hist(s, 200, alpha=0.5)
-xlabel(r'$\ln$(Period/days)')
+hist(T/log(10.), 300, alpha=0.5)
+xlabel(r'$\log_{10}$(Period/days)')
+xlim([0, 4])
 show()
 
+data[:,0] -= data[:,0].min()
 t = linspace(data[:,0].min(), data[:,0].max(), 1000)
 
 saveFrames = False # For making movies
@@ -29,13 +31,14 @@ if saveFrames:
 ion()
 for i in xrange(0, posterior_sample.shape[0]):
   hold(False)
-  plot(data[:,0], data[:,1], 'b.')
+  errorbar(data[:,0], data[:,1], fmt='b.', yerr=data[:,2])
   hold(True)
   plot(t, posterior_sample[i, 0:1000], 'r')
+  xlim([-0.05*data[:,0].max(), 1.05*data[:,0].max()])
   ylim([-1.5*max(abs(data[:,1])), 1.5*max(abs(data[:,1]))])
-  axhline(0., color='k')
-  xlabel('Time', fontsize=16)
-  ylabel('y', fontsize=16)
+  #axhline(0., color='k')
+  xlabel('Time (days)', fontsize=16)
+  ylabel('Radial Velocity (m/s)', fontsize=16)
   draw()
   if saveFrames:
     savefig('Frames/' + '%0.4d'%(i+1) + '.png', bbox_inches='tight')

### 2014-11-14 display.py
Show amplitudes and eccentricities

@@ -1,7 +1,7 @@
 from pylab import *
 import os
 
-data = loadtxt('nu_oph.txt')
+data = loadtxt('HD208487.txt')
 posterior_sample = atleast_2d(loadtxt('posterior_sample.txt'))
 
 hist(posterior_sample[:,1007], 100)
@@ -9,7 +9,12 @@ xlabel('Number of Planets')
 show()
 
 T = posterior_sample[:,1008:1018]
-T = T[T != 0].flatten()
+A = posterior_sample[:,1018:1028]
+E = posterior_sample[:,1038:1048]
+which = T != 0
+T = T[which].flatten()
+A = A[which].flatten()
+E = E[which].flatten()
 # Trim
 #s = sort(T)
 #left, middle, right = s[0.25*len(s)], s[0.5*len(s)], s[0.75*len(s)]
@@ -19,6 +24,17 @@ T = T[T != 0].flatten()
 hist(T/log(10.), 300, alpha=0.5)
 xlabel(r'$\log_{10}$(Period/days)')
 xlim([0, 4])
+show()
+
+subplot(2,1,1)
+plot(T/log(10.), A, 'b.', markersize=1)
+xlim([0, 4])
+ylabel('Amplitude (m/s)')
+subplot(2,1,2)
+plot(T/log(10.), E, 'b.', markersize=1)
+xlim([0, 4])
+xlabel(r'$\log_{10}$(Period/days)')
+ylabel('Ellipticity')
 show()
 
 data[:,0] -= data[:,0].min()

### 2014-11-14 display.py
Load correct file

@@ -1,7 +1,7 @@
 from pylab import *
 import os
 
-data = loadtxt('HD208487.txt')
+data = loadtxt('nu_oph.txt')
 posterior_sample = atleast_2d(loadtxt('posterior_sample.txt'))
 
 hist(posterior_sample[:,1007], 100)

### 2014-11-19 Paper/ms.tex
Eccentricity prior

@@
 five parameters: $T_i$, the orbital period, $A_i$, the amplitude (in metres
-per second) of the RV signal, $\phi_i$, the phase of the signal, $v_i$, an
-``initial velocity'' parameter describing the eccentricity of the orbit
-(see Section~\ref{sec:orbits}), and $\theta_i$, a ``viewing angle'' parameter
+per second) of the RV signal, $\phi_i$, the phase of the signal, $e_i$, the eccentricity of the orbit, and $\theta_i$, a ``viewing angle'' parameter
 describing the position of the observer with respect to the major axis of the
@@
 hyperparameters used to define the prior for the properties of the planets,
-and $\psi_i = \{T_i, A_i, \phi_i, v_i, \theta_i\}$
+and $\psi_i = \{T_i, A_i, \phi_i, e_i, \theta_i\}$
 are the properties of planet $i$.
@@
 $\phi_i$	&	Phase of signal	&	Uniform($0, 2\pi$)\\
-$v_i$		&	Initial speed of orbit	&	Uniform$(0.4, 1)$\\
+$e_i$		&	Eccentricity of orbit	&	Beta$(1, 3.1)T(0, 0.8)$\\
 $\theta_i$	&	Viewing angle	&	Uniform$(0, 2\pi)$\\
@@
 requires a ``viewing angle'' parameter $\theta$.
-Although we parameterised the orbits using the initial velocity $v$,
-the standard eccentricity $\epsilon$ can be recovered using the relation
-$\epsilon = 1 - v^2$.
-
+The eccentricity of the orbit, in terms of $v$, is $e = 1 - v^2$.
 By precomputing a set of orbits before running the MCMC, we are able to

### 2014-11-19 display.py
Fake data like nu oph

@@ -1,7 +1,8 @@
 from pylab import *
 import os
 
-data = loadtxt('nu_oph.txt')
+data = loadtxt('fake_data_like_nuoph.txt')
+truth = loadtxt('fake_data_like_nuoph.truth')
 posterior_sample = atleast_2d(loadtxt('posterior_sample.txt'))
 
 hist(posterior_sample[:,1007], 100)
@@ -24,17 +25,22 @@ E = E[which].flatten()
 hist(T/log(10.), 300, alpha=0.5)
 xlabel(r'$\log_{10}$(Period/days)')
 xlim([0, 4])
+for i in xrange(1008, 1008 + int(truth[1007])):
+  axvline(truth[i]/log(10.), color='r')
 show()
 
 subplot(2,1,1)
 plot(T/log(10.), A, 'b.', markersize=1)
 xlim([0, 4])
 ylabel('Amplitude (m/s)')
+plot(truth[1008:1008 + int(truth[1007])]/log(10.), truth[1018:1018 + int(truth[1007])], 'ro')
+
 subplot(2,1,2)
 plot(T/log(10.), E, 'b.', markersize=1)
 xlim([0, 4])
 xlabel(r'$\log_{10}$(Period/days)')
 ylabel('Ellipticity')
+plot(truth[1008:1008 + int(truth[1007])]/log(10.), truth[1038:1038 + int(truth[1007])], 'ro')
 show()
 
 data[:,0] -= data[:,0].min()

### 2014-11-19 main.cpp

@@ -10,7 +10,7 @@ using namespace DNest3;
 int main(int argc, char** argv)
 {
 	// Load the data
-	Data::get_instance().load("nu_oph.txt");
+	Data::get_instance().load("fake_data_like_nuoph.txt");
 
 	Lookup::get_instance().load();
 

### 2014-11-19 display.py
hold true

@@ -31,12 +31,14 @@ show()
 
 subplot(2,1,1)
 plot(T/log(10.), A, 'b.', markersize=1)
+hold(True)
 xlim([0, 4])
 ylabel('Amplitude (m/s)')
 plot(truth[1008:1008 + int(truth[1007])]/log(10.), truth[1018:1018 + int(truth[1007])], 'ro')
 
 subplot(2,1,2)
 plot(T/log(10.), E, 'b.', markersize=1)
+hold(True)
 xlim([0, 4])
 xlabel(r'$\log_{10}$(Period/days)')
 ylabel('Ellipticity')

### 2014-11-19 Paper/ms.tex
Updated nu oph figures

@@
 The marginal likelihood for our model was
-$\log(\mathcal{Z}) \approx -217.04$ and the information
+$\log(\mathcal{Z}) \approx -217.65$ and the information
 (Kullback-Leibler divergence from the prior to the posterior)
-was $\mathcal{H} \approx 80.0$ nats.
+was $\mathcal{H} \approx 77.7$ nats.
 

### 2014-11-20 showresults.py
New Gliese581 figures

@@ -1,4 +1,4 @@
 import postprocess
-postprocess.postprocess()
+postprocess.postprocess(cut=0.5)
 import display
 

### 2014-11-20 Paper/ms.tex
Improved discussion of Robertson

@@
 (d and e, with periods 66 and 3.15 days respectively)
-whose existence is mostly accepted, and another two
+whose existence was mostly accepted, and another two
 (f and g, with periods 433 and 36.5 days respectively)
-whose existence is generally doubted.
+whose existence was generally doubted. However, \citet{2014Sci...345..440R}
+found that planets d and g do not exist but are signals due to stellar activity.
+While our model cannot account for stellar variability and contribute to that
+particular discussion, it is a challenging and interesting dataset from an
+inference point of view.
 To run our code on the combined dataset from the HARPS and HIRES spectrographs,

### 2014-11-20 Paper/ms.tex
Editing nu oph discussion

@@
 We apply the method to $\nu$ Oph and Gliese 581, finding moderate evidence
-(posterior probability of existence = 70\%) for a third signal in
-$\nu$ Oph with an orbital period of 75.7 $\pm$ 1.8 days. The results also
-suggest Gliese 581 hosts many (7-15) ``planets'' (or other causes of other periodic
+for additional signals in
+$\nu$ Oph with periods of 36.12 $\pm$ 0.04 days and
+75.57 $\pm$ 0.76 days; the posterior probability that at least one of these
+exists is 79\%. The results also suggest Gliese 581 hosts many (7-15) ``planets''
+(or other causes of other periodic
 signals), but only four have well determined periods. The analysis of both
@@
 Figure~\ref{fig:nu_oph_N}, showing the most probable number of planets is 3,
-and the posterior probability that $N \geq 3$ is about 95\%. Of course, this
-third signal is not necessarily a planet but a feature in the data which is
-better explained by a periodic signal than by noise.
+and the posterior probability that $N \geq 3$ is about 95\%. Of course, these
+extra possible signals are not necessarily a planet but a feature in the data
+which is better explained by a periodic signal than by noise (and may have been
+explained by correlated noise, had we included it).
 
@@
 \end{eqnarray}
-The two rightmost peaks are the commonly accepted periods, and the third
-prominent peak corresponds to a 75.7 $\pm$ 1.8 day period. The probability
-this signal exists (technically, the probability a period exists between
-67.6 and 85.1 days)
-is 70\%, and given that it exists, its velocity amplitude
-is 18.6 $\pm$ 8.3 m s$^{-1}$ and its eccentricity (or velocity parameter $v$)
-is not well determined.
+The two rightmost peaks are the commonly accepted periods, and the two next
+most prominent peaks have periods of 36.12 $\pm$ 0.04 days and
+75.57 $\pm$ 0.76 days.
+The probability
+at least one of these exists (the fraction of our samples which
+have one of these signals) is 79\%.
+Given that they exist, their amplitudes are low, around 10-15 metres per second,
+which we note is above the noise level. We also found that the 75.7 $\pm$ 1.8 day
+period prefers an eccentricity of about $0.37 \pm 0.14$.
 An example model fit to the data is shown in Figure~\ref{fig:nuoph}.
@@
 \caption{The posterior distribution for the number of planets $N$ orbiting
-$\nu$ Oph.\label{fig:nu_oph_N}}
+$\nu$ Oph. The posterior probability of $N > 2$ is about 96\%, however the
+prior probability of $N > 2$ was already high due to the uniform prior for $N$.
+\label{fig:nu_oph_N}}
 \end{figure}
@@
 system. The two dashed vertical lines are the commonly accepted periods
-of $530.3$ and $3190$ days.\label{fig:nu_oph_periods}}
+of $530.3$ and $3190$ days. The next most prominent peaks with well determined
+periods are at 36.12 $\pm$ 0.04 days and
+75.57 $\pm$ 0.76 days.\label{fig:nu_oph_periods}}
 \end{figure}
@@
 \caption{The $\nu$ Oph radial velocity data and an example model fit which
-includes the third period.\label{fig:nuoph}}
+includes a third period. The amplitude of this additional signal is low but
+is about twice the reported errorbars on the measurements.\label{fig:nuoph}}
 \end{figure}
@@
 There are several phase transitions (concave-up regions) present. The one at
-$\log(X) \approx -75$ separates models which contain the third signal from
+$\log(X) \approx -75$ separates models which contain additional signals from
 models that do not. Without using Nested Sampling it would be very difficult
 to mix between these two situations and calculate the posterior probability
-for the existence of the third signal.
+for the existence of the additional signals.
 \label{fig:logl0}}
@@
 right of the phase transition. Mixing between these two phases is crucial for
-accurately computing the posterior probability that the third signal exists.
+accurately computing the posterior probability that extra signals exist.
 However, if we had simply tried to sample the posterior distribution, it would
-not mix easily between these two phases and would give a misleading posterior
-probability for the existence of the third signal.
+not mix easily between these two phases and could easily fail to detect the
+extra signals.
 
@@
 In $\nu$ Oph, we found some evidence for a signal with a period of
-75.7 $\pm$ 1.8 days: the posterior probability it exists, given our assumptions,
-is 70\%. The posterior distribution contains models with three signals and
-models with two, however, these are separated by a phase transition. Therefore
-mixing between the two situations would be very difficult if we simply tried
-to sample the posterior distribution.
+either 36.12 $\pm$ 0.04 days or
+75.57 $\pm$ 0.76 days: the posterior probability one of these exists, given our
+assumptions, is 79\%. The posterior distribution contains models with several
+signals and models with two, however, these are separated by a phase transition.
+Therefore mixing between the two situations would be very difficult if we simply tried to sample the posterior distribution.
 

### 2014-11-22 Paper/ms.tex
Data in latex comment
Minor fixes

@@
 to distinguish from planetary signals. However, even doing the inference
-under a given set of siplifying assumptions (e.g. no stellar activity) can
+under a given set of simplifying assumptions (e.g. no stellar activity) can
 be difficult. It is common for the posterior distribution for the
@@
 The number of known extrasolar planets has exploded in the last two
-decades, driven by improvements in
+decades. This has been driven by improvements in
 all of the different techniques used to detect and characterise exoplanets,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% DATA USED
+
+% Nu Oph
+% Time (days), Radial velocity (m/s), Radial velocity errorbar (m/s)
+%2.337364819999999781e+03 -2.061999999999999886e+02 4.299999999999999822e+00
+%2.487101369999999861e+03 2.615000000000000000e+02 4.200000000000000178e+00
+%2.507077150000000074e+03 3.403000000000000114e+02 3.799999999999999822e+00
+%2.541955269999999928e+03 4.198999999999999773e+02 5.799999999999999822e+00
+%2.541958299999999781e+03 4.293999999999999773e+02 4.000000000000000000e+00
+%2.736306950000000143e+03 -2.519999999999999929e+01 5.299999999999999822e+00
+%2.857037139999999908e+03 -2.150000000000000000e+01 5.000000000000000000e+00
+%3.100276800000000094e+03 5.113000000000000114e+02 3.700000000000000178e+00
+%3.499178260000000137e+03 1.225000000000000000e+02 5.299999999999999822e+00
+%3.636025169999999889e+03 3.305000000000000000e+02 6.200000000000000178e+00
+%3.963951230000000123e+03 -2.350999999999999943e+02 4.900000000000000355e+00
+%4.309113010000000031e+03 -2.686999999999999886e+02 5.200000000000000178e+00
+%4.338994730000000345e+03 -3.363999999999999773e+02 3.399999999999999911e+00
+%4.560239679999999680e+03 -1.160000000000000000e+02 3.500000000000000000e+00
+%4.588269440000000031e+03 -2.230000000000000071e+01 5.000000000000000000e+00
+%4.674023040000000037e+03 1.691999999999999886e+02 4.099999999999999645e+00
+%4.674025380000000041e+03 1.686999999999999886e+02 4.099999999999999645e+00
+%4.677067339999999604e+03 1.716999999999999886e+02 3.799999999999999822e+00
+%4.928330149999999776e+03 -3.908000000000000114e+02 3.500000000000000000e+00
+%4.952301330000000235e+03 -3.595000000000000000e+02 4.500000000000000000e+00
+%4.953310220000000299e+03 -3.531999999999999886e+02 4.099999999999999645e+00
+%5.132930250000000342e+03 7.940000000000000568e+01 6.700000000000000178e+00
+%5.267333859999999731e+03 1.450000000000000000e+02 3.500000000000000000e+00
+%5.296329950000000281e+03 4.579999999999999716e+01 3.500000000000000000e+00
+%5.324284969999999703e+03 -3.520000000000000284e+01 4.000000000000000000e+00
+%5.347153669999999693e+03 -1.170000000000000000e+02 4.099999999999999645e+00
+%5.384187179999999898e+03 -1.946999999999999886e+02 4.200000000000000178e+00
+%5.398047959999999875e+03 -2.165999999999999943e+02 3.700000000000000178e+00
+%5.436026149999999689e+03 -2.645000000000000000e+02 3.500000000000000000e+00
+%5.467928210000000036e+03 -2.518000000000000114e+02 3.700000000000000178e+00
+%5.613381190000000061e+03 6.979999999999999716e+01 3.500000000000000000e+00
+%5.656341529999999693e+03 2.221999999999999886e+02 3.299999999999999822e+00
+%5.657340799999999945e+03 2.265000000000000000e+02 3.799999999999999822e+00
+%5.690210130000000390e+03 3.385000000000000000e+02 6.000000000000000000e+00
+%5.712159810000000107e+03 4.010000000000000000e+02 3.799999999999999822e+00
+%5.763070190000000366e+03 3.993000000000000114e+02 3.399999999999999911e+00
+%5.766075829999999769e+03 3.908999999999999773e+02 3.600000000000000089e+00
+%5.770066890000000058e+03 3.886000000000000227e+02 4.000000000000000000e+00
+%5.784950219999999717e+03 3.693000000000000114e+02 4.099999999999999645e+00
+%5.787005500000000211e+03 3.641999999999999886e+02 3.600000000000000089e+00
+%5.788002169999999751e+03 3.663000000000000114e+02 3.700000000000000178e+00
+%5.804928630000000339e+03 2.938000000000000114e+02 3.799999999999999822e+00
+%5.809972679999999855e+03 2.825000000000000000e+02 3.700000000000000178e+00
+%5.846919119999999566e+03 1.640999999999999943e+02 4.900000000000000355e+00
+
+% Gliese 581
+% First 122 points are HIRES, then HARPS follows
+%1409.76222   6.96 1.89 
+%1586.14605 -10.24 3.22 
+%1704.91213   0.47 2.89 
+%2003.95507  -4.37 3.65 
+%2100.86678 -19.45 2.22 
+%2161.73096   3.35 2.19 
+%2162.73165  10.73 2.38 
+%2335.15024   6.87 2.33 
+%2487.79326  -7.30 3.77 
+%2712.05433 -10.96 2.10 
+%2828.87708   7.75 2.25 
+%3077.15398  -7.18 3.18 
+%3239.82508  -2.50 2.60 
+%3426.14551  -2.74 4.82 
+%3748.16323  -4.97 1.96 
+%3754.14596  -8.68 1.97 
+%3932.80588   2.40 1.95 
+%3960.83771  12.14 1.90 
+%3961.77279   4.38 1.57 
+%3962.77418  -4.97 2.40 
+%3963.81985  -9.92 1.78 
+%3982.78069  14.04 1.92 
+%4130.14636 -14.79 1.84 
+%4131.16220  -4.58 2.58 
+%4139.16155   1.60 1.92 
+%4246.81907  -4.50 1.18 
+%4247.92767  -9.21 1.84 
+%4248.82815  -7.84 1.48 
+%4249.90281   6.45 1.38 
+%4250.81292  18.17 1.95 
+%4251.81942   7.61 1.30 
+%4255.79771   7.35 3.62 
+%4277.76996  17.11 1.68 
+%4278.78138  14.06 1.71 
+%4279.78247 -10.60 2.03 
+%4285.78969 -15.08 2.54 
+%4294.84381  -0.04 2.04 
+%4300.81847  -0.76 1.83 
+%4304.79872  16.63 2.24 
+%4305.80545   2.61 1.45 
+%4306.83201  -8.05 2.31 
+%4307.83943  -8.03 1.47 
+%4308.82016   1.74 1.27 
+%4309.80568  13.32 1.26 
+%4310.80271  11.19 1.26 
+%4311.79457  -4.40 1.23 
+%4312.79306  -7.14 1.13 
+%4313.80176   8.14 1.45 
+%4314.82230  13.67 1.35 
+%4318.79357  -5.91 1.44 
+%4335.75166   5.83 1.35 
+%4336.78425  14.68 1.84 
+%4339.72676  -8.58 2.69 
+%4343.74861  -0.29 2.06 
+%4345.74397  -5.82 1.94 
+%4461.17458   3.57 2.13 
+%4545.09518   4.98 1.95 
+%4546.07813  17.48 1.52 
+%4547.07715   9.06 2.74 
+%4547.99454  -3.97 2.45 
+%4600.98896  13.54 2.03 
+%4601.90368  -8.33 2.69 
+%4633.86049  -6.47 1.49 
+%4634.86134 -12.30 1.27 
+%4635.87975  -5.69 1.19 
+%4636.88690  12.26 2.07 
+%4637.88502  15.38 1.56 
+%4638.94288  -1.08 1.40 
+%4639.91382 -11.27 1.39 
+%4640.92943  -6.50 1.95 
+%4641.90267   5.64 1.33 
+%4643.88456   0.95 1.74 
+%4666.84246  -7.60 1.25 
+%4667.84194  -8.11 2.07 
+%4671.85435  -9.71 1.33 
+%4672.83453  -3.24 1.00 
+%4673.84973   8.11 1.03 
+%4686.82794   2.60 2.22 
+%4688.77854  -9.57 1.96 
+%4689.77880   3.13 1.29 
+%4701.79211  17.51 1.17 
+%4702.81250  11.35 1.72 
+%4703.80853  -1.73 1.54 
+%4704.80341  -7.99 1.20 
+%4717.72887  14.81 1.48 
+%4718.73784   5.10 2.03 
+%4719.73866  -4.93 1.86 
+%4720.73430 -14.79 1.46 
+%4721.73009  -9.86 1.33 
+%4722.74036   3.80 1.58 
+%4723.73908   6.85 1.62 
+%4903.15067 -14.41 2.08 
+%5021.91811 -10.11 1.43 
+%5022.91286   1.39 1.86 
+%5024.95012   9.32 1.57 
+%5049.86010   4.91 0.90 
+%5050.80706  13.45 0.79 
+%5051.84043   5.28 1.17 
+%5052.86710  -8.12 2.09 
+%5053.79905 -10.89 0.80 
+%5201.15720  16.10 1.58 
+%5202.15734   9.80 1.61 
+%5203.14807  -8.37 1.53 
+%5258.15584   2.44 2.22 
+%5260.13379  16.54 2.11 
+%5338.81312  -4.84 0.53 
+%5338.90357  -3.10 1.59 
+%5338.98795  -1.62 1.33 
+%5339.06181  -1.56 1.82 
+%5340.05371   9.91 1.63 
+%5340.86600  11.62 1.38 
+%5340.95260   9.87 0.58 
+%5341.03351   6.69 0.76 
+%5341.79524  -1.63 0.80 
+%5342.00395  -1.67 0.88 
+%5342.06383  -5.26 1.46 
+%5369.93610 -15.90 1.40 
+%5370.87580 -14.21 1.48 
+%5371.93429   6.73 1.13 
+%5408.89194  -0.40 2.00 
+%5409.76684  15.21 1.48 
+%5409.86682  15.68 1.98 
+%3.152712894000113010e+3 -1.076949579832486847e+01 1.100000000000000089e+00
+%3.158663459999952465e+03 -1.927949579832421989e+01 1.199999999999999956e+00
+%3.511773341000080109e+03 -7.469495798325453961e+00 1.199999999999999956e+00
+%3.520744746000040323e+03 9.360504201675112768e+00 1.300000000000000044e+00
+%3.573512039000168443e+03 -7.194957983251981659e-01 1.199999999999999956e+00
+%3.574522328999824822e+03 8.800504201674996807e+00 1.100000000000000089e+00
+%3.575480748999863863e+03 3.360504201674885394e+00 1.000000000000000000e+00
+%3.576536046000197530e+03 -7.339495798325046394e+00 1.000000000000000000e+00
+%3.577592602999880910e+03 -1.124949579832446034e+01 1.199999999999999956e+00
+%3.578510712999850512e+03 -8.694957983248485789e-01 9.000000000000000222e-01
+%3.578629602000117302e+03 1.740504201675818763e+00 1.100000000000000089e+00
+%3.579462557000108063e+03 1.217050420167531399e+01 9.000000000000000222e-01
+%3.579621048999950290e+03 1.388050420167452614e+01 1.000000000000000000e+00
+%3.585461769999936223e+03 5.590504201675727813e+00 1.000000000000000000e+00
+%3.586465158000122756e+03 -3.939495798324088582e+00 6.999999999999999556e-01
+%3.587464703999925405e+03 -1.753949579832436712e+01 1.399999999999999911e+00
+%3.588538062000181526e+03 -8.759495798324579141e+00 2.299999999999999822e+00
+%3.589462024000007659e+03 4.800504201675437344e+00 6.999999999999999556e-01
+%3.590463895000051707e+03 1.240050420167548850e+01 8.000000000000000444e-01
+%3.591466483999975026e+03 6.640504201675057061e+00 6.999999999999999556e-01
+%3.592464813000056893e+03 -6.219495798324814473e+00 8.000000000000000444e-01
+%3.606551678999792784e+03 1.405050420167519576e+01 1.899999999999999911e+00
+%3.607507526999805123e+03 9.530504201675782383e+00 1.000000000000000000e+00
+%3.608482642999850214e+03 -4.709495798325136207e+00 1.199999999999999956e+00
+%3.609488452000077814e+03 -1.422949579832533118e+01 1.500000000000000000e+00
+%3.757877319000195712e+03 3.670504201675584000e+00 1.000000000000000000e+00
+%3.760875475000124425e+03 -3.069495798325050373e+00 1.300000000000000044e+00
+%3.761859215999953449e+03 6.250504201675610716e+00 1.300000000000000044e+00
+%3.811846942999865860e+03 4.690504201676048979e+00 1.199999999999999956e+00
+%3.813827016999945045e+03 -1.071949579832498500e+01 9.000000000000000222e-01
+%3.830836957000195980e+03 -2.109495798324090288e+00 9.000000000000000222e-01
+%3.862701441000215709e+03 -5.994957983244120214e-01 9.000000000000000222e-01
+%3.864713661999907345e+03 1.355050420167458469e+01 1.000000000000000000e+00
+%3.867752171000000089e+03 -9.449495798325102669e+00 1.100000000000000089e+00
+%3.870696603000164032e+03 5.030504201675611853e+00 1.100000000000000089e+00
+%3.882657763000112027e+03 -1.153949579832413974e+01 9.000000000000000222e-01
+%3.887690737999975681e+03 -5.299495798324116436e+00 9.000000000000000222e-01
+%3.918621751000173390e+03 7.230504201675813647e+00 1.100000000000000089e+00
+%3.920594947000034153e+03 -1.805949579832422103e+01 1.000000000000000000e+00
+%3.945543122000060976e+03 9.820504201675461786e+00 1.000000000000000000e+00
+%3.951485927000176162e+03 5.480504201674563092e+00 9.000000000000000222e-01
+%3.975471595999784768e+03 -7.649495798323968643e+00 1.000000000000000000e+00
+%3.979543974999804050e+03 -8.039495798325191345e+00 1.199999999999999956e+00
+%4.166874181999824941e+03 -1.481949579832431141e+01 1.100000000000000089e+00
+%4.170853961000218987e+03 7.430504201675347531e+00 9.000000000000000222e-01
+%4.194872349000070244e+03 -1.367949579832483664e+01 1.100000000000000089e+00
+%4.196750384000129998e+03 1.457050420167504967e+01 1.199999999999999956e+00
+%4.197845038999803364e+03 1.538050420167458299e+01 1.199999999999999956e+00
+%4.198855508999899030e+03 -3.269495798324584257e+00 1.199999999999999956e+00
+%4.199732869000174105e+03 -5.909495798324115867e+00 1.000000000000000000e+00
+%4.200910914999898523e+03 -2.694957983244705702e-01 1.000000000000000000e+00
+%4.201868553999811411e+03 9.050504201676190519e+00 1.000000000000000000e+00
+%4.202882594999857247e+03 1.327050420167452671e+01 1.000000000000000000e+00
+%4.228741560999769717e+03 8.380504201674909837e+00 1.100000000000000089e+00
+%4.229700478000100702e+03 1.080050420167566472e+01 1.399999999999999911e+00
+%4.230762136000208557e+03 -1.799495798325168039e+00 1.000000000000000000e+00
+%4.234645916000008583e+03 1.381050420167539983e+01 1.100000000000000089e+00
+%4.253633167999796569e+03 -9.309495798325073679e+00 1.000000000000000000e+00
+%4.254664809999987483e+03 -4.359495798324175553e+00 1.000000000000000000e+00
+%4.291568849999923259e+03 -7.179495798323998201e+00 1.300000000000000044e+00
+%4.292590813999995589e+03 -4.949579832391748369e-02 9.000000000000000222e-01
+%4.293625868000090122e+03 9.980504201674733622e+00 1.000000000000000000e+00
+%4.295639448000118136e+03 -1.049949579832443192e+01 1.100000000000000089e+00
+%4.296606114000082016e+03 -2.112949579832523739e+01 1.199999999999999956e+00
+%4.297641939000226557e+03 -7.539495798324580278e+00 1.000000000000000000e+00
+%4.298567600000184029e+03 7.790504201675929608e+00 1.100000000000000089e+00
+%4.299622194999828935e+03 1.060050420167613083e+01 1.500000000000000000e+00
+%4.300619109999854118e+03 1.005042016757329293e-01 1.000000000000000000e+00
+%4.315507493999786675e+03 1.331050420167478876e+01 1.600000000000000089e+00
+%4.317480847000144422e+03 -8.449495798323880535e+00 1.000000000000000000e+00
+%4.319490526999812573e+03 2.190504201674770002e+00 1.300000000000000044e+00
+%4.320544071999844164e+03 1.182050420167612970e+01 1.000000000000000000e+00
+%4.323507050000131130e+03 -1.125949579832408176e+01 3.200000000000000178e+00
+%4.340555780999828130e+03 -1.549495798323974327e+00 9.000000000000000222e-01
+%4.342486204000189900e+03 1.844050420167597792e+01 1.000000000000000000e+00
+%4.349515163000207394e+03 -1.179949579832495488e+01 1.000000000000000000e+00
+%4.530855659999884665e+03 6.150504201675843774e+00 1.000000000000000000e+00
+%4.550831273999996483e+03 6.140504201676222351e+00 9.000000000000000222e-01
+%4.553803722000215203e+03 -1.330949579832463314e+01 8.000000000000000444e-01
+%4.563838002000004053e+03 -3.339495798325486930e+00 9.000000000000000222e-01
+%4.566761146999895573e+03 1.505042016756164003e-01 1.100000000000000089e+00
+%4.567791670999955386e+03 8.300504201676162097e+00 1.000000000000000000e+00
+%4.569793302000034600e+03 -1.422949579832533118e+01 1.000000000000000000e+00
+%4.570804248000029474e+03 -1.572949579832538802e+01 1.000000000000000000e+00
+%4.571818376000039279e+03 3.530504201675555009e+00 1.100000000000000089e+00
+%4.587861967999953777e+03 2.970504201675439049e+00 1.500000000000000000e+00
+%4.588838798999786377e+03 8.170504201675754530e+00 1.100000000000000089e+00
+%4.589827492999844253e+03 5.160504201676019420e+00 1.100000000000000089e+00
+%4.590819633999839425e+03 -5.959495798323999338e+00 1.000000000000000000e+00
+%4.591817119999788702e+03 -2.045949579832395671e+01 1.500000000000000000e+00
+%4.592827337000053376e+03 -8.829495798325481815e+00 9.000000000000000222e-01
+%4.610742932000197470e+03 1.883050420167542427e+01 1.100000000000000089e+00
+%4.611713477000128478e+03 8.080504201675609011e+00 9.000000000000000222e-01
+%4.616713028000202030e+03 8.550504201675579452e+00 1.300000000000000044e+00
+%4.639686739000026137e+03 -1.078949579832411132e+01 1.100000000000000089e+00
+%4.640654409000184387e+03 -1.055949579832393681e+01 1.300000000000000044e+00
+%4.641631705999840051e+03 2.030504201675498166e+00 1.000000000000000000e+00
+%4.643644998000003397e+03 2.210504201675789204e+00 1.300000000000000044e+00
+%4.644587028000038117e+03 -1.027949579832387883e+01 1.199999999999999956e+00
+%4.646625357000157237e+03 -7.129495798324114730e+00 1.199999999999999956e+00
+%4.647579119000118226e+03 7.150504201675289551e+00 1.100000000000000089e+00
+%4.648484817000105977e+03 9.600504201674908700e+00 1.100000000000000089e+00
+%4.661553706000093371e+03 -1.198949579832486734e+01 1.199999999999999956e+00
+%4.662549407999962568e+03 -7.094957983238003862e-01 1.300000000000000044e+00
+%4.663544865999836475e+03 1.364050420167473021e+01 1.100000000000000089e+00
+%4.664553042000159621e+03 1.352050420167572042e+01 1.199999999999999956e+00
+%4.665569376000203192e+03 5.950504201674533533e+00 1.000000000000000000e+00
+%4.672531723000109196e+03 -5.569495798324552993e+00 1.800000000000000044e+00
+%4.674524120000191033e+03 9.440504201675636864e+00 1.300000000000000044e+00
+%4.677505113999824971e+03 -8.939495798324870179e+00 1.100000000000000089e+00
+%4.678556785000022501e+03 1.560504201675527725e+00 1.100000000000000089e+00
+%4.679504033999983221e+03 1.368050420167499226e+01 1.500000000000000000e+00
+%4.681514142999891192e+03 3.410504201674768865e+00 1.399999999999999911e+00
+%4.682503343000076711e+03 -7.349495798324667817e+00 1.300000000000000044e+00
+%4.701485073999967426e+03 1.412050420167609843e+01 1.199999999999999956e+00
+%4.703513042000122368e+03 -2.069495798323828240e+00 1.199999999999999956e+00
+%4.708479050000198185e+03 -4.589495798324350062e+00 1.199999999999999956e+00
+%4.721473032999783754e+03 -1.558949579832535903e+01 1.199999999999999956e+00
+%4.722472370999865234e+03 2.590504201675614127e+00 1.100000000000000089e+00

### 2014-11-23 Paper/ms.tex
Lookup table

@@
 
-\section{Orbit Simulations}\label{sec:orbits}
+\section{Orbit Look-Up Table}\label{sec:orbits}
 The expected (noise-free) signal due to an exoplanet is periodic, but

### 2014-12-10 Paper/ms.tex
Period renamed to P

--- 2014-11-22 Paper/ms.tex
@@
 the properties of each planet. For simplicity, we describe each planet by
-five parameters: $T_i$, the orbital period, $A_i$, the amplitude (in metres
+five parameters: $P_i$, the orbital period, $A_i$, the amplitude (in metres
 per second) of the RV signal, $\phi_i$, the phase of the signal, $e_i$, the eccentricity of the orbit, and $\theta_i$, a ``viewing angle'' parameter
@@
 where $N$ is the number of planets,
-$\boldsymbol{\alpha} = \{\mu_T, \sigma_T, \mu_A\}$ are the
+$\boldsymbol{\alpha} = \{\mu_P, \sigma_P, \mu_A\}$ are the
 hyperparameters used to define the prior for the properties of the planets,
-and $\psi_i = \{T_i, A_i, \phi_i, e_i, \theta_i\}$
+and $\psi_i = \{P_i, A_i, \phi_i, e_i, \theta_i\}$
 are the properties of planet $i$.
@@
 preferred value, but since these have very heavy tails, the assumption is
-quite fail safe. For example, the prior for $\mu_T$, the typical orbital period,
+quite fail safe. For example, the prior for $\mu_P$, the typical orbital period,
 is centered around 1 year.
@@
 $N$		& Number of planets	& Uniform$(\{0, 1, ..., N_{\rm max}\})$\\
-$\mu_T$		&	Median orbital period	& $\log(\mu_T) \sim$ Cauchy$(5.9, 1)T(-15.3, 27.1)$\\
-$w_T$		&	Diversity of orbital periods & $w_T \sim$ Uniform$(0.1, 3)$\\
+$\mu_P$		&	Median orbital period	& $\log(\mu_P) \sim$ Cauchy$(5.9, 1)T(-15.3, 27.1)$\\
+$w_P$		&	Diversity of orbital periods & $w_P \sim$ Uniform$(0.1, 3)$\\
 $\mu_A$		&	Mean amplitude	& $\log(\mu_A) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
@@
 {\bf Planet Parameters}\\
-$T_i$		&	Orbital period	&	$\log(T_i) \sim $ Biexponential($\log(\mu_T), w_T$)\\
+$P_i$		&	Orbital period	&	$\log(P_i) \sim $ Biexponential($\log(\mu_P), w_P$)\\
 $A_i$		&	Amplitude of signal	& Exponential$(\textnormal{mean}=\mu_A)$\\

### 2014-12-11 Paper/ms.tex
More words on t distribution

@@
 parameters allowing for a non-gaussian distribution for the noise in the
-measurements. When the model-predicted RV signal is $m(t)$, the sampling
-distribution for each measurement was a student-$t$ distribution with
-scale parameter $\sqrt{\sigma_i^2 + s_0^2}$ and shape parameter $\nu$. In the limit
-$(s_0 \to 0, \nu \to \inf)$ the sampling distribution reduces to the standard
-assumption of a normal distribution with standard deviation $\sigma_i$ known
-from the error bars in the data set.
+measurements.
+
+A standard assumption for the probability distribution of the data given the
+parameters (known as the sampling distribution, which becomes the likelihood
+function when the dataset is known) is a normal distribution with
+standard deviation $\sigma_i$ known from the error bars in the data set.
+However, it is usually recommended to put in ``safety features'', in case the
+data set contains any discrepant measurements, or in case the error bars in the
+data set are underestimated. To achieve this, we used a student-$t$ distribution
+instead of a normal distribution, with
+scale parameter $\sqrt{\sigma_i^2 + s_0^2}$ and shape parameter $\nu$. The
+parameter $s_0$ is an ``extra noise'' parameter that effectively increases the
+size of the error bars, and the shape parameter $\nu$ allows for heavier tails
+than a normal distribution. If $\nu$ is large, the student-$t$ distribution
+is approximately a normal distribution, and if $\nu$ is small the noise
+distribution has much heavier tails. For instance, when $\nu=1$ the student-$t$
+distribution becomes a Cauchy distribution.
 

### 2014-12-13 Paper/ms.tex
Commas

@@
 the properties of each planet. For simplicity, we describe each planet by
-five parameters: $P_i$, the orbital period, $A_i$, the amplitude (in metres
-per second) of the RV signal, $\phi_i$, the phase of the signal, $e_i$, the eccentricity of the orbit, and $\theta_i$, a ``viewing angle'' parameter
-describing the position of the observer with respect to the major axis of the
-star's elliptical trajectory.
+five parameters: the orbital period $P_i$, the semi-amplitude (in metres
+per second) of the RV signal $A_i$, the phase of the signal
+$\phi_i$, the eccentricity $e_i$, and the ``viewing angle'' $\theta_i$.
 
@@
 
-The prior for the velocity amplitudes $A_i$ is important and will affect the
+The prior for the velocity semi-amplitudes $\{A_i\}$ is important and will affect the
 inferred number of planets $N$. For example, if the prior allows for very
-low amplitude planets to exist, the data will not rule this out. The prior
+low amplitude planets to exist, the data will not rule this out, and the posterior
+for $N$ will include the possibility of large numbers of planets. The prior
 will influence how many of these low amplitude planets will exist: if we

### 2014-12-13 Paper/ms.tex
More words about sampling distribution

@@
 where $p(D | \theta)$ is called the likelihood, once the actual dataset
-is substituted in.
+is substituted in. Note that some authors do not distinguish between a
+sampling distribution and a likelihood. Throughout this paper we use the term
+sampling distribution for $p(D|\theta)$ if we are discussing a probability
+distribution (actually a family of them, indexed by $\theta$)
+over the set of possible datasets. We use the term {\it likelihood}
+when the actual dataset has been plugged in, when $p(D|\theta)$
+becomes a scalar function (not a probability distribution)
+over the parameter space.
+
 The denominator, often called the {\it evidence} or

### 2014-12-13 Paper/ms.tex
Interpretation of KL divergence

@@
 The marginal likelihood for our model was
-$\log(\mathcal{Z}) \approx -217.65$ and the information
-(Kullback-Leibler divergence from the prior to the posterior)
-was $\mathcal{H} \approx 77.7$ nats.
+$\log(\mathcal{Z}) \approx -217.65$. Nested Sampling also allows for calculation
+of the ``information'', or Kullback-Leibler divergence from the prior to the
+posterior, which quantifies how much we learned about the parameters:
+\begin{eqnarray}
+\mathcal{H} &=& \int p(\theta | D) \log\left[\frac{p(\theta | D)}{p(\theta)}\right] \, d\theta
+\end{eqnarray}
+An intuitive interpretation of this quantity is the number of times the
+prior distribution had to be compressed by a factor of $e$ (if the logarithm
+in the formula is a natural logarithm) to get to the posterior distribution.
+For the $\nu$ Oph data the information was $\mathcal{H} \approx 77.7$ nats or
+111 bits, so the posterior is roughly $e^{-77}$ times the volume of the prior.
 

### 2014-12-13 Paper/ms.tex
nats

@@
 in the formula is a natural logarithm) to get to the posterior distribution.
-For the $\nu$ Oph data the information was $\mathcal{H} \approx 77.7$ nats or
-111 bits, so the posterior is roughly $e^{-77}$ times the volume of the prior.
+For the $\nu$ Oph data the information was $\mathcal{H} \approx 77.7$ nats
+(natural units) or 111 bits, so the posterior is roughly $e^{-77}$ times the
+volume of the prior.
 

### 2014-12-13 Paper/ms.tex
Comment on kl divergence

@@
 $\log(\mathcal{Z}) \approx -217.65$. Nested Sampling also allows for calculation
-of the ``information'', or Kullback-Leibler divergence from the prior to the
-posterior, which quantifies how much we learned about the parameters:
+of the ``information'', or Kullback-Leibler divergence (a quantity in information
+theory) from the prior to the posterior, which quantifies how much we learned about the parameters:
 \begin{eqnarray}

### 2014-12-13 Paper/ms.tex
sigma_extra

@@
 \begin{eqnarray}
-\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\psi}\}_{i=1}^N, y_0, s, \nu\right\}
+\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\psi}\}_{i=1}^N, y_0, \sigma_{\rm extra}, \nu\right\}
 \end{eqnarray}
@@
 are the properties of planet $i$.
-The parameter $y_0$ describes a DC offset in the data, and $s$ and $\nu$ are
-parameters allowing for a non-gaussian distribution for the noise in the
-measurements.
+The parameter $y_0$ describes a DC offset in the data, and
+$\sigma_{\rm extra}$ and $\nu$ are parameters of the noise distribution.
 
@@
 instead of a normal distribution, with
-scale parameter $\sqrt{\sigma_i^2 + s_0^2}$ and shape parameter $\nu$. The
-parameter $s_0$ is an ``extra noise'' parameter that effectively increases the
+scale parameter $\sqrt{\sigma_i^2 + \sigma_{\rm extra}^2}$ and shape parameter $\nu$. The
+parameter $\sigma_{\rm extra}$ is an ``extra noise'' parameter that effectively increases the
 size of the error bars, and the shape parameter $\nu$ allows for heavier tails
@@
 All of the model assumptions are specified in detail in
-Table~\ref{tab:priors}. Cauchy distributions were used when there is a
+Table~\ref{tab:priors}. Most of these priors were chosen to represent vague
+prior knowledge, rather than the judgement of an informed expert on extrasolar
+planets.
+
+Cauchy distributions were used when there is a
 preferred value, but since these have very heavy tails, the assumption is

### 2014-12-13 Paper/ms.tex
Started simulated data section

@@
 
-\section{Reanalysis of $\nu$ Oph}
+\section{Demonstration on Simulated Data}
+To test our proposed methodology, we generated a simulated dataset for a
+system with $N=5$ planets. The dataset was ``inspired by'' the $\nu$ Oph
+dataset (Section~\ref{sec:nu_oph}), and contains two large signals with
+periods of 531 and 3180 days, whose semi-amplitudes are
+322 ms$^{-1}$ and 214 ms$^{-1}$ respectively. The other three planets have
+much lower semi-amplitudes, ranging from 1-9 ms$^{-1}$. The standard deviation
+for the noise in the data was 5 ms$^{-1}$, so some of these low-amplitude
+signals should be detectable.
+
+\section{Application to $\nu$ Oph}\label{sec:nu_oph}
 The $\nu$ Oph system is generally accepted to have two confirmed planets
@@
 
-\section{Reanalysis of Gliese 581}
+\section{Application to Gliese 581}
 The red dwarf star Gliese 581 is thought to host several planets. Exactly

### 2014-12-13 Paper/ms.tex
More work in phase transition section

@@
 
-Even if the slab contains a very small amount of posterior mass, its existence
-can still cause problems with the calculation of the marginal likelihood
+If the slab contains a very small amount of posterior probability, it is not
+a problem if an MCMC algorithm spends all its time in the spike. However, this
+situation could still cause problems with the calculation of the marginal likelihood
 if annealing methods are used. The thermodynamic integral formula gives
-the log of the marginal likelihood as an average of log likelihoods:
+the log of the marginal likelihood $\mathcal{Z}$ as an average of log likelihoods:
 \begin{eqnarray}
@@
 distribution with ``inverse temperature'' $\beta$, proportional to
-$\pi(\theta)L(\theta)^\beta$. Even if the slab contains virtually zero mass
-at $\beta=1$ (i.e. the posterior), at some inverse temperature between 0 and 1
-the slab and the spike ought to contribute equally to the distribution. At
-this critical temperature the MCMC will fail to mix and will give a misleadingly
-low estimate of the average log likelihood at that temperature.
+$\pi(\theta)L(\theta)^\beta$. Even if the slab contains virtually zero probability
+when $\beta=1$ (i.e. the posterior), for some values of the inverse temperature
+$\beta$ the slab and the spike will both be important. At
+these temperatures the MCMC will fail to mix (it will incorrectly spend all its
+time in either the slab or the spike, rather than mixing between the two)
+and will give a misleading estimate of the average log likelihood at that temperature.
 
@@
 Nested Sampling, and variants such as DNS, are not affected
-by phase transitions. Part of their output, the relationship between the
-likelihood $L$ and the enclosed prior mass $X$, can be used to diagnose whether the
+by phase transitions because the exploration only makes use of likelihood
+{\it rankings}, rather than likelihood values themselves.
+Part of their output, the relationship between the
+likelihood $L$ and the enclosed prior mass
+$X(L) = \int \pi(\theta) \mathds{1}\left[L(\theta) > L\right]\, d^n\theta$
+can be used to diagnose whether the
 problem contains a phase transition. In particular, if the graph of $\log(L)$
-vs. $\log(X)$ is convex at some point, then a phase transition exists.
+vs. $\log(X)$ is convex at some point, then a phase transition exists
+\citep{skilling}.
 

### 2014-12-14 Paper/ms.tex
DC offset is m_0
Fake dataset

@@
 \begin{eqnarray}
-\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\psi}\}_{i=1}^N, y_0, \sigma_{\rm extra}, \nu\right\}
+\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\psi}\}_{i=1}^N, m_0, \sigma_{\rm extra}, \nu\right\}
 \end{eqnarray}
@@
 are the properties of planet $i$.
-The parameter $y_0$ describes a DC offset in the data, and
+The parameter $m_0$ describes a DC offset in the data, and
 $\sigma_{\rm extra}$ and $\nu$ are parameters of the noise distribution.
@@
 for the noise in the data was 5 ms$^{-1}$, so some of these low-amplitude
-signals should be detectable.
+signals should be detectable. The simulated data is shown in
+Figure~\ref{fig:fake_data}, along with a model fit.
+
+\begin{figure}
+\includegraphics[scale=0.4]{Figures/fake_data.eps}
+\caption{A simulated dataset that ``resembles'' the $\nu$ Oph dataset, which
+we used to test our methodology. The dominant signal is from two large
+planets with periods periods of 531 and 3180 days and semi-amplitudes of
+322 ms$^{-1}$ and 214 ms$^{-1}$ respectively. There are also three
+much smaller planets which contribute small additional effects to the data.
+\label{fig:fake_data}}
+\end{figure}
 

### 2014-12-14 display.py
Tighter plot margins and log amplitude

@@ -24,22 +24,23 @@ E = E[which].flatten()
 
 hist(T/log(10.), 300, alpha=0.5)
 xlabel(r'$\log_{10}$(Period/days)')
-xlim([0, 4])
+xlim([1, 4])
 for i in xrange(1008, 1008 + int(truth[1007])):
   axvline(truth[i]/log(10.), color='r')
 show()
 
 subplot(2,1,1)
-plot(T/log(10.), A, 'b.', markersize=1)
+plot(T/log(10.), log10(A), 'b.', markersize=1)
 hold(True)
-xlim([0, 4])
-ylabel('Amplitude (m/s)')
-plot(truth[1008:1008 + int(truth[1007])]/log(10.), truth[1018:1018 + int(truth[1007])], 'ro')
+xlim([1, 4])
+ylim([-1, 3])
+ylabel(r'$\log_{10}$[Amplitude (m/s)$]$')
+plot(truth[1008:1008 + int(truth[1007])]/log(10.), log10(truth[1018:1018 + int(truth[1007])]), 'ro')
 
 subplot(2,1,2)
 plot(T/log(10.), E, 'b.', markersize=1)
 hold(True)
-xlim([0, 4])
+xlim([1, 4])
 xlabel(r'$\log_{10}$(Period/days)')
 ylabel('Eccentricity')
 plot(truth[1008:1008 + int(truth[1007])]/log(10.), truth[1038:1038 + int(truth[1007])], 'ro')

### 2014-12-14 display.py
Figure tweaks

@@ -1,12 +1,17 @@
 from pylab import *
 import os
 
+rc("font", size=14, family="serif", serif="Computer Sans")
+rc("text", usetex=True)
+
 data = loadtxt('fake_data_like_nuoph.txt')
 truth = loadtxt('fake_data_like_nuoph.truth')
 posterior_sample = atleast_2d(loadtxt('posterior_sample.txt'))
 
 hist(posterior_sample[:,1007], 100)
 xlabel('Number of Planets')
+ylabel('Number of Posterior Samples')
+xlim([-0.5, 10.5])
 show()
 
 T = posterior_sample[:,1008:1018]
@@ -22,28 +27,29 @@ E = E[which].flatten()
 #iqr = right - left
 #s = s[logical_and(s > middle - 5*iqr, s < middle + 5*iqr)]
 
-hist(T/log(10.), 300, alpha=0.5)
+hist(T/log(10.), 500, alpha=0.5)
 xlabel(r'$\log_{10}$(Period/days)')
 xlim([1, 4])
 for i in xrange(1008, 1008 + int(truth[1007])):
   axvline(truth[i]/log(10.), color='r')
+ylabel('Number of Posterior Samples')
 show()
 
 subplot(2,1,1)
-plot(T/log(10.), log10(A), 'b.', markersize=1)
+plot(truth[1008:1008 + int(truth[1007])]/log(10.), log10(truth[1018:1018 + int(truth[1007])]), 'ro', markersize=7)
 hold(True)
 xlim([1, 4])
 ylim([-1, 3])
 ylabel(r'$\log_{10}$[Amplitude (m/s)$]$')
-plot(truth[1008:1008 + int(truth[1007])]/log(10.), log10(truth[1018:1018 + int(truth[1007])]), 'ro')
+plot(T/log(10.), log10(A), 'b.', markersize=1)
 
 subplot(2,1,2)
-plot(T/log(10.), E, 'b.', markersize=1)
+plot(truth[1008:1008 + int(truth[1007])]/log(10.), truth[1038:1038 + int(truth[1007])], 'ro', markersize=7)
 hold(True)
 xlim([1, 4])
 xlabel(r'$\log_{10}$(Period/days)')
 ylabel('Eccentricity')
-plot(truth[1008:1008 + int(truth[1007])]/log(10.), truth[1038:1038 + int(truth[1007])], 'ro')
+plot(T/log(10.), E, 'b.', markersize=1)
 show()
 
 data[:,0] -= data[:,0].min()

### 2014-12-15 Paper/ms.tex
New figures, fake data results
Fake data truth

@@
 signals should be detectable. The simulated data is shown in
-Figure~\ref{fig:fake_data}, along with a model fit.
+Figure~\ref{fig:fake_data}, along with the true radial velocity curve $m(t)$
+that was used to generate the data.
 
@@
 \label{fig:fake_data}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.4]{Figures/fake_data_N.eps}
+\caption{\label{fig:fake_data_N}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.4]{Figures/fake_data_periods.eps}
+\caption{\label{fig:fake_data_periods}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.4]{Figures/fake_data_posterior.eps}
+\caption{\label{fig:fake_data_posterior}}
 \end{figure}

### 2014-12-15 Paper/ms.tex
Working on the fake data section

@@
 
-\section{Demonstration on Simulated Data}
+\section{Demonstration on Simulated Data}\label{sec:fake_data}
 To test our proposed methodology, we generated a simulated dataset for a
@@
 periods of 531 and 3180 days, whose semi-amplitudes are
-322 ms$^{-1}$ and 214 ms$^{-1}$ respectively. The other three planets have
+322 m s$^{-1}$ and 214 m s$^{-1}$ respectively. The other three planets have
 much lower semi-amplitudes, ranging from 1-9 ms$^{-1}$. The standard deviation
-for the noise in the data was 5 ms$^{-1}$, so some of these low-amplitude
+for the noise in the data was 5 m s$^{-1}$, so some of these low-amplitude
 signals should be detectable. The simulated data is shown in
@@
 planets with periods periods of 531 and 3180 days and semi-amplitudes of
-322 ms$^{-1}$ and 214 ms$^{-1}$ respectively. There are also three
+322 m s$^{-1}$ and 214 m s$^{-1}$ respectively. There are also three
 much smaller planets which contribute small additional effects to the data.
@@
 
+We ran the MCMC on the simulated dataset to obtain the posterior distribution
+for the model parameters. The posterior distribution for $N$, the number of
+planets, is shown in Figure~\ref{fig:fake_data_N}. The true number of planets,
+5, is also the most probable value according to the posterior distribution,
+simply by chance. The posterior distribution also suggests that $N=4$ has
+substantial probability.
+The posterior distribution for the periods $\{P_i\}_{i=1}^N$ is shown in
+Figure~\ref{fig:fake_data_periods}. Because of the label switching degeneracy,
+the posterior distribution for each period is identical, so we pooled the
+results for all periods. Defining the log-periods by
+$S_i =  \log_{10}\left[P_i/(\textnormal{1 day})\right]$,
+Figure~\ref{fig:fake_data_periods} is a Monte Carlo representation
+of the mixture distribution
+\begin{eqnarray}
+f(S) &=& \sum_{N=0}^{10} p(N | D)\sum_{i=1}^N p(S_i | N, D).
+\end{eqnarray}
+The posterior distribution for the periods, shown in
+Figure~\ref{fig:fake_data_periods}, shows that four of the true periods were
+recovered, with probability close to 1. One period which was actually present
+was not detected because it had a very small amplitude. There is also some
+evidence for periods which did not exist, however the posterior probabilities
+for these peaks are not close to 1.
+
 \begin{figure}
 \includegraphics[scale=0.4]{Figures/fake_data_N.eps}
-\caption{\label{fig:fake_data_N}}
+\caption{The posterior distribution for the number of planets $N$ given the
+simulated dataset. The true number of planets was 5, which is also the most
+probable value (this is a coincidence and will not occur in general).
+\label{fig:fake_data_N}}
 \end{figure}
@@
 \includegraphics[scale=0.4]{Figures/fake_data_periods.eps}
-\caption{\label{fig:fake_data_periods}}
+\caption{The posterior distribution for the periods\label{fig:fake_data_periods}}
 \end{figure}
@@
 The posterior distribution for the logarithms of the periods is shown in
-Figure~\ref{fig:nu_oph_periods}. Because of the label-switching degeneracy,
+Figure~\ref{fig:nu_oph_periods}. As in Section~\ref{sec:fake_data},
 the posterior samples for all periods were combined to make this figure, which
-shows three prominent peaks. Defining the log-periods by
-$S_i =  \log_{10}\left[T/(\textnormal{1 day})\right]$,
-Figure~\ref{fig:nu_oph_periods} is a Monte Carlo representation
-of the mixture distribution
-\begin{eqnarray}
-f(S) &=& \sum_{N=0}^{10} p(N | D)\sum_{i=1}^N p(S_i | N, D).
-\end{eqnarray}
+shows three prominent peaks. 
 The two rightmost peaks are the commonly accepted periods, and the two next

### 2014-12-15 Paper/ms.tex
More work on simulated data section

@@
 
-We ran the MCMC on the simulated dataset to obtain the posterior distribution
-for the model parameters. The posterior distribution for $N$, the number of
+We ran the MCMC on the simulated dataset to obtain samples from the
+posterior distribution. We obtained 590 samples.
+The posterior distribution for $N$, the number of
 planets, is shown in Figure~\ref{fig:fake_data_N}. The true number of planets,
@@
 the posterior distribution for each period is identical, so we pooled the
-results for all periods. Defining the log-periods by
+samples for all periods. Defining the log-periods by
 $S_i =  \log_{10}\left[P_i/(\textnormal{1 day})\right]$,
@@
 Figure~\ref{fig:fake_data_periods}, shows that four of the true periods were
-recovered, with probability close to 1. One period which was actually present
-was not detected because it had a very small amplitude. There is also some
-evidence for periods which did not exist, however the posterior probabilities
-for these peaks are not close to 1.
+recovered, with probability close to 1. One period (with $\log_{10}P_i \approx 1.8$)
+which was actually present was not ``detected'' because it had a very small
+amplitude. We note that the posterior probability near this period should not be
+precisely zero. There is also some
+evidence for periods which did not actually exist, however the posterior
+probabilities for these peaks are not close to 1.
 
@@
 \includegraphics[scale=0.4]{Figures/fake_data_periods.eps}
-\caption{The posterior distribution for the periods\label{fig:fake_data_periods}}
-\end{figure}
+\caption{The posterior distribution for the periods given the simulated
+dataset. The solid lines are the true periods, four of which were detected
+with high posterior probability.\label{fig:fake_data_periods}}
+\end{figure}
+
+The joint posterior distribution for the periods and the amplitudes of the
+signals is shown in Figure~\ref{fig:fake_data_posterior} along with the
+eccentricities. As with Figure~\ref{fig:fake_data_periods}, the samples for
+all planets were combined. The true values are also plotted as circles. Clearly, the
+reason the period of $\log_{10}(P_i) \approx 1.8$ was not ``detected'' was that
+it had a very low amplitude of approximately 1 m s$^{-1}$ which is one fifth
+of the noise level.
 
@@
 \includegraphics[scale=0.4]{Figures/fake_data_posterior.eps}
-\caption{\label{fig:fake_data_posterior}}
+\caption{The joint posterior distribution for the periods and the amplitudes
+(top panel) and the periods and eccentricities (bottom panel).
+\label{fig:fake_data_posterior}}
 \end{figure}

### 2014-12-15 Paper/ms.tex
Caption words

@@
 \caption{The joint posterior distribution for the periods and the amplitudes
-(top panel) and the periods and eccentricities (bottom panel).
+(top panel) and the periods and eccentricities (bottom panel) from the
+simulated dataset. The true values are plotted as circles.
 \label{fig:fake_data_posterior}}

### 2014-12-15 Paper/ms.tex
Priors editing

@@
 prior knowledge, rather than the judgement of an informed expert on extrasolar
-planets.
-
-Cauchy distributions were used when there is a
+planets. Uniform distributions were used for parameters such as phases, where
+time-translation symmetry seems plausible.
+Truncated Cauchy distributions were used when there is a
 preferred value, but since these have very heavy tails, the assumption is
-quite fail safe. For example, the prior for $\mu_P$, the typical orbital period,
-is centered around 1 year.
+quite fail safe relative to other possible assignments such as normal distributions.
+For example, the prior for $\mu_P$, the typical orbital period,
+is centered around 1 year but could be as low as $\approx e^{-21}$ years or as high
+as $\approx e^{21}$ years, a very wide range. Truncation intervals are
+specified with the $T(a,b)$ notation in Table~\ref{tab:priors}.
 

### 2014-12-16 Orbit.cpp
Semi amplitude fix (need to redo runs)

@@ -46,18 +46,21 @@ vector<double> Orbit::evaluate(const std::vector<double>& arg_to_sin,
 	double S = sin(viewing_angle);
 	vector<double> vr = vx;
 	double vrmax = -1E300;
+	double vrmin =  1E300;
 	int closest = 0;
 	for(size_t i=0; i<vr.size(); i++)
 	{
 		vr[i] = C*vx[i] + S*vy[i];
-		if(abs(vr[i]) > vrmax)
-			vrmax = abs(vr[i]);
+		if(vr[i] > vrmax)
+			vrmax = vr[i];
+		if(vr[i] < vrmin)
+			vrmin = vr[i];
 		if(abs(vr[i]) < abs(vr[closest]))
 			closest = i;
 	}
 
 	for(size_t i=0; i<vr.size(); i++)
-		vr[i] /= vrmax;
+		vr[i] = 2*(vr[i] - vrmin)/(vrmax - vrmin) - 1.;
 
 	vector<double> result = arg_to_sin;
 	int index;

### 2014-12-16 Orbits/orbit_lookup.r

@@ -18,6 +18,8 @@ evaluate = function(arg_to_sin, v=0.75, viewing_angle=0.)
 
   radial_velocity = cos(viewing_angle)*trajectories[[orbit]][,1] + sin(viewing_angle)*trajectories[[orbit]][,2]
   radial_velocity = radial_velocity/max(abs(radial_velocity))
+  radial_velocity = (radial_velocity - min(radial_velocity))/(max(radial_velocity) - min(radial_velocity))
+  radial_velocity = 2*radial_velocity - 1
   closest_to_zero = which(abs(radial_velocity) == min(abs(radial_velocity)))
 
   cc = 2*pi*(closest_to_zero - 1)/N

### 2014-12-17 Orbit.cpp
RV function like cos not sin

@@ -38,7 +38,7 @@ void Orbit::load(const char* filename)
 	fin.close();
 }
 
-vector<double> Orbit::evaluate(const std::vector<double>& arg_to_sin,
+vector<double> Orbit::evaluate(const std::vector<double>& arg_to_cos,
 						double viewing_angle) const
 {
 	// Create radial velocities
@@ -47,7 +47,7 @@ vector<double> Orbit::evaluate(const std::vector<double>& arg_to_sin,
 	vector<double> vr = vx;
 	double vrmax = -1E300;
 	double vrmin =  1E300;
-	int closest = 0;
+	int highest = 0;
 	for(size_t i=0; i<vr.size(); i++)
 	{
 		vr[i] = C*vx[i] + S*vy[i];
@@ -55,19 +55,19 @@ vector<double> Orbit::evaluate(const std::vector<double>& arg_to_sin,
 			vrmax = vr[i];
 		if(vr[i] < vrmin)
 			vrmin = vr[i];
-		if(abs(vr[i]) < abs(vr[closest]))
-			closest = i;
+		if(vr[i] > vr[highest])
+			highest = i;
 	}
 
 	for(size_t i=0; i<vr.size(); i++)
 		vr[i] = 2*(vr[i] - vrmin)/(vrmax - vrmin) - 1.;
 
-	vector<double> result = arg_to_sin;
+	vector<double> result = arg_to_cos;
 	int index;
-	double cc = 2.*M_PI*((double)(closest))/vr.size();
+	double cc = 2.*M_PI*((double)(highest))/vr.size();
 	for(size_t i=0; i<result.size(); i++)
 	{
-		index = (int)(vr.size()*mod(arg_to_sin[i] + cc, 2.*M_PI)/(2*M_PI));
+		index = (int)(vr.size()*mod(arg_to_cos[i] + cc, 2.*M_PI)/(2*M_PI));
 		result[i] = vr[index];
 	}
 

### 2014-12-17 Orbit.h

@@ -13,7 +13,7 @@ class Orbit
 		void load(const char* filename);
 
 		std::vector<double> evaluate
-				(const std::vector<double>& arg_to_sin,
+				(const std::vector<double>& arg_to_cos,
 						double viewing_angle) const;
 
 		// Unit test

### 2014-12-17 Orbits/orbit_lookup.r

@@ -20,17 +20,21 @@ evaluate = function(arg_to_sin, v=0.75, viewing_angle=0.)
   radial_velocity = radial_velocity/max(abs(radial_velocity))
   radial_velocity = (radial_velocity - min(radial_velocity))/(max(radial_velocity) - min(radial_velocity))
   radial_velocity = 2*radial_velocity - 1
-  closest_to_zero = which(abs(radial_velocity) == min(abs(radial_velocity)))
+  closest_to_max = which(radial_velocity == max(radial_velocity))[1]
 
-  cc = 2*pi*(closest_to_zero - 1)/N
+  cc = 2*pi*(closest_to_max - 1)/N
   index = 1 + N*((arg_to_sin + cc) %% (2*pi))/(2*pi)
 
   return(radial_velocity[index])
 }
 
-t = seq(-10, 10, by=0.01)
-y = evaluate(t, v=(0.4 + 0.6*runif(1)), viewing_angle=2.*pi*runif(1))
-plot(t, y, type='l')
-abline(v=2*pi)
-print(any(is.na(y)))
+t = seq(0, 10, by=0.01)
+y = evaluate(2*pi*t/5, v=sqrt(1. - 0.5), viewing_angle=-0.8)
+plot(t, y)
+y = -evaluate(2*pi*t/5 - 0.1, v=sqrt(1. - 0.5), viewing_angle=pi - 0.8)
+lines(t, y)
+
+
+#data = as.matrix(read.csv("kepler.txt", sep=" "))
+#lines(data[,1], data[,2])
 

### 2014-12-17 Orbits/kepler.py
kepler equation script

@@ -1,0 +1,47 @@
+from pylab import *
+
+t = linspace(0, 10, 10001)
+
+e = 0.5		# Eccentricity
+P = 5.		# Period
+K = 1.		# Semi-amplitude
+phi = 1.	# Longitude of ascending node
+wbar = pi/2	# Longitude of periastron
+omega = 2.*pi/P
+
+# Solve for E
+def update(E, t):
+  E = E - (E - e*sin(E) - omega*t - phi)/(1. - e*cos(E))
+  return E
+
+ion()
+hold(False)
+E = omega*t + phi
+for i in xrange(0, 20):
+  hold(True)
+  plot(t, E)
+  E = update(E, t)
+  title(i+1)
+  draw()
+
+ioff()
+show()
+
+E = mod(E, 2*pi)
+cosf = (cos(E) - e)/(1. - e*cos(E))
+f = arccos(cosf)
+f[E > pi] = 2*pi - f[E > pi]
+v = K*(sin(f + wbar) + e*sin(wbar))
+v -= (v.min() + v.max())/2
+print(v.max() - v.min())
+plot(t, v)
+
+data = empty((10001, 2))
+data[:,0], data[:,1] = t, v
+savetxt('kepler.txt', data)
+show()
+
+# EQUIVALENCE
+# kepler.py			Old code (but using cosine alignment)
+# (phi, wbar) = (0, pi/2)		(phase, viewing angle) = (0, 0)
+# (phi, wbar) = 

### 2014-12-22 Paper/ms.tex
Updated Zs and Hs

@@
 The marginal likelihood for our model was
-$\log(\mathcal{Z}) \approx -217.65$. Nested Sampling also allows for calculation
+$\log(\mathcal{Z}) \approx -220.5$. Nested Sampling also allows for calculation
 of the ``information'', or Kullback-Leibler divergence (a quantity in information
@@
 in the formula is a natural logarithm) to get to the posterior distribution.
-For the $\nu$ Oph data the information was $\mathcal{H} \approx 77.7$ nats
-(natural units) or 111 bits, so the posterior is roughly $e^{-77}$ times the
+For the $\nu$ Oph data the information was $\mathcal{H} \approx 76.6$ nats
+(natural units) or 111 bits, so the posterior is roughly $e^{-76.6}$ times the
 volume of the prior.
@@
 
-The marginal likelihood was $\log(\mathcal{Z}) \approx -618.04$ and the
-information was $\mathcal{H} \approx 140.2$ nats. This compares favourably
+The marginal likelihood was $\log(\mathcal{Z}) \approx -616.3$ and the
+information was $\mathcal{H} \approx 130.4$ nats. This compares favourably
 to the marginal likelihood of $-640.1$ (for a 6-planet model)
@@
 interesting papers on it. We also thank Tom Loredo (Cornell) and Dan
-Foreman-Mackey (NYU) for interesting conversations and feedback.
+Foreman-Mackey (NYU) for interesting conversations and feedback, and Ben
+Montet (Caltech) and Geraint Lewis (Sydney) for helpful discussions.
 

### 2014-12-22 Paper/ms.tex
Nu oph numbers up to date

@@
 for additional signals in
-$\nu$ Oph with periods of 36.12 $\pm$ 0.04 days and
-75.57 $\pm$ 0.76 days; the posterior probability that at least one of these
-exists is 79\%. The results also suggest Gliese 581 hosts many (7-15) ``planets''
+$\nu$ Oph with periods of 36.11 $\pm$ 0.034 days,
+75.58 $\pm$ 0.80 days, and 1709 $\pm$ 183 days;
+the posterior probability that at least one of these
+exists is 85\%. The results also suggest Gliese 581 hosts many (7-15) ``planets''
 (or other causes of other periodic
@@
 The posterior distribution for $N$, the number of planets, is shown in
-Figure~\ref{fig:nu_oph_N}, showing the most probable number of planets is 3,
-and the posterior probability that $N \geq 3$ is about 95\%. Of course, these
+Figure~\ref{fig:nu_oph_N}, showing that $N$ could be anywhere from 2 to 10,
+and the posterior probability that $N \geq 3$ is about 88\%. Of course, these
 extra possible signals are not necessarily a planet but a feature in the data
@@
 the posterior samples for all periods were combined to make this figure, which
-shows three prominent peaks. 
-The two rightmost peaks are the commonly accepted periods, and the two next
-most prominent peaks have periods of 36.12 $\pm$ 0.04 days and
-75.57 $\pm$ 0.76 days.
-The probability
-at least one of these exists (the fraction of our samples which
-have one of these signals) is 79\%.
-Given that they exist, their amplitudes are low, around 10-15 metres per second,
-which we note is above the noise level. We also found that the 75.7 $\pm$ 1.8 day
-period prefers an eccentricity of about $0.37 \pm 0.14$.
-An example model fit to the data is shown in Figure~\ref{fig:nuoph}.
+shows several prominent peaks. 
+The two peaks with vertical dashed lines are the commonly accepted periods of
+530 and 3190 days, and the other prominent peaks (i.e. signals which have a
+moderate probability of existence) have periods of 36.11 $\pm$ 0.034 days,
+75.58 $\pm$ 0.80 days, and 1709 $\pm$ 183 days.
+The probability that
+at least one of these ``extra'' signals exists, given by the fraction of
+samples that have one of these signals, is 85\%.
+Given that they exist, their amplitudes are low, around 5-40 metres per second,
+which we note is above the noise level. An example model fit to the data is
+shown in Figure~\ref{fig:nuoph}.
 
@@
 \caption{The posterior distribution for the number of planets $N$ orbiting
-$\nu$ Oph. The posterior probability of $N > 2$ is about 96\%, however the
+$\nu$ Oph. The posterior probability of $N > 2$ is about 88\%, however the
 prior probability of $N > 2$ was already high due to the uniform prior for $N$.
@@
 of $530.3$ and $3190$ days. The next most prominent peaks with well determined
-periods are at 36.12 $\pm$ 0.04 days and
-75.57 $\pm$ 0.76 days.\label{fig:nu_oph_periods}}
+periods are at log periods of around 1.6, 2.4, and 3.3, corresponding to
+periods of 36.11 $\pm$ 0.034 days,
+75.58 $\pm$ 0.80 days, and 1709 $\pm$ 183 days.\label{fig:nu_oph_periods}}
 \end{figure}
@@
 We applied the code to two well-studied RV datasets, $\nu$ Oph and Gliese 581.
-In $\nu$ Oph, we found some evidence for a signal with a period of
-either 36.12 $\pm$ 0.04 days or
-75.57 $\pm$ 0.76 days: the posterior probability one of these exists, given our
-assumptions, is 79\%. The posterior distribution contains models with several
-signals and models with two, however, these are separated by a phase transition.
-Therefore mixing between the two situations would be very difficult if we simply tried to sample the posterior distribution.
+In $\nu$ Oph, we found some evidence for additional signals with low amplitude,
+but with several possible solutions for their periods. Given our modelling
+assumptions, the posterior probability at least one of these additional signals
+is real is 85\%. The posterior distribution contains models both with and without
+these additional signals, however, these are separated by a phase transition.
+Therefore mixing between the two situations would be infrequent if we simply
+tried to sample the posterior distribution.
 

### 2014-12-22 Paper/ms.tex
Updated fake data results

@@
 To test our proposed methodology, we generated a simulated dataset for a
-system with $N=5$ planets. The dataset was ``inspired by'' the $\nu$ Oph
+system with $N=7$ planets. The dataset was ``inspired by'' the $\nu$ Oph
 dataset (Section~\ref{sec:nu_oph}), and contains two large signals with
-periods of 531 and 3180 days, whose semi-amplitudes are
-322 m s$^{-1}$ and 214 m s$^{-1}$ respectively. The other three planets have
-much lower semi-amplitudes, ranging from 1-9 ms$^{-1}$. The standard deviation
+periods of 530 and 3120 days, whose semi-amplitudes are
+291 m s$^{-1}$ and 181 m s$^{-1}$ respectively. The other five planets have
+much lower semi-amplitudes, ranging from 4-30 ms$^{-1}$. The standard deviation
 for the noise in the data was 5 m s$^{-1}$, so some of these low-amplitude
@@
 we used to test our methodology. The dominant signal is from two large
-planets with periods periods of 531 and 3180 days and semi-amplitudes of
-322 m s$^{-1}$ and 214 m s$^{-1}$ respectively. There are also three
+planets with periods periods of 530 and 3120 days and semi-amplitudes of
+291 m s$^{-1}$ and 181 m s$^{-1}$ respectively. There are also five
 much smaller planets which contribute small additional effects to the data.
@@
 We ran the MCMC on the simulated dataset to obtain samples from the
-posterior distribution. We obtained 590 samples.
+posterior distribution. We obtained 520 posterior samples.
 The posterior distribution for $N$, the number of
 planets, is shown in Figure~\ref{fig:fake_data_N}. The true number of planets,
-5, is also the most probable value according to the posterior distribution,
-simply by chance. The posterior distribution also suggests that $N=4$ has
-substantial probability.
+7, is not the most probable value, but it does have substantial probability.
+The posterior distribution suggests that $N$ could be anywhere from 6 to 10.
 The posterior distribution for the periods $\{P_i\}_{i=1}^N$ is shown in
@@
 The posterior distribution for the periods, shown in
-Figure~\ref{fig:fake_data_periods}, shows that four of the true periods were
-recovered, with probability close to 1. One period (with $\log_{10}P_i \approx 1.8$)
+Figure~\ref{fig:fake_data_periods}, shows that six of the true periods were
+recovered, with probability close to 1. One period (with $\log_{10}P_i \approx 1.15$)
 which was actually present was not ``detected'' because it had a very small
@@
 \caption{The posterior distribution for the number of planets $N$ given the
-simulated dataset. The true number of planets was 5, which is also the most
-probable value (this is a coincidence and will not occur in general).
+simulated dataset. The true number of planets was 7.
 \label{fig:fake_data_N}}
@@
 \caption{The posterior distribution for the periods given the simulated
-dataset. The solid lines are the true periods, four of which were detected
+dataset. The solid lines are the true periods, six of which were detected
 with high posterior probability.\label{fig:fake_data_periods}}
@@
 all planets were combined. The true values are also plotted as circles. Clearly, the
-reason the period of $\log_{10}(P_i) \approx 1.8$ was not ``detected'' was that
-it had a very low amplitude of approximately 1 m s$^{-1}$ which is one fifth
-of the noise level.
+reason the period of $\log_{10}(P_i) \approx 1.15$ was not ``detected'' was that
+it had a very low amplitude of approximately 4 m s$^{-1}$ which is below the
+noise level.
 

### 2014-12-22 Paper/ms.tex
Numbers updated in Gliese 581 section

@@
 There are several phase transitions (concave-up regions) present. The one at
-$\log(X) \approx -75$ separates models which contain additional signals from
-models that do not. Without using Nested Sampling it would be very difficult
+$\log(X) \approx -70$ separates models which contain additional signals from
+models that do not. Without using Nested Sampling it would be more difficult
 to mix between these two situations and calculate the posterior probability
@@
 analysis, the models with a third signal exist to the left of the phase
-transition at $\log(X) \approx -75$, and models without the signal exist to the
+transition at $\log(X) \approx -70$, and models without the signal exist to the
 right of the phase transition. Mixing between these two phases is crucial for
 accurately computing the posterior probability that extra signals exist.
-However, if we had simply tried to sample the posterior distribution, it would
-not mix easily between these two phases and could easily fail to detect the
-extra signals.
 
@@
 and shows strong evidence for at least eight periods
-($P(N \leq 7 | D) = 0.004$).
+($P(N \leq 7 | D) = 0.012$).
 It is now recognised that
@@
 Montet (Caltech) and Geraint Lewis (Sydney) for helpful discussions.
+The referee also provided excellent suggestions for improving the manuscript.
 

### 2014-12-22 Paper/ms.tex
Initials

@@
 \section*{Acknowledgements}
-It is a pleasure to thank Fengji Hou and David Hogg (NYU) for inspiring me to
+It is a pleasure to thank Fengji Hou and David Hogg (NYU) for inspiring me (BJB) to
 finally work on this problem, and Phil Gregory (UBC) for writing so many

### 2014-12-25 Paper/ms.tex
Removed lookup table talk from introduction

@@
 GNU General Public Licence version 3.
-
-Computing the expected RV signal from
-a planet in an elliptical orbit can be slow. Therefore we use a lookup table
-of precomputed orbits, allowing us to do $\sim 15000$ MCMC steps per second
-per CPU core.
 
@@
 Sampling run for each, to get the marginal likelihood as a function of $a$.
-Yet when the hypothesis space for $a$ is discrete to begin with, the wasteful
-alternative has somehow become standard practice.
+When the hypothesis space for $a$ is discrete, MCMC is still possible and there
+is no reason to switch to the wasteful alternative.
 

### 2014-12-25 Orbits/kepler.py
Agreement with Ohta

@@ -5,7 +5,7 @@ t = linspace(0, 10, 10001)
 e = 0.5		# Eccentricity
 P = 5.		# Period
 K = 1.		# Semi-amplitude
-phi = 1.	# Longitude of ascending node
+phi = 0.	# Longitude of ascending node
 wbar = pi/2	# Longitude of periastron
 omega = 2.*pi/P
 
@@ -43,5 +43,5 @@ show()
 
 # EQUIVALENCE
 # kepler.py			Old code (but using cosine alignment)
-# (phi, wbar) = (0, pi/2)		(phase, viewing angle) = (0, 0)
+# (phi, wbar) = (0, pi/2)		(phase, viewing angle) = (0, -pi/2)
 # (phi, wbar) = 

### 2014-12-25 Orbits/orbit_lookup.r

@@ -1,4 +1,4 @@
-
+system("python2 kepler.py")
 
 # LOAD ALL THE FILES
 trajectories = list()
@@ -29,12 +29,9 @@ evaluate = function(arg_to_sin, v=0.75, viewing_angle=0.)
 }
 
 t = seq(0, 10, by=0.01)
-y = evaluate(2*pi*t/5, v=sqrt(1. - 0.5), viewing_angle=-0.8)
+y = evaluate(2*pi*t/5, v=sqrt(1. - 0.5), viewing_angle=-pi/2)
 plot(t, y)
-y = -evaluate(2*pi*t/5 - 0.1, v=sqrt(1. - 0.5), viewing_angle=pi - 0.8)
-lines(t, y)
 
-
-#data = as.matrix(read.csv("kepler.txt", sep=" "))
-#lines(data[,1], data[,2])
+data = as.matrix(read.csv("kepler.txt", sep=" "))
+lines(data[,1], data[,2])
 

### 2014-12-25 Orbits/kepler.py
More coincident points

@@ -5,8 +5,8 @@ t = linspace(0, 10, 10001)
 e = 0.5		# Eccentricity
 P = 5.		# Period
 K = 1.		# Semi-amplitude
-phi = 0.	# Longitude of ascending node
-wbar = pi/2	# Longitude of periastron
+phi = 1.8	# Longitude of ascending node
+wbar = 2.7	# Longitude of periastron
 omega = 2.*pi/P
 
 # Solve for E
@@ -44,4 +44,10 @@ show()
 # EQUIVALENCE
 # kepler.py			Old code (but using cosine alignment)
 # (phi, wbar) = (0, pi/2)		(phase, viewing angle) = (0, -pi/2)
-# (phi, wbar) = 
+# (phi, wbar) = (0, pi)			(phase, viewing angle) = (0.6, -pi)
+# (phi, wbar) = (0, 0.45)		(phase, viewing angle) = (5.9, -0.45)
+# (phi, wbar) = (1.1, 0.45)		(phase, viewing angle) = (0.7, -0.45)
+# (phi, wbar) = (1.1, 0.45)		(phase, viewing angle) = (0.7, -0.45)
+# (phi, wbar) = (1.8, 1.9)		(phase, viewing angle) = (1.9, -1.9)
+# (phi, wbar) = (1.8, 2.7)		(phase, viewing angle) = (2.2, -2.7)
+

### 2014-12-25 Orbits/orbit_lookup.r

@@ -29,7 +29,7 @@ evaluate = function(arg_to_sin, v=0.75, viewing_angle=0.)
 }
 
 t = seq(0, 10, by=0.01)
-y = evaluate(2*pi*t/5, v=sqrt(1. - 0.5), viewing_angle=-pi/2)
+y = evaluate(2*pi*t/5 + 2.2, v=sqrt(1. - 0.5), viewing_angle=-2.7)
 plot(t, y)
 
 data = as.matrix(read.csv("kepler.txt", sep=" "))

### 2014-12-25 Paper/ms.tex
ohta reference

@@
 Since the our orientation with respect to the orbits is unknown, each planet
-requires a ``viewing angle'' parameter $\theta$.
+requires a ``viewing angle'' parameter $\theta$ also known as the longitude
+of the line of sight \citep{ohta}.
 The eccentricity of the orbit, in terms of $v$, is $e = 1 - v^2$.
@@
 O'Hagan, A., Forster,~J., 2004, Bayesian inference. London: Arnold.
+
+\bibitem[\protect\citeauthoryear{Ohta, Taruya, 
+\& Suto}{2005}]{ohta} Ohta Y., Taruya A., Suto Y., 2005, ApJ, 622, 1118
 

### 2014-12-25 Paper/ms.tex
phase definition clarification

@@
 per second) of the RV signal $A_i$, the phase of the signal
-$\phi_i$, the eccentricity $e_i$, and the ``viewing angle'' $\theta_i$.
+$\phi_i$ (defined such that $\psi=0$ gives an RV signal whose maximum is at
+$t=0$),
+the eccentricity $e_i$, and the ``viewing angle'' $\theta_i$.
 

### 2014-12-25 Paper/ms.tex
minor edits

--- 2014-12-25 Paper/ms.tex
@@
 per second) of the RV signal $A_i$, the phase of the signal
-$\phi_i$, the eccentricity $e_i$, and the ``viewing angle'' $\theta_i$.
+$\phi_i$ (defined such that $\psi=0$ gives an RV signal whose maximum is at
+$t=0$), the eccentricity $e_i$, and the ``viewing angle'' $\theta_i$
+(also known as the longitude of the line of sight).
 
@@
 hyperparameters used to define the prior for the properties of the planets,
-and $\psi_i = \{P_i, A_i, \phi_i, e_i, \theta_i\}$
+and $\psi_i = \{P_i, A_i, e_i, \phi_i, \theta_i\}$
 are the properties of planet $i$.
 The parameter $m_0$ describes a DC offset in the data, and
-$\sigma_{\rm extra}$ and $\nu$ are parameters of the noise distribution.
+$\sigma_{\rm extra}$ and $\nu$ are parameters of the noise distribution
+which are discussed further below.
 
@@
 is centered around 1 year but could be as low as $\approx e^{-21}$ years or as high
-as $\approx e^{21}$ years, a very wide range. Truncation intervals are
+as $\approx e^{21}$ years, a very generous range. Truncation intervals are
 specified with the $T(a,b)$ notation in Table~\ref{tab:priors}.
@@
 are not available in closed form for the normal distribution.
-
 The prior for the velocity semi-amplitudes $\{A_i\}$ is important and will affect the

### 2014-12-25 Paper/ms.tex
Circular orbit --> cosine

@@
 per second) of the RV signal $A_i$, the phase of the signal
-$\phi_i$ (defined such that $\psi=0$ gives an RV signal whose maximum is at
+$\phi_i$ (defined such that $\phi=0$ gives an RV signal whose maximum is at
 $t=0$), the eccentricity $e_i$, and the ``viewing angle'' $\theta_i$
-(also known as the longitude of the line of sight).
+(also known as the longitude of the line of sight). We defined our parameters
+such that in the limit of zero eccentricity, the RV signal of a planet
+reduces to $A_i\cos\left(2\pi t/P_i + \phi_i\right)$.
 

### 2014-12-26 Paper/ms.tex
rjobject reference
omega for viewing angle

@@
 $\phi_i$ (defined such that $\phi=0$ gives an RV signal whose maximum is at
-$t=0$), the eccentricity $e_i$, and the ``viewing angle'' $\theta_i$
+$t=0$), the eccentricity $e_i$, and the ``viewing angle'' $\omega_i$
 (also known as the longitude of the line of sight). We defined our parameters
@@
 hyperparameters used to define the prior for the properties of the planets,
-and $\psi_i = \{P_i, A_i, e_i, \phi_i, \theta_i\}$
+and $\psi_i = \{P_i, A_i, e_i, \phi_i, \omega_i\}$
 are the properties of planet $i$.
@@
 $e_i$		&	Eccentricity of orbit	&	Beta$(1, 3.1)T(0, 0.8)$\\
-$\theta_i$	&	Viewing angle	&	Uniform$(0, 2\pi)$\\
+$\omega_i$	&	Viewing angle	&	Uniform$(0, 2\pi)$\\
 \hline
@@
 from the origin, they will measure $m(t) = \dot{x}(t)$. However, if the
-observer is located at an angle $\theta$ with respect to the $x$-axis, then
+observer is located at an angle $\omega$ with respect to the $x$-axis, then
 the radial velocity measured will instead be
-$m(t) = \cos(\theta)\dot{x}(t) + \sin(\theta)\dot{y}(t)$.
+$m(t) = \cos(\omega)\dot{x}(t) + \sin(\omega)\dot{y}(t)$.
 Since the our orientation with respect to the orbits is unknown, each planet
-requires a ``viewing angle'' parameter $\theta$ also known as the longitude
+requires a ``viewing angle'' parameter $\omega$ also known as the longitude
 of the line of sight \citep{ohta}.
@@
 \bibitem[\protect\citeauthoryear{Brewer}{2014}]{rjobject} Brewer, B. J., 2014,
-Bayesian Analysis, submitted.
+submitted. ArXiv: 1411.3921
 

### 2014-12-26 Paper/ms.tex
Eccentricity prior

@@
 
-An apparently strange choice is the conditional prior for the orbital
+An apparently strange choice is the conditional prior for the (logarithms of)
+the orbital
 periods, which is a biexponential distribution. A more conventional choice
@@
 may be many low amplitude planets. However, their other properties, such as
-their orbital periods, will not be well determined.
+their orbital periods, will not be well determined. The Beta prior for
+eccentricity was suggested by Gregory (priv. comm) and is an approximation
+to the inferred frequency distribution of eccentricities in the population
+\citep{kipping}.
 
@@
 \& Hogg}{2014}]{fengji} Hou F., Goodman J., Hogg D.~W., 2014, arXiv, arXiv:1401.6128
+
+\bibitem[\protect\citeauthoryear{Kipping}{2013}]{kipping} 
+Kipping D.~M., 2013, MNRAS, 434, L51
 

### 2014-12-26 Paper/ms.tex
prior words

@@
 the orbital
-periods, which is a biexponential distribution. A more conventional choice
+periods, which is a biexponential distribution given $\mu_P$ and $w_P$.
+Rather than assigning independent priors to the log periods, the hierarchical
+model allows for the periods to ``cluster around'' a typical period $\mu_P$ if
+there is evidence for this. On the other hand, independent priors for the
+periods would imply a strong prior commitment to the hypothesis that the
+periods are spread out across the whole prior volume.
+A more conventional choice for the conditional prior given $\mu_P$ and $w_P$
 would have been a normal distribution. However, the \citet{rjobject} software
@@
 are not available in closed form for the normal distribution.
+
 The prior for the velocity semi-amplitudes $\{A_i\}$ is important and will affect the

### 2014-12-26 Paper/ms.tex
Semi amplitude

@@
 $P_i$		&	Orbital period	&	$\log(P_i) \sim $ Biexponential($\log(\mu_P), w_P$)\\
-$A_i$		&	Amplitude of signal	& Exponential$(\textnormal{mean}=\mu_A)$\\
+$A_i$		&	Semi-mplitude of signal	& Exponential$(\textnormal{mean}=\mu_A)$\\
 $\phi_i$	&	Phase of signal	&	Uniform($0, 2\pi$)\\

### 2014-12-26 Paper/ms.tex
slight edits

@@
 preferred value, but since these have very heavy tails, the assumption is
-quite fail safe relative to other possible assignments such as normal distributions.
+quite fail safe relative to other possible ``informative'' assignments such as normal distributions.
 For example, the prior for $\mu_P$, the typical orbital period,
@@
 the orbital
-periods, which is a biexponential distribution given $\mu_P$ and $w_P$.
+periods, which is a biexponential distribution given a location parameter
+$\mu_P$ and scale parameter, or width, $w_P$.
 Rather than assigning independent priors to the log periods, the hierarchical
@@
 periods would imply a strong prior commitment to the hypothesis that the
-periods are spread out across the whole prior volume.
+periods are spread out across the whole prior volume (equivalent to assuming
+a fixed large value for $w_P$.
 A more conventional choice for the conditional prior given $\mu_P$ and $w_P$

### 2014-12-30 Paper/ms.tex
A bit more prior editing

@@
 planets. Uniform distributions were used for parameters such as phases, where
-time-translation symmetry seems plausible.
+time-translation symmetry seems plausible. For some parameters we assigned the
+distribution in terms of the log of the parameter, rather than the parameter
+itself, when the parameter is positive and uncertain by orders of magnitude.
 Truncated Cauchy distributions were used when there is a
@@
 is centered around 1 year but could be as low as $\approx e^{-21}$ years or as high
-as $\approx e^{21}$ years, a very generous range. Truncation intervals are
-specified with the $T(a,b)$ notation in Table~\ref{tab:priors}.
+as $\approx e^{21}$ years, a very generous range. A uniform distribution for
+$\log(\mu_P)$ would have been more conventional, whereas the Cauchy distribution
+expresses a slight preference for $\mu_P$ being of order one year.
 
@@
 periods are spread out across the whole prior volume (equivalent to assuming
-a fixed large value for $w_P$.
+a fixed large value for $w_P$).
 A more conventional choice for the conditional prior given $\mu_P$ and $w_P$
@@
 to 10. The prior limits for $m_0$ were set to the minimum and maximum value
-in the dataset, which is not strictly a valid strategy.\label{tab:priors}}
+in the dataset, which is not strictly a valid strategy. The 
+Truncation intervals are specified with the $T(a,b)$ notation.\label{tab:priors}}
 \end{table*}

### 2014-12-30 Paper/ms.tex
invariance

@@
 by phase transitions because the exploration only makes use of likelihood
-{\it rankings}, rather than likelihood values themselves.
+{\it rankings}, rather than likelihood values themselves, and are therefore
+invariant under monotonic transformations of the likelihood function.
 Part of their output, the relationship between the

### 2014-12-30 Paper/ms.tex
Going through the whole paper for a final edit

@@
 to be multimodal and to have other awkward features. In
-addition, when $N$ is unknown, the marginal likelihood (or evidence), as a
-function of $N$, is required. Rather than doing separate runs with different
-values of $N$, we propose an alternative
+addition, when $N$ is unknown, the marginal likelihood (or evidence) as a
+function of $N$ is required. Rather than doing separate runs with different
+trial values of $N$, we propose an alternative
 approach using a trans-dimensional Markov Chain Monte Carlo method within
 Nested Sampling. The posterior distribution for $N$ can be obtained with a
-single run, which takes $\sim$ 15 minutes on a typical modern desktop computer.
-We apply the method to $\nu$ Oph and Gliese 581, finding moderate evidence
+single run. We apply the method to $\nu$ Oph and Gliese 581, finding moderate evidence
 for additional signals in
@@
 exoplanetary system from observational data can be challenging.
-In the particular case of radial velocity data,
+In the case of radial velocity data,
 the expected signal due to an exoplanet is periodic, and the goal is to
@@
 2011A&A...528L...5T, fengji}. Bayesian model fitting via Markov Chain Monte
-Carlo (MCMC) tends to be very computationally intensive, especially if we
+Carlo (MCMC) tends to be computationally intensive, especially if we
 want to calculate the posterior distribution for $N$, the number of planets.
@@
 posterior distribution in a single run.
-The exploration is done using Diffusive Nested Sampling
+In addition, rather than trying to sample the posterior distribution, 
+we use Diffusive Nested Sampling
 \citep[DNS][]{dnest} which replaces the posterior distribution with an alternative
@@
 at {\tt https://github.com/eggplantbren/Exoplanet} under the terms of the
-GNU General Public Licence version 3.
+GNU General Public Licence.
 
@@
 \end{eqnarray}
-where $p(D | \theta)$ is called the likelihood, once the actual dataset
-is substituted in. Note that some authors do not distinguish between a
+where $p(D | \theta)$ as a function of $\theta$ is called the likelihood,
+once the actual dataset has been substituted in. Note that some authors do not
+distinguish between a
 sampling distribution and a likelihood. Throughout this paper we use the term
@@
 \citep{skilling} and ideas related to thermodynamic integration
-\citep[e.g.][]{neal}.
+\citep[e.g.][]{neal}. Relationships between these methods are discussed by
+\citet{cameron} and~\citet{scott}.
 
@@
 simplest of these are birth-death moves. More complicated moves, such as
-split-and-merge \citep[e.g.][]{umstatter} are possible but usually not necessary.
+split-and-merge are possible but not always necessary.
+Trans-dimensional MCMC is a natural tool for a wide range of astronomical
+data analysis problems \citep[e.g.][]{umstatter, walmswell, starfield}.
 
@@
 time in either the slab or the spike, rather than mixing between the two)
-and will give a misleading estimate of the average log likelihood at that temperature.
-
-In data analysis problems, phase transitions can exist when the data contains
+and will give a misleading estimate of the average log likelihood at that temperature
+and therefore an incorrect marginal likelihood estimate.
+
+Phase transitions are well known in statistical mechanics, but can also occur
+in Bayesian data analysis. Typically this occors when the data contains
 a ``big'' effect which provides a lot of information about some parameters, and
-a ``small'' subtle effect.
+a ``small'' subtle effect as well.
 Nested Sampling, and variants such as DNS, are not affected
@@
 
-We assigned hierarchical priors to some of the planet's parameters. This allows the
-model to capture the idea that knowing the values of some planet's parameters
-provides some information about the parameters of another planet. Not using
-hierarchical priors usually implies a strong prior commitment to the hypothesis
-that the properties of the planets are spread out across the whole domain of
-possible values, which is not necessarily the case.
-
 The unknown parameters are:
 \begin{eqnarray}
-\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\psi}\}_{i=1}^N, m_0, \sigma_{\rm extra}, \nu\right\}
+\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\psi}\}_{i=1}^N, m_0,
+\sigma_{\rm extra}, \nu\right\}
 \end{eqnarray}
 where $N$ is the number of planets,
-$\boldsymbol{\alpha} = \{\mu_P, \sigma_P, \mu_A\}$ are the
+$\boldsymbol{\alpha} = \{\mu_P, \sigma_P, \mu_A\}$ are hyperparameters
 hyperparameters used to define the prior for the properties of the planets,
@@
 All of the model assumptions are specified in detail in
-Table~\ref{tab:priors}. Most of these priors were chosen to represent vague
+Table~\ref{tab:priors}. We assigned hierarchical priors to some of the planet's
+parameters (i.e. the prior for the planets' parameters is defined conditional
+on some hyperparameters). This allows the
+model to capture the idea that knowing the values of some planet's parameters
+provides some information about the parameters of another planet. Not using
+hierarchical priors usually implies a strong prior commitment to the hypothesis
+that the properties of the planets are spread out across the whole domain of
+possible values, which is not necessarily the case.
+
+Most of our priors were chosen to represent vague
 prior knowledge, rather than the judgement of an informed expert on extrasolar
@@
 periods, which is a biexponential distribution given a location parameter
-$\mu_P$ and scale parameter, or width, $w_P$.
+$\mu_P$ and a scale parameter $w_P$ which determines the width of the
+distribution\footnote{A biexponential distribution with location parameter
+$\mu$ and scale parameter $w$ has probability density function
+$p(x|\mu, w) = \frac{1}{2w} \exp\left(-\frac{|x - \mu|}{w}\right)$.}.
 Rather than assigning independent priors to the log periods, the hierarchical
@@
 
+\bibitem[\protect\citeauthoryear{Brewer, Foreman-Mackey, 
+\& Hogg}{2013}]{starfield} Brewer B.~J., Foreman-Mackey D., Hogg D.~W., 2013, AJ, 146, 7
+
 \bibitem[\protect\citeauthoryear{Brewer}{2014}]{rjobject} Brewer, B. J., 2014,
@@
 
+\bibitem[\protect\citeauthoryear{Cameron 
+\& Pettitt}{2014}]{cameron} Cameron E., Pettitt A., 2014,
+Statistical Science, Vol. 29, No. 3, 397-419
+
 \bibitem[\protect\citeauthoryear{Feroz 
@@
 \& Suto}{2005}]{ohta} Ohta Y., Taruya A., Suto Y., 2005, ApJ, 622, 1118
+
+\bibitem[\protect\citeauthoryear{Polson 
+\& Scott}{2014}]{scott} Polson N.~G., Scott J.~G., 2014, arXiv, arXiv:1409.3601
 
@@
 022001
+
+\bibitem[\protect\citeauthoryear{Walmswell et 
+al.}{2013}]{walmswell} Walmswell J.~J., Eldridge J.~J., Brewer 
+B.~J., Tout C.~A., 2013, MNRAS, 435, 2171
 

### 2014-12-31 Paper/ms.tex
More editing

@@
 
-The prior for the velocity semi-amplitudes $\{A_i\}$ is important and will affect the
-inferred number of planets $N$. For example, if the prior allows for very
-low amplitude planets to exist, the data will not rule this out, and the posterior
-for $N$ will include the possibility of large numbers of planets. The prior
-will influence how many of these low amplitude planets will exist: if we
+For the velocity semi-amplitudes $\{A_i\}$, we chose an exponential distribution
+given the hyperparameter $\mu_A$. Our prior for $\mu_A$ spans many orders of
+magnitude but expresses a slight preference for $\mu_A$ being of order unity,
+using a Cauchy distribution. The prior for the semi-amplitudes
+will influence how many of these low amplitude planets will be inferred: if we
 believe there are many, and the data are uninformative about low amplitude
@@
 
-We ran the MCMC on the simulated dataset to obtain samples from the
+We ran our algorithm on the simulated dataset to obtain samples from the
 posterior distribution. We obtained 520 posterior samples.
@@
 The posterior distribution suggests that $N$ could be anywhere from 6 to 10.
+
+\begin{figure}
+\includegraphics[scale=0.4]{Figures/fake_data_N.eps}
+\caption{The posterior distribution for the number of planets $N$ given the
+simulated dataset. The true number of planets was 7.
+\label{fig:fake_data_N}}
+\end{figure}
+
 The posterior distribution for the periods $\{P_i\}_{i=1}^N$ is shown in
@@
 \end{eqnarray}
+If a certain period is accurately measured (i.e. its probability of existence
+is close to 1 and the posterior distribution for its period is very narrow)
+then it will appear in Figure 3 with a height of $\sim 520$. If the uncertainty
+in the period is larger than the histogram bin width then the peak will be
+spread over several bins.
+
 The posterior distribution for the periods, shown in
@@
 probabilities for these peaks are not close to 1.
-
-\begin{figure}
-\includegraphics[scale=0.4]{Figures/fake_data_N.eps}
-\caption{The posterior distribution for the number of planets $N$ given the
-simulated dataset. The true number of planets was 7.
-\label{fig:fake_data_N}}
-\end{figure}
 
@@
 Sampling algorithms) for the $\nu$ Oph analysis.
-There are several phase transitions (concave-up regions) present. The one at
+There are several phase transitions (concave-up regions) present. The small one at
 $\log(X) \approx -70$ separates models which contain additional signals from
@@
 For the $\nu$ Oph data the information was $\mathcal{H} \approx 76.6$ nats
-(natural units) or 111 bits, so the posterior is roughly $e^{-76.6}$ times the
-volume of the prior.
+(natural units) or 111 bits, so the posterior occupies about $e^{-76.6}$ times
+the prior volume.
 
@@
 and shows strong evidence for at least eight periods
-($P(N \leq 7 | D) = 0.012$).
+($P(N \leq 7 | D) = 0.012$). Some authors \citep[e.g.][]{2011A&A...528L...5T}
+recommend that the probability of $N$ planets should be $\geq$ 150 times
+greater than the probability of $N-1$ planets existing before making a
+claim that $N$ planets have been definitively detected. Such a decision rule
+is presumably equivalent to a utility function where false positives are much
+worse than false negatives. We note that applying this rule to our results, we
+would assert $N=6$, even though this has a tiny probability.
+
 It is now recognised that
@@
 we find many more signals in the data than previous authors.
-
 By inspecting the posterior distribution for the periods
@@
 are well determined, corresponding to the known periods of Gliese 581 b, c, d,
-and e. The other ``periods'' are very uncertain. It is possible that a
+and e. The other ``periods'' are very uncertain. It is quite likely that a
 non-sinusoidal signal due to stellar activity is being modelled as several

### 2014-12-31 Paper/ms.tex
Slight tweak

@@
 single run, which provides posterior samples and an estimate of the marginal
-likelihood. By embedding the MCMC in a Nested Sampling algorithm, instead of
+likelihood. By using Diffusive Nested Sampling, instead of
 directly trying to sample the posterior distribution, we can overcome difficult
-features in the problem, such as phase transitions.
+features in the problem, such as phase transitions and (to some extent)
+multiple modes.
 

### 2014-12-31 Paper/ms.tex
mention orbital stability
Prior table names arguments

@@
 are not available in closed form for the normal distribution.
+Our prior for $w_P$, which controls the diversity of the log-periods, was
+uniform between 0.1 and 3, since it is unlikely that many planets have
+extremely similar or extremely different (over several orders of magnitude)
+orbital periods.
 
 For the velocity semi-amplitudes $\{A_i\}$, we chose an exponential distribution
-given the hyperparameter $\mu_A$. Our prior for $\mu_A$ spans many orders of
+given the hyperparameter $\mu_A$ which sets the mean of the exponential distribution.
+Our prior for $\mu_A$ spans many orders of
 magnitude but expresses a slight preference for $\mu_A$ being of order unity,
@@
 $N$		& Number of planets	& Uniform$(\{0, 1, ..., N_{\rm max}\})$\\
-$\mu_P$		&	Median orbital period	& $\log(\mu_P) \sim$ Cauchy$(5.9, 1)T(-15.3, 27.1)$\\
-$w_P$		&	Diversity of orbital periods & $w_P \sim$ Uniform$(0.1, 3)$\\
-$\mu_A$		&	Mean amplitude	& $\log(\mu_A) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
+$\mu_P$		&	Median orbital period (years)	& $\log(\mu_P) \sim$ Cauchy$({\rm location}=5.9, {\rm scale}=1)T(-15.3, 27.1)$\\
+$w_P$		&	Diversity of orbital periods & $w_P \sim$ Uniform$({\rm min}=0.1, {\rm max}=3)$\\
+$\mu_A$		&	Mean amplitude (metres per second)	& $\log(\mu_A) \sim$ Cauchy$({\rm location}=0, {\rm scale}=1)T(-21.2, 21.2)$\\
 \hline
 {\bf Planet Parameters}\\
-$P_i$		&	Orbital period	&	$\log(P_i) \sim $ Biexponential($\log(\mu_P), w_P$)\\
-$A_i$		&	Semi-mplitude of signal	& Exponential$(\textnormal{mean}=\mu_A)$\\
-$\phi_i$	&	Phase of signal	&	Uniform($0, 2\pi$)\\
-$e_i$		&	Eccentricity of orbit	&	Beta$(1, 3.1)T(0, 0.8)$\\
-$\omega_i$	&	Viewing angle	&	Uniform$(0, 2\pi)$\\
+$P_i$		&	Orbital period	&	$\log(P_i) \sim $ Biexponential$({\rm location}=\log(\mu_P), {\rm scale}=w_P$)\\
+$A_i$		&	Semi-amplitude of signal	& Exponential$(\textnormal{mean}=\mu_A)$\\
+$\phi_i$	&	Phase of signal	&	Uniform($\textnormal{min}=0, \textnormal{max}=2\pi$)\\
+$e_i$		&	Eccentricity of orbit	&	Beta$(\alpha=1, \beta=3.1)T(0, 0.8)$\\
+$\omega_i$	&	Viewing angle	&	Uniform$({\rm min}=0, {\rm max}=2\pi)$\\
 \hline
 {\bf Other}\\
-$m_0$		&	Constant DC offset	&	Uniform$(m_{0, \rm min}, m_{0, \rm max})$\\
-$\sigma_{\rm extra}$	& ``Extra noise'' parameter	& $\log(\sigma) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
-$\nu$		& Shape parameter for $t$-distribution for noise & $\log(\nu) \sim$ Uniform$(\log(0.01), \log(1000))$\\
+$m_0$		&	Constant DC offset	&	Uniform$({\rm min}=m_{0, \rm min}, {\rm max}=m_{0, \rm max})$\\
+$\sigma_{\rm extra}$	& ``Extra noise'' parameter	& $\log(\sigma) \sim$ Cauchy$({\rm location}=0, {\rm scale}=1)T(-21.2, 21.2)$\\
+$\nu$		& Shape parameter for $t$-distribution for noise & $\log(\nu) \sim$ Uniform$({\rm min}=\log(0.01), {\rm max}=\log(1000))$\\
 \hline
@@
 $Y_i$		& Radial velocity measurements	&
-		Student-$t\left(m(t_i), \sqrt{\sigma_i^2 + \sigma_{\rm extra}^2}, \nu\right)$
+		Student-$t\left({\rm location}=m(t_i), {\rm scale}=\sqrt{\sigma_i^2 + \sigma_{\rm extra}^2}, {\rm shape}=\nu\right)$
 \end{tabular}
@@
 to 10. The prior limits for $m_0$ were set to the minimum and maximum value
-in the dataset, which is not strictly a valid strategy. The 
-Truncation intervals are specified with the $T(a,b)$ notation.\label{tab:priors}}
+in the dataset, which is not strictly a valid strategy. Some of the distributions
+were truncated for numerical reasons, the truncated intervals are specified with the
+$T(a,b)$ notation.\label{tab:priors}}
 \end{table*}
@@
 oscillations due to stellar activity: any oscillations found in the dataset
-will be described as ``planets'' by the model. However it is interesting that
+will be described as ``planets'' by the model. Another important consideration
+is the physical stability of the orbital system which is ignored by this type
+of analysis \citep{orbital}. However it is interesting that
 we find many more signals in the data than previous authors.
@@
 \bibitem[Stephens(2000)]{birthdeath} Stephens, M., 2000, ``Bayesian analysis of mixture models with an unknown number of components-an alternative to reversible jump methods.'', Annals of Statistics, 40-74.
+
+\bibitem[\protect\citeauthoryear{T{\'o}th 
+\& Nagy}{2014}]{orbital} T{\'o}th Z., Nagy I., 2014, MNRAS, 442, 454
 

### 2015-01-26 Paper/ms.tex
ellipticity --> eccentricity
New references suggested by the referee

@@
 
-It is now well known that RV datasets can contain periodic signals
-resulting from stellar activity rather than planets \citep{2014Sci...345..440R}.
+It is well known that RV datasets can contain periodic signals
+resulting from stellar activity rather than planets,
+which can affect the conclusions we draw about
+exoplanet systems \citep[e.g.][]{2001A&A...379..279Q, 2007A&A...474..293B, 2014Sci...345..440R}.
 Therefore, it is important to develop models which attempt to distinguish
@@
 present paper. Rather, we consider the problem of inferring the number $N$ of
-Keplerian signals in an RV dataset in a computationally efficient way.
+Keplerian signals in an RV dataset in a computationally efficient way, under
+the simplifying assumption that only Keplerian signals are present in the data.
 
@@
 To save time, we pre-computed the properties of orbits as a function of
-ellipticity. We also made the standard assumption that the planets do not
+eccentricity. We also made the standard assumption that the planets do not
 interact, so the expected signal due to several planets is the sum of the
@@
 
+\bibitem[Bonfils et 
+al.(2007)]{2007A&A...474..293B} Bonfils, X., Mayor, M., Delfosse, X., et al.\ 2007, A\&A, 474, 293
+
 \bibitem[\protect\citeauthoryear{Brewer 
@@
 \bibitem[\protect\citeauthoryear{Brewer}{2014}]{rjobject} Brewer, B. J., 2014,
-submitted. ArXiv: 1411.3921
+ArXiv: 1411.3921
 
@@
 \& Scott}{2014}]{scott} Polson N.~G., Scott J.~G., 2014, arXiv, arXiv:1409.3601
+
+\bibitem[Queloz et 
+al.(2001)]{2001A&A...379..279Q} Queloz, D., Henry, G.~W., Sivan, J.~P., et al.\ 2001, A\&A, 379, 279
 

### 2015-01-26 Paper/ms.tex
Fixed minor issue with citation

@@
 including the radial velocity (RV) method \citep[e.g.][]{2012PASJ...64..135S},
-the transit method \citep[e.g.]{2014PNAS..11112647B},
+the transit method \citep[e.g.][]{2014PNAS..11112647B},
 and gravitational microlensing

### 2015-01-27 Paper/ms.tex
Occam's razor and nu Oph numbers

@@
 $\sigma_{\rm extra}$ and $\nu$ are parameters of the noise distribution
-which are discussed further below.
+which are discussed further below. Note that our parameter $\omega_i$ is
+standard, however $\phi_i$ is non-standard because we assert that $\phi_i=0$
+always implies the signal is at its maximum at $t=0$. Our parameter space is
+equivalent to the standard one, we are just using a different coordinate system.
 
@@
 moderate probability of existence) have periods of 36.11 $\pm$ 0.034 days,
-75.58 $\pm$ 0.80 days, and 1709 $\pm$ 183 days.
-The probability that
-at least one of these ``extra'' signals exists, given by the fraction of
-samples that have one of these signals, is 85\%.
+75.58 $\pm$ 0.80 days, and 1709 $\pm$ 183 days. As with any MCMC output,
+if we are interested in the probability of any proposition $Q$ (for example,
+``$Q \equiv$ a planet exists with period between 35 and 37 days''), we can
+calculate the proportion of the posterior samples for which $Q$ is true, which
+(if we have a lot of samples) is a Monte Carlo estimate of the posterior
+probability of $Q$. For $\nu$ Oph, we calculated the probability that
+at least one of these ``extra'' signals (beyond the two commonly accepted ones)
+exists, as 85\%.
 Given that they exist, their amplitudes are low, around 5-40 metres per second,
@@
 worse than false negatives. We note that applying this rule to our results, we
-would assert $N=6$, even though this has a tiny probability.
+would assert $N=6$, even though this has a very small posterior probability.
 
@@
 there are many possible sources for oscillations in a data set and not all
-such oscillations should be claimed as planets \citep{2014Sci...345..440R}. 
+such oscillations should be claimed as planets. 
 Our model cannot distinguish between oscillations due to planets and
@@
 oscillation \citep{gaussproc}, the number of periods detected may be reduced
-substantially. An example model fit for Gliese 581 is shown in
+substantially. With these kinds of models, the posterior distribution for $N$
+can also be influenced by the prior for the amplitudes $A_i$. Many authors
+assign independent broad priors to the amplitudes, and this causes the
+``Occam's razor'' penalty for adding extra signals to be quite strong. Since we
+use a hierarchical prior for the amplitudes, if some amplitudes are found to
+be low, $\mu_A$ will become small. When $\mu_A$ is small, it is likely that any
+extra signals will have small amplitudes, so the ``Occam's razor'' effect is
+weaker. An example model fit for Gliese 581 is shown in
 Figure~\ref{fig:gliese581}.
@@
 \end{figure}
+
+As with $\nu$ Oph, we can calculate the posterior probability of any hypothesis
+about Gliese 581 by computing the fraction of the posterior samples that have
+that property. The posterior probabilities for planets b, c, d, and e, are
+close to 1. The posterior probability a signal exists with $\log_{10}(P)
+\in [1.55, 1.57]$ is 88\%, and the probability for a signal with
+$\log_{10}(P) \in [2.6, 2.8]$ is 85\%.
 

### 2015-01-27 Paper/ms.tex
Jones et al reference

@@
 simplest of these are birth-death moves. More complicated moves, such as
-split-and-merge are possible but not always necessary.
+split-and-merge, are possible but not always necessary.
 Trans-dimensional MCMC is a natural tool for a wide range of astronomical
-data analysis problems \citep[e.g.][]{umstatter, walmswell, starfield}.
+data analysis problems \citep[e.g.][]{umstatter, walmswell, starfield, 2014arXiv1411.7447J}.
 
@@
 \& Hogg}{2014}]{fengji} Hou F., Goodman J., Hogg D.~W., 2014, arXiv, arXiv:1401.6128
+
+\bibitem[\protect\citeauthoryear{Jones, Kashyap, 
+\& van Dyk}{2014}]{2014arXiv1411.7447J} Jones D.~E., Kashyap V.~L., van Dyk D.~A., 2014, arXiv, arXiv:1411.7447
 

### 2015-01-27 Paper/ms.tex
A bit of reordering

@@
 (or other causes of other periodic
-signals), but only four have well determined periods. The analysis of both
+signals), but only 4-6 have well determined periods. The analysis of both
 of these datasets shows phase transitions exist which are difficult to
@@
 are well determined, corresponding to the known periods of Gliese 581 b, c, d,
-and e. The other ``periods'' are very uncertain. It is quite likely that a
+and e. The other ``periods'' are more uncertain. As with $\nu$ Oph, we can
+calculate the posterior probability of any hypothesis
+about Gliese 581 by computing the fraction of the posterior samples that have
+that property. The posterior probabilities for planets b, c, d, and e, are
+close to 1. The posterior probability a signal exists with $\log_{10}(P)
+\in [1.55, 1.57]$ is 88\%, and the probability for a signal with
+$\log_{10}(P) \in [2.6, 2.8]$ is 85\%.
+
+One possible explanation for the large number of inferred signals that a
 non-sinusoidal signal due to stellar activity is being modelled as several
@@
 \end{figure}
-
-As with $\nu$ Oph, we can calculate the posterior probability of any hypothesis
-about Gliese 581 by computing the fraction of the posterior samples that have
-that property. The posterior probabilities for planets b, c, d, and e, are
-close to 1. The posterior probability a signal exists with $\log_{10}(P)
-\in [1.55, 1.57]$ is 88\%, and the probability for a signal with
-$\log_{10}(P) \in [2.6, 2.8]$ is 85\%.
 

### 2015-01-27 Paper/ms.tex
A bit more discussion

@@
 non-sinusoidal signal due to stellar activity is being modelled as several
-periods \citep{astero}, and if the model were extended to include a ``stochastic''
-oscillation \citep{gaussproc}, the number of periods detected may be reduced
-substantially. With these kinds of models, the posterior distribution for $N$
+periods \citep[e.g. as happens in asteroseismology when sinusoidal models are
+used;][]{astero}, and if the model were extended to include a ``stochastic''
+oscillation \citep[e.g.][]{gaussproc}, the number of periods detected may be reduced
+substantially. Another contributing factor is the prior for the amplitudes.
+With these kinds of models, the posterior distribution for $N$
 can also be influenced by the prior for the amplitudes $A_i$. Many authors

### 2015-01-27 Paper/ms.tex
mention posterior prob of existence as calculated from samples

@@
 \end{eqnarray}
-If a certain period is accurately measured (i.e. its probability of existence
-is close to 1 and the posterior distribution for its period is very narrow)
+If a certain period is accurately measured (i.e. it appears in close to 100 of
+the posterior samples and the distribution for its period is very narrow)
 then it will appear in Figure 3 with a height of $\sim 520$. If the uncertainty
@@
 By inspecting the posterior distribution for the periods
-(Figure~\ref{fig:gliese581_periods}), we see that only four of the periods
-are well determined, corresponding to the known periods of Gliese 581 b, c, d,
-and e. The other ``periods'' are more uncertain. As with $\nu$ Oph, we can
+(Figure~\ref{fig:gliese581_periods}), we see that only 4 of the periods
+are well determined and have a posterior probability close to 1 (i.e. they
+are present in all samples), corresponding to the known periods of Gliese 581
+b, c, d, and e. The other ``periods'' are more uncertain.
+As with $\nu$ Oph, we can
 calculate the posterior probability of any hypothesis
@@
 With these kinds of models, the posterior distribution for $N$
-can also be influenced by the prior for the amplitudes $A_i$. Many authors
+can be influenced by the prior for the amplitudes $A_i$. Many authors
 assign independent broad priors to the amplitudes, and this causes the

### 2015-01-27 Paper/ms.tex
% sign

@@
 \end{eqnarray}
-If a certain period is accurately measured (i.e. it appears in close to 100 of
+If a certain period is accurately measured (i.e. it appears in close to 100\% of
 the posterior samples and the distribution for its period is very narrow)
